---
title: "03.5_Sparse_Modeling"
output: html_document
date: "2026-05-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Objectives

# GOAL: Use sparse regression (LASSO + SuSiE) to identify which
# geospatial covariates best explain daily DOC and NO3 concentrations
# across sites. Following Modelscapes workflow (Zenodo 17101921).

Script : 03.5_Sparse_Modeling.R
Project: rc_sfa-rc-3-wenas-meta
Purpose: Introduction to Sparse Modeling in R
          
Follows this workflow: https://zenodo.org/records/17101921
  # title: "Sparse Modeling Introduction and Practice"
  # author: "Modelscapes"
  # date: "2025-05-15"
  # output: html_document
  # self-contained: true
Inputs :
   - Output_for_analysis/03_merge_geospatial/03_master_merged.csv

 Output :
   - 


 # Authorship
 Author: Jake Cavaiani
 13 May 2026
 
#  Libraries 
```{r read in libraries}
# install.packages("devtools")
#  library(devtools)
#  devtools::install_github("DOI-USGS/streamMetabolizer")

rm(list=ls(all=T)) #this clears your Environment

library(pacman)
p_load(tidyverse,
       here,
       glmnet, # LASSO
       susieR,  # SuSiE
       corrplot, # covariate collinearity check
       janitor) # clean names


```

#  1. Read inputs 
```{r read in master file}
master_df <- read_csv(
  here("Output_for_analysis", "03_merge_geospatial",
       "03_master_merged.csv"),
  na = c("", "NA", "-9999", "N/A"),
  show_col_types = FALSE
)

```
# Define responses and covariates
```{r define variables}
# --- Response variables ---
response_vars <- c("DOC_Interp_mg_C_L", "NO3_Interp_mg_N_L")

# --- Metadata columns to EXCLUDE from covariates ---
id_vars <- c("Study_ID", "Comparison_ID", "Pair", "Site", "Sampling_Date",
             "latitude", "longitude", "comid", "fire_years_used",
             "fdate", "resolution", "gnis_id", "gnis_name", "reachcode",
             "flowdir", "ftype", "wbareatype", "hwtype", "rpuid", "vpuid",
             "huc2", "huc4", "huc6", "huc8", "huc10", "huc12",
             "Burn_Unburn", "Climate",     # categorical - handle separately if desired
             "Days_Between_Samples",       # observational metadata
             "burn_sev_NA")                # pivot artifact

# --- Covariate candidates = everything else numeric ---
candidate_covars <- master_df %>%
  dplyr::select(-all_of(c(response_vars, id_vars))) %>%
  dplyr::select(where(is.numeric)) %>%
  names()

length(candidate_covars)   # should be ~150+ geospatial vars

```

# 3 Aggregate to one row per site (important!)
```{r aggregate - site}
site_df <- master_df %>%
  group_by(Study_ID, Site, comid) %>%
  summarise(
    # Response: mean DOC and NO3 per site
    DOC_mean = mean(DOC_Interp_mg_C_L, na.rm = TRUE),
    NO3_mean = mean(NO3_Interp_mg_N_L, na.rm = TRUE),
    # Covariates: first value (they're constant per site)
    across(all_of(candidate_covars), ~ first(na.omit(.x))),
    .groups = "drop"
  ) %>%
  mutate(across(c(DOC_mean, NO3_mean), ~ if_else(is.nan(.x), NA_real_, .x)))

dim(site_df)

doc_site <- site_df %>% 
  select(Study_ID, DOC_mean) %>% 
  na.omit()

no3_site <- site_df %>% 
  select(Study_ID, NO3_mean) %>% 
  na.omit()

# ---- 2. Yearly (one row per site-year) ----
yearly_df <- master_df %>%
  mutate(Year = lubridate::year(Sampling_Date)) %>%
  group_by(Study_ID, Site, comid, Year) %>%
  summarise(
    DOC_mean = mean(DOC_Interp_mg_C_L, na.rm = TRUE),
    NO3_mean = mean(NO3_Interp_mg_N_L, na.rm = TRUE),
    # Time-varying covariate(s): take mean within year
    TSF_mean = mean(Time_Since_Fire, na.rm = TRUE),   
    # Static covariates: first non-NA
    across(all_of(setdiff(candidate_covars, "TSF")),
           ~ first(na.omit(.x))),
    n_obs = sum(!is.na(DOC_Interp_mg_C_L) | !is.na(NO3_Interp_mg_N_L)),
    .groups = "drop"
  ) %>%
  mutate(across(c(DOC_mean, NO3_mean), ~ if_else(is.nan(.x), NA_real_, .x)))

doc_site <- yearly_df %>% 
  select(Study_ID, DOC_mean, Year, Time_Since_Fire) %>% 
  na.omit()

no3_site <- yearly_df %>% 
  select(Study_ID, NO3_mean, Year, Time_Since_Fire) %>% 
  na.omit()



# ---- 3. Monthly (one row per site-year-month) ----
monthly_df <- master_df %>%
  mutate(Year  = lubridate::year(Sampling_Date),
         Month = lubridate::month(Sampling_Date),
         YearMonth = lubridate::floor_date(Sampling_Date, "month")) %>%
  group_by(Study_ID, Site, comid, Year, Month, YearMonth) %>%
  summarise(
    DOC_mean = mean(DOC_Interp_mg_C_L, na.rm = TRUE),
    NO3_mean = mean(NO3_Interp_mg_N_L, na.rm = TRUE),
    TSF_mean = mean(Time_Since_Fire, na.rm = TRUE),
    across(all_of(setdiff(candidate_covars, "TSF")),
           ~ first(na.omit(.x))),
    n_obs = sum(!is.na(DOC_Interp_mg_C_L) | !is.na(NO3_Interp_mg_N_L)),
    .groups = "drop"
  ) %>%
  mutate(across(c(DOC_mean, NO3_mean), ~ if_else(is.nan(.x), NA_real_, .x)))

doc_site <- monthly_df %>% 
  select(Study_ID,  DOC_mean, Year, Month, Time_Since_Fire) %>% 
  na.omit()

no3_site <- monthly_df %>% 
  select(Study_ID, NO3_mean, Year, Month, Time_Since_Fire) %>% 
  na.omit()

```

# 3.5 EXPLORATORY ANALYSIS: BOOT STRAPPING
```{r}
# Covariates of interest:
# Watershed area - areasqkm
# Maximum Watershed elevation smooth - maxelevsmo - should correlate with slope?
# Watershed Slope - slope
# 30-year max air temp - tmax8110ws
# 30-year mean air temp - tmean8110ws
# 30-year min air temp - tmin8110ws
# Forest cover 
  # pctconif2019ws - Evergreen
  # pctdecid2019ws - Deciduous 
  # pctmxfst2019ws - Mixed forest
# Urban cover:
  # pcturbhi2019ws - high intensity
  # pcturbmd2019ws - medium intensity 
  # pcturblo2019ws - low intensity
  # pcturbop2019ws - Developed open space
# Grassland cover - pctgrs2019ws
# Wetland cover
  # pcthbwet2019ws - Herbaceous Wetlands
  # pctwdwet2019ws - Woody wetlands
# Ag cover
  # pctcrop2019ws - cultivated crop
  # pcthay2019ws - Hay pasture
# OM mean - omws
# Mean Bedrock depth - rckdepws
# Carbonate Residual Material - pctcarbresidws
# Non-carbonate Residual material - pctnoncarbresidws
# Silicic Residual materia
# Glacier Till
  # pctglactilclayws - clayey glacial till
  # pctglactilcrsws - coarse glacial till
  # pctglactilloamws - loamy glacial till
# Saline Lake Sediment - pctsallakews
# Baseflow index - bfiws
# Soil permeability - permws
# Mean runoff - runoffws
# Watershed burn percentage - burn_percent_fire_year
# burn severity
  # burn_sev_high
  # burn_sev_mod
  # burn_sev_low
# TSF
set.seed(42)
n_boot <- 1000

na_frac <- colMeans(is.na(X_full))
keep_covars <- names(na_frac)[na_frac < 0.2]

X_keep <- site_df %>% select(all_of(keep_covars)) %>% as.matrix()
X_imp  <- apply(X_keep, 2, function(col) {
  col[is.na(col)] <- median(col, na.rm = TRUE); col
})

ok <- !is.na(site_df$NO3_mean)
X_ok <- X_imp[ok, , drop = FALSE]
y_ok <- site_df$NO3_mean[ok]

cat("n =", nrow(X_ok), "p =", ncol(X_ok), "unique y =", length(unique(y_ok)), "\n")

fit_orig <- cv.glmnet(X_ok, y_ok, alpha = 1)

beta_orig <- as.numeric(coef(fit_orig, s = "lambda.1se"))[-1]

# Steps 2–6 inside each bootstrap iteration
boot_betas <- matrix(NA, n_boot, length(candidate_covars))
colnames(boot_betas) <- candidate_covars

for (b in 1:n_boot) {
  resampled <- sample(sites, N, replace = TRUE)
  d <- purrr::map_dfr(resampled, ~ filter(site_df, Site == .x))
  X <- as.matrix(d[, candidate_covars])
  y <- d$NO3_mean
  ok <- complete.cases(X, y)
  if (sum(ok) < 10) next
  fit <- cv.glmnet(X[ok,], y[ok], alpha = 1, nfolds = 10)
  boot_betas[b, ] <- as.numeric(coef(fit, s = "lambda.1se"))[-1]
}

# Stability selection (right panel of Fig 6.4 in your chapter)
selection_prob <- colMeans(boot_betas != 0, na.rm = TRUE)

# Coefficient distributions (boxplots, left panel of Fig 6.4)
boot_betas_df <- as.data.frame(boot_betas) %>%
  tidyr::pivot_longer(everything(), names_to = "var", values_to = "coef")

# Plot
ggplot(boot_betas_df, aes(x = reorder(var, coef, median), y = coef)) +
  geom_boxplot(outlier.size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  labs(x = NULL, y = "Bootstrap LASSO coefficient",
       title = "Nonparametric cluster bootstrap (1000 reps)")


#### OPTION 2 ####

boot_betas_nested <- matrix(NA, n_boot, length(candidate_covars))

for (b in 1:n_boot) {
  resampled_sites <- sample(sites, N, replace = TRUE)
  
  d <- purrr::imap_dfr(resampled_sites, function(s, i) {
    master_df %>%
      filter(Site == s) %>%
      slice_sample(prop = 1, replace = TRUE) %>%   # temporal resample
      summarise(
        DOC_mean = mean(DOC_Interp_mg_C_L, na.rm = TRUE),
        NO3_mean = mean(NO3_Interp_mg_N_L, na.rm = TRUE),
        across(all_of(candidate_covars), ~ first(na.omit(.x)))
      ) %>%
      mutate(boot_site = paste0(s, "_", i))
  })
  
  X <- as.matrix(d[, candidate_covars])
  y <- d$NO3_mean
  ok <- complete.cases(X, y)
  if (sum(ok) < 10) next
  fit <- cv.glmnet(X[ok,], y[ok], alpha = 1, nfolds = 10)
  boot_betas_nested[b, ] <- as.numeric(coef(fit, s = "lambda.1se"))[-1]
}

```

# 4. Prep X and y (separately for DOC and NO3)
```{r prep - DOC}
# ---- DOC ----
doc_dat <- site_df %>%
  dplyr::select(Site, DOC_mean, all_of(candidate_covars)) %>%
  drop_na()                        # complete cases only

# Drop zero-variance and near-constant covariates
nzv <- doc_dat %>%
  dplyr::select(-Site, -DOC_mean) %>%
  summarise(across(everything(), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "variance") %>%
  filter(variance == 0 | is.na(variance)) %>%
  pull(var)

X_doc <- doc_dat %>%
  dplyr::select(-Site, -DOC_mean, -any_of(nzv)) %>%
  as.matrix() %>%
  scale()                          # standardize (LASSO best practice)

y_doc <- doc_dat$DOC_mean

message("DOC: ", nrow(X_doc), " sites x ", ncol(X_doc), " covariates")

```

```{r prep - NO3}
# ---- NO3 ----
no3_dat <- site_df %>%
  dplyr::select(Site, NO3_mean, all_of(candidate_covars)) %>%
  drop_na()

nzv_no3 <- no3_dat %>%
  dplyr::select(-Site, -NO3_mean) %>%
  summarise(across(everything(), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "variance") %>%
  filter(variance == 0 | is.na(variance)) %>%
  pull(var)

X_no3 <- no3_dat %>%
  dplyr::select(-Site, -NO3_mean, -any_of(nzv_no3)) %>%
  as.matrix() %>%
  scale()

y_no3 <- no3_dat$NO3_mean

message("NO3: ", nrow(X_no3), " sites x ", ncol(X_no3), " covariates")

```

# 5 Check collinearity (optional but useful)
```{r collinearity}
cor_mat <- cor(X_doc, use = "pairwise.complete.obs")
# Flag highly correlated pairs
high_cor <- which(abs(cor_mat) > 0.9 & abs(cor_mat) < 1, arr.ind = TRUE)
head(high_cor)
# Optional: corrplot(cor_mat, method = "color", tl.cex = 0.4)

```

# 6 Run LASSO — DOC
```{r lasso - DOC}
set.seed(42)
cv_doc <- cv.glmnet(x = X_doc, y = y_doc, alpha = 1)

plot(cv_doc)    # dashed lines show best lambda values [1]
abline(v = log(cv_doc$lambda.min), col = "red")
abline(v = log(cv_doc$lambda.1se), col = "blue")

best_lambda_doc <- cv_doc$lambda.1se   # more parsimonious than lambda.min

model_doc <- glmnet(x = X_doc, y = y_doc, alpha = 1,
                    lambda = best_lambda_doc)

coef_doc <- coef(model_doc)[-1]        # drop intercept
selected_doc <- which(coef_doc != 0)   # nonzero covariates [1]

doc_results <- tibble(
  covariate   = colnames(X_doc)[selected_doc],
  coefficient = coef_doc[selected_doc]
) %>% arrange(desc(abs(coefficient)))

doc_results

# deviance explained (pseudo-R2) [1]
model_doc$dev.ratio

```

# 7. Run LASSO — NO3
```{r lasso - NO3}
set.seed(42)
cv_no3 <- cv.glmnet(x = X_no3, y = y_no3, alpha = 1)
plot(cv_no3)

best_lambda_no3 <- cv_no3$lambda.1se
model_no3 <- glmnet(x = X_no3, y = y_no3, alpha = 1,
                    lambda = best_lambda_no3)

coef_no3 <- coef(model_no3)[-1]
selected_no3 <- which(coef_no3 != 0)

no3_results <- tibble(
  covariate   = colnames(X_no3)[selected_no3],
  coefficient = coef_no3[selected_no3]
) %>% arrange(desc(abs(coefficient)))

no3_results
model_no3$dev.ratio

```

# 8. Run SuSiE — DOC (complementary view)
```{r susie - DOC}
susie_doc <- susie(X = X_doc, y = y_doc,
                   L = 10)          # max number of "signals"

# Credible sets: groups of correlated covariates that likely contain an effect
susie_doc$sets$cs

# Posterior inclusion probabilities
pip_doc <- tibble(
  covariate = colnames(X_doc),
  pip       = susie_doc$pip
) %>% arrange(desc(pip))

head(pip_doc, 20)

```

```{r susie - NO3}
susie_no3 <- susie(X = X_no3, y = y_no3, L = 10)
susie_no3$sets$cs

pip_no3 <- tibble(
  covariate = colnames(X_no3),
  pip       = susie_no3$pip
) %>% arrange(desc(pip))

head(pip_no3, 20)

```

# 10. Compare LASSO and SuSiE results
```{r compare}
doc_compare <- full_join(
  doc_results %>% rename(lasso_coef = coefficient),
  pip_doc     %>% rename(susie_pip  = pip),
  by = "covariate"
) %>% arrange(desc(susie_pip))

no3_compare <- full_join(
  no3_results %>% rename(lasso_coef = coefficient),
  pip_no3     %>% rename(susie_pip  = pip),
  by = "covariate"
) %>% arrange(desc(susie_pip))

doc_compare
no3_compare

```

# 11. Visualize selected covariates
```{r plot results}


```



# ============================================================

# ============================================================


# ATTEMPT 2 #

# ============================================================

# ============================================================

```{r setup, message = FALSE}
library(pacman)
p_load(tidyverse,
       here,
       glmnet, # LASSO
       susieR,  # SuSiE
       corrplot, # covariate collinearity check
       janitor) # clean names
```

## 1. Load data

```{r load-data}
master_df <- read_csv(here("Output_for_analysis", "03_merge_geospatial",
                           "03_master_merged.csv"),
                      show_col_types = FALSE)

dim(master_df)
```

## 2. Define responses and covariates

SuSiE requires a matrix X (each column = covariate, each row = observation) and a vector y of response values [1].

```{r define-vars}
# --- Response variables ---
response_vars <- c("DOC_Interp_mg_C_L", "NO3_Interp_mg_N_L")

# --- Metadata columns to EXCLUDE from covariates ---
id_vars <- c("Study_ID", "Comparison_ID", "Pair", "Site", "Sampling_Date",
             "latitude", "longitude", "comid", "fire_years_used",
             "fdate", "resolution", "gnis_id", "gnis_name", "reachcode",
             "flowdir", "ftype", "wbareatype", "hwtype", "rpuid", "vpuid",
             "huc2", "huc4", "huc6", "huc8", "huc10", "huc12",
             "Burn_Unburn", "Climate",     # categorical - handle separately if desired
             "Days_Between_Samples",       # observational metadata
             "burn_sev_NA")                # pivot artifact

# --- Covariate candidates = everything else numeric ---
candidate_covars <- master_df %>%
  dplyr::select(-all_of(c(response_vars, id_vars))) %>%
  dplyr::select(where(is.numeric)) %>%
  names()

candidate_covars <- (c("Area_watershed_km", "Time_Since_Fire", "burn_sev_high", "burn_sev_mod", "burn_sev_low",
                       "streamorde", "maxelevraw", "pctconif2019ws", "pcthbwet2019ws", "omws", ))

length(candidate_covars)   # should be ~150+ geospatial vars
```

## 3. Aggregate to one row per site

Your data is ~51k rows because it's daily. Since geospatial covariates are constant within a site, we'll aggregate to one row per site. Start simple, iterate later.

```{r aggregate-site}
site_df <- master_df %>%
  group_by(Study_ID, Site, comid) %>%
  summarise(
    # Response: mean DOC and NO3 per site
    DOC_mean = mean(DOC_Interp_mg_C_L, na.rm = TRUE),
    NO3_mean = mean(NO3_Interp_mg_N_L, na.rm = TRUE),
    # Covariates: first value (constant per site)
    across(all_of(candidate_covars), ~ first(na.omit(.x))),
    .groups = "drop"
  ) %>%
  mutate(across(c(DOC_mean, NO3_mean), ~ if_else(is.nan(.x), NA_real_, .x)))

dim(site_df)

# ---- 2. Calculate NA counts for each candidate covariate ----
covar_completeness <- site_df %>%
  dplyr::select(all_of(candidate_covars)) %>%
  summarise(across(everything(),
                   list(
                     n_missing = ~ sum(is.na(.x)),
                     n_total   = ~ length(.x),
                     pct_complete = ~ round(100 * sum(!is.na(.x)) / length(.x), 1)
                   ),
                   .names = "{.col}__{.fn}")) %>%
  pivot_longer(everything(),
               names_to = c("covariate", "metric"),
               names_sep = "__",
               values_to = "value") %>%
  pivot_wider(names_from = metric, values_from = value) %>%
  arrange(desc(n_missing))

covar_completeness

# ---- 3. Summary: how many covariates at each completeness level? ----
completeness_summary <- covar_completeness %>%
  mutate(completeness_bin = case_when(
    pct_complete == 100          ~ "100% (complete)",
    pct_complete >= 95           ~ "95-99%",
    pct_complete >= 90           ~ "90-94%",
    pct_complete >= 75           ~ "75-89%",
    pct_complete >= 50           ~ "50-74%",
    TRUE                          ~ "<50%"
  )) %>%
  count(completeness_bin, name = "n_covariates") %>%
  arrange(desc(n_covariates))

completeness_summary

# ---- 4. Visualize completeness distribution ----
p_completeness <- covar_completeness %>%
  ggplot(aes(x = pct_complete)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  geom_vline(xintercept = 100, linetype = "dashed", color = "red") +
  labs(title = "Covariate completeness across sites",
       subtitle = paste0(nrow(site_df), " total sites, ",
                         length(candidate_covars), " candidate covariates"),
       x = "Percent of sites with non-NA value",
       y = "Number of covariates") +
  theme_minimal(base_size = 11)

p_completeness
message("Dropped: ", length(candidate_covars) - length(candidate_covars_complete))

# ---- 5. Identify which sites are driving NAs ----
# Which sites have the most missing covariate data?
site_missingness <- site_df %>%
  dplyr::select(Site, all_of(candidate_covars)) %>%
  rowwise() %>%
  mutate(n_missing_covars = sum(is.na(c_across(all_of(candidate_covars))))) %>%
  ungroup() %>%
  dplyr::select(Site, n_missing_covars) %>%
  arrange(desc(n_missing_covars))

site_missingness

# ---- 6. Down-select covariates to those with 100% coverage ----
complete_covars <- covar_completeness %>%
  filter(pct_complete > 50) %>%
  pull(covariate)

message("Covariates with 100% coverage: ", length(complete_covars),
        " of ", length(candidate_covars))

# Optional: less strict thresholds
covars_95 <- covar_completeness %>% filter(pct_complete >= 95) %>% pull(covariate)
covars_90 <- covar_completeness %>% filter(pct_complete >= 90) %>% pull(covariate)

message("Covariates with >=95% coverage: ", length(covars_95))
message("Covariates with >=90% coverage: ", length(covars_90))

# ---- 7. Preview the complete covariates ----
tibble(
  covariate = complete_covars,
  category = case_when(
    str_detect(complete_covars, "^qa_|^qc_|^qe_|^va_|^vc_|^ve_") ~ "NHD Flow",
    str_detect(complete_covars, "ws$") & str_detect(complete_covars, "^pct") ~ "StreamCat land cover/lithology",
    str_detect(complete_covars, "ws$") ~ "StreamCat soil/climate",
    str_detect(complete_covars, "^burn") ~ "Fire (StreamCat MTBS)",
    str_detect(complete_covars, "^huc") ~ "HUC",
    TRUE ~ "Other NHD"
  )
) %>%
  count(category, name = "n_covariates") %>%
  arrange(desc(n_covariates))

# ---- 9. Redefine candidate_covars to use only complete ones ----
# This is what gets passed to LASSO / SuSiE going forward
candidate_covars_complete <- complete_covars
```

## 4. Prep X and y (separately for DOC and NO3)

LASSO/SuSiE don't handle NAs — we need complete cases per response [1].

```{r prep-doc}
# ---- DOC ----
doc_dat <- site_df %>%
  dplyr::select(Site, DOC_mean, all_of(candidate_covars_complete)) %>%
  drop_na()

# Drop zero-variance covariates (SuSiE and scale() can't handle them)
nzv <- doc_dat %>%
  dplyr::select(-Site, -DOC_mean) %>%
  summarise(across(everything(), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "variance") %>%
  filter(variance == 0 | is.na(variance)) %>%
  pull(var)

X_doc <- doc_dat %>%
  dplyr::select(-Site, -DOC_mean, -any_of(nzv)) %>%
  as.matrix() %>%
  scale()

y_doc <- doc_dat$DOC_mean

message("DOC: ", nrow(X_doc), " sites x ", ncol(X_doc), " covariates")
```

```{r prep-no3}
# ---- NO3 ----
no3_dat <- site_df %>%
  dplyr::select(Site, NO3_mean, all_of(candidate_covars_complete)) %>%
  drop_na()

nzv_no3 <- no3_dat %>%
  dplyr::select(-Site, -NO3_mean) %>%
  summarise(across(everything(), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "variance") %>%
  filter(variance == 0 | is.na(variance)) %>%
  pull(var)

X_no3 <- no3_dat %>%
  dplyr::select(-Site, -NO3_mean, -any_of(nzv_no3)) %>%
  as.matrix() %>%
  scale()

y_no3 <- no3_dat$NO3_mean

message("NO3: ", nrow(X_no3), " sites x ", ncol(X_no3), " covariates")
```

## 5. Check collinearity (optional but useful)

SuSiE's group selection handles correlated variables well, but it's still good to know how collinear your covariates are.

```{r collinearity}
cor_mat <- cor(X_doc, use = "pairwise.complete.obs")
high_cor <- which(abs(cor_mat) > 0.9 & abs(cor_mat) < 1, arr.ind = TRUE)
head(high_cor)
corrplot(cor_mat, method = "color", tl.cex = 0.4)
```

## 6. LASSO — DOC

```{r lasso-doc}
set.seed(42)
cv_doc <- cv.glmnet(x = X_doc, y = y_doc, alpha = 1)

plot(cv_doc)
abline(v = log(cv_doc$lambda.min), col = "red")
abline(v = log(cv_doc$lambda.1se), col = "blue")

best_lambda_doc <- cv_doc$lambda.1se   # more parsimonious than lambda.min

model_lasso_doc <- glmnet(x = X_doc, y = y_doc, alpha = 1,
                          lambda = best_lambda_doc)

coef_doc <- coef(model_lasso_doc)[-1]          # drop intercept
selected_doc <- which(coef_doc != 0)

doc_lasso_results <- tibble(
  covariate   = colnames(X_doc)[selected_doc],
  coefficient = coef_doc[selected_doc]
) %>% arrange(desc(abs(coefficient)))

doc_lasso_results
model_lasso_doc$dev.ratio     # pseudo-R2
```

## 7. LASSO — NO3

```{r lasso-no3}
set.seed(42)
cv_no3 <- cv.glmnet(x = X_no3, y = y_no3, alpha = 1)
plot(cv_no3)

best_lambda_no3 <- cv_no3$lambda.1se
model_lasso_no3 <- glmnet(x = X_no3, y = y_no3, alpha = 1,
                          lambda = best_lambda_no3)

coef_no3 <- coef(model_lasso_no3)[-1]
selected_no3 <- which(coef_no3 != 0)

no3_lasso_results <- tibble(
  covariate   = colnames(X_no3)[selected_no3],
  coefficient = coef_no3[selected_no3]
) %>% arrange(desc(abs(coefficient)))

no3_lasso_results
model_lasso_no3$dev.ratio
```

## 8. SuSiE — DOC

SuSiE is a Bayesian method similar to forward step-wise selection. It handles correlated variables through group selection, producing "credible sets" — groups of correlated variables that likely contain a true effect but can't be distinguished individually [1].

```{r susie-doc}
model_susie_doc <- susie(X = X_doc, y = y_doc, L = 10)

# Summary with credible sets and inclusion probabilities [1]
summary(model_susie_doc)

# Posterior inclusion probabilities [1]
pip_doc <- tibble(
  covariate = colnames(X_doc),
  pip       = model_susie_doc$pip
) %>% arrange(desc(pip))

head(pip_doc, 20)

# Credible sets: groups of correlated covariates likely containing a true effect [1]
model_susie_doc$sets$cs

# R-squared for SuSiE model [1]
rsq_doc <- cor(predict(model_susie_doc, X_doc), y_doc)^2
rsq_doc
```

```{r susie-doc-plots}
# SuSiE's built-in plot - color-coded by credible set [1]
susie_plot(model_susie_doc, y = "PIP")

# Coefficient plot
plot(coef(model_susie_doc)[-1],
     pch = 20,
     main = "SuSiE covariates for DOC",
     ylab = "Effect on DOC concentration",
     xlab = "Explanatory variables")
abline(h = 0, col = "gray")
# Label variables with notable effects
big_effects <- which(abs(coef(model_susie_doc)[-1]) > 0.1)
if(length(big_effects) > 0){
  text(x = big_effects,
       y = coef(model_susie_doc)[-1][big_effects],
       labels = colnames(X_doc)[big_effects],
       pos = 3, cex = 0.7)
}

coef_df <- tibble(
  covariate = colnames(X_doc),
  coef      = coef(model_susie_doc)[-1],
  index     = seq_along(coef)
) %>%
  mutate(label = if_else(abs(coef) > 0.1, covariate, NA_character_))

ggplot(coef_df, aes(x = index, y = coef)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(size = 2) +
  geom_text(aes(label = label), nudge_y = 0.02, size = 3, na.rm = TRUE) +
  labs(title = "SuSiE covariates for DOC",
       x = "Explanatory variables",
       y = "Effect on DOC concentration") +
  theme_minimal()
```

## 9. SuSiE — NO3

```{r susie-no3}
model_susie_no3 <- susie(X = X_no3, y = y_no3, L = 10)
summary(model_susie_no3)

pip_no3 <- tibble(
  covariate = colnames(X_no3),
  pip       = model_susie_no3$pip
) %>% arrange(desc(pip))

head(pip_no3, 20)

model_susie_no3$sets$cs

rsq_no3 <- cor(predict(model_susie_no3, X_no3), y_no3)^2
rsq_no3
```

```{r susie-no3-plots}
susie_plot(model_susie_no3, y = "PIP")

plot(coef(model_susie_no3)[-1],
     pch = 20,
     main = "SuSiE covariates for NO3",
     ylab = "Effect on NO3 concentration",
     xlab = "Explanatory variables")
abline(h = 0, col = "gray")
big_effects_no3 <- which(abs(coef(model_susie_no3)[-1]) > 0.1)
if(length(big_effects_no3) > 0){
  text(x = big_effects_no3,
       y = coef(model_susie_no3)[-1][big_effects_no3],
       labels = colnames(X_no3)[big_effects_no3],
       pos = 3, cex = 0.7)
}

library(ggplot2)

# ---- SuSiE PIP plot (built-in) ----
susie_plot(model_susie_no3, y = "PIP")

# ---- Coefficient plot with ggplot ----
coef_df_no3 <- tibble(
  covariate = colnames(X_no3),
  coef      = coef(model_susie_no3)[-1],
  index     = seq_along(coef)
) %>%
  mutate(label = if_else(abs(coef) > 0.1, covariate, NA_character_))

p_susie_coef_no3 <- ggplot(coef_df_no3, aes(x = index, y = coef)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(size = 2) +
  geom_text(aes(label = label), nudge_y = 0.02, size = 3, na.rm = TRUE) +
  labs(title = "SuSiE covariates for NO3",
       x = "Explanatory variables",
       y = "Effect on NO3 concentration") +
  theme_minimal()

p_susie_coef_no3



```

## 10. Compare LASSO and SuSiE results

```{r compare}
doc_compare <- full_join(
  doc_lasso_results %>% rename(lasso_coef = coefficient),
  pip_doc           %>% rename(susie_pip  = pip),
  by = "covariate"
) %>% arrange(desc(susie_pip))

no3_compare <- full_join(
  no3_lasso_results %>% rename(lasso_coef = coefficient),
  pip_no3           %>% rename(susie_pip  = pip),
  by = "covariate"
) %>% arrange(desc(susie_pip))

doc_compare
no3_compare
```
tmin8110ws: lasso_coef = NA, susie_pip = 0.55 → LASSO zeroed it out, but SuSiE gives it moderate support.
pctsallakews: lasso_coef = 2.78, susie_pip = 0.20 → LASSO selected it, SuSiE is less confident.
bfiws: lasso_coef = NA, susie_pip = 0.03 → neither method gave it much weight, but it appears because SuSiE reports PIPs for every covariate.

## 11. Visualize selected covariates

```{r plot-lasso-doc, fig.width=8, fig.height=6}
p_lasso_doc <- doc_lasso_results %>%
  mutate(covariate = fct_reorder(covariate, coefficient)) %>%
  ggplot(aes(x = coefficient, y = covariate,
             fill = coefficient > 0)) +
  geom_col() +
  scale_fill_manual(values = c("firebrick", "steelblue"),
                    labels = c("Negative", "Positive"),
                    name = "Direction") +
  labs(title = "LASSO-selected covariates for DOC",
       subtitle = paste0("Deviance explained: ",
                         round(model_lasso_doc$dev.ratio, 3)),
       x = "Standardized coefficient",
       y = NULL) +
  theme_minimal(base_size = 11)

p_lasso_doc
```

```{r plot-lasso-no3, fig.width=8, fig.height=6}
p_lasso_no3 <- no3_lasso_results %>%
  mutate(covariate = fct_reorder(covariate, coefficient)) %>%
  ggplot(aes(x = coefficient, y = covariate,
             fill = coefficient > 0)) +
  geom_col() +
  scale_fill_manual(values = c("firebrick", "steelblue"),
                    labels = c("Negative", "Positive"),
                    name = "Direction") +
  labs(title = "LASSO-selected covariates for NO3",
       subtitle = paste0("Deviance explained: ",
                         round(model_lasso_no3$dev.ratio, 3)),
       x = "Standardized coefficient",
       y = NULL) +
  theme_minimal(base_size = 11)

p_lasso_no3
```

```{r plot-susie-pip, fig.width=8, fig.height=6}
# SuSiE PIP bar charts - top 20 variables for each response
p_pip_doc <- pip_doc %>%
  slice_max(pip, n = 20) %>%
  mutate(covariate = fct_reorder(covariate, pip)) %>%
  ggplot(aes(x = pip, y = covariate)) +
  geom_col(fill = "darkorange") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "gray40") +
  labs(title = "SuSiE posterior inclusion probabilities - DOC",
       subtitle = paste0("Model R2 = ", round(rsq_doc, 3)),
       x = "Posterior inclusion probability (PIP)",
       y = NULL) +
  xlim(0, 1) +
  theme_minimal(base_size = 11)

p_pip_no3 <- pip_no3 %>%
  slice_max(pip, n = 20) %>%
  mutate(covariate = fct_reorder(covariate, pip)) %>%
  ggplot(aes(x = pip, y = covariate)) +
  geom_col(fill = "darkorange") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "gray40") +
  labs(title = "SuSiE posterior inclusion probabilities - NO3",
       subtitle = paste0("Model R2 = ", round(rsq_no3, 3)),
       x = "Posterior inclusion probability (PIP)",
       y = NULL) +
  xlim(0, 1) +
  theme_minimal(base_size = 11)

p_pip_doc
p_pip_no3
```

```{r plot-method-comparison, fig.width=8, fig.height=6}
# Highlight covariates selected by BOTH methods
p_compare_doc <- doc_compare %>%
  mutate(
    selected_by = case_when(
      !is.na(lasso_coef) & susie_pip > 0.5 ~ "Both",
      !is.na(lasso_coef)                    ~ "LASSO only",
      susie_pip > 0.5                       ~ "SuSiE only",
      TRUE                                  ~ "Neither"
    )
  ) %>%
  filter(selected_by != "Neither") %>%
  mutate(covariate = fct_reorder(covariate, susie_pip)) %>%
  ggplot(aes(x = susie_pip, y = covariate, color = selected_by)) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Both" = "purple",
                                "LASSO only" = "steelblue",
                                "SuSiE only" = "darkorange")) +
  labs(title = "Method agreement for DOC covariates",
       x = "SuSiE posterior inclusion probability",
       y = NULL,
       color = "Selected by") +
  theme_minimal(base_size = 11)

p_compare_doc

p_compare_no3 <- no3_compare %>%
  mutate(
    selected_by = case_when(
      !is.na(lasso_coef) & susie_pip > 0.5 ~ "Both",
      !is.na(lasso_coef)                    ~ "LASSO only",
      susie_pip > 0.5                       ~ "SuSiE only",
      TRUE                                  ~ "Neither"
    )
  ) %>%
  filter(selected_by != "Neither") %>%
  mutate(covariate = fct_reorder(covariate, susie_pip)) %>%
  ggplot(aes(x = susie_pip, y = covariate, color = selected_by)) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Both" = "purple",
                                "LASSO only" = "steelblue",
                                "SuSiE only" = "darkorange")) +
  labs(title = "Method agreement for NO3 covariates",
       x = "SuSiE posterior inclusion probability",
       y = NULL,
       color = "Selected by") +
  theme_minimal(base_size = 11)

p_compare_no3
```

## 12. Save outputs

```{r save-outputs}
out_dir <- here("Output_for_analysis", "03.5_Sparse_Modeling")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

# --- Tables ---
write_csv(doc_lasso_results, file.path(out_dir, "doc_lasso_selected.csv"))
write_csv(no3_lasso_results, file.path(out_dir, "no3_lasso_selected.csv"))
write_csv(pip_doc,           file.path(out_dir, "doc_susie_pip.csv"))
write_csv(pip_no3,           file.path(out_dir, "no3_susie_pip.csv"))
write_csv(doc_compare,       file.path(out_dir, "doc_lasso_vs_susie.csv"))
write_csv(no3_compare,       file.path(out_dir, "no3_lasso_vs_susie.csv"))

# --- Save aggregated site-level data (useful for later re-runs) ---
write_csv(site_df, file.path(out_dir, "site_level_master_df.csv"))

# --- Plots ---
ggsave(file.path(out_dir, "doc_lasso_coefficients.png"),
       p_lasso_doc, width = 8, height = 6, dpi = 150)
ggsave(file.path(out_dir, "no3_lasso_coefficients.png"),
       p_lasso_no3, width = 8, height = 6, dpi = 150)
ggsave(file.path(out_dir, "doc_susie_pip.png"),
       p_pip_doc, width = 8, height = 6, dpi = 150)
ggsave(file.path(out_dir, "no3_susie_pip.png"),
       p_pip_no3, width = 8, height = 6, dpi = 150)
ggsave(file.path(out_dir, "doc_method_comparison.png"),
       p_compare_doc, width = 8, height = 6, dpi = 150)
ggsave(file.path(out_dir, "no3_method_comparison.png"),
       p_compare_no3, width = 8, height = 6, dpi = 150)

# --- Save fitted model objects (for reuse / inspection) ---
saveRDS(model_lasso_doc, file.path(out_dir, "model_lasso_doc.rds"))
saveRDS(model_lasso_no3, file.path(out_dir, "model_lasso_no3.rds"))
saveRDS(model_susie_doc, file.path(out_dir, "model_susie_doc.rds"))
saveRDS(model_susie_no3, file.path(out_dir, "model_susie_no3.rds"))

message("All results saved to: ", out_dir)
```

## 13. Summary report

```{r summary}
cat("============================================================\n")
cat("SPARSE MODELING SUMMARY\n")
cat("============================================================\n\n")

cat("--- DOC ---\n")
cat("Sites with complete data:", nrow(X_doc), "\n")
cat("Candidate covariates:", ncol(X_doc), "\n")
cat("LASSO-selected covariates:", nrow(doc_lasso_results), "\n")
cat("LASSO deviance explained:", round(model_lasso_doc$dev.ratio, 3), "\n")
cat("SuSiE R-squared:", round(rsq_doc, 3), "\n")
cat("SuSiE credible sets:", length(model_susie_doc$sets$cs), "\n")
cat("SuSiE PIP > 0.5 covariates:", sum(pip_doc$pip > 0.5), "\n\n")

cat("--- NO3 ---\n")
cat("Sites with complete data:", nrow(X_no3), "\n")
cat("Candidate covariates:", ncol(X_no3), "\n")
cat("LASSO-selected covariates:", nrow(no3_lasso_results), "\n")
cat("LASSO deviance explained:", round(model_lasso_no3$dev.ratio, 3), "\n")
cat("SuSiE R-squared:", round(rsq_no3, 3), "\n")
cat("SuSiE credible sets:", length(model_susie_no3$sets$cs), "\n")
cat("SuSiE PIP > 0.5 covariates:", sum(pip_no3$pip > 0.5), "\n")
```

## 14. Next steps (optional extensions)

Based on the Modelscapes workflow, consider these follow-ups [1]:

1. **Cross-validation for predictive performance**: Split sites into train/test folds, evaluate RMSE for LASSO vs SuSiE vs linear regression to see which method generalizes best [1].

2. **Inspect credible sets closely**: When SuSiE puts multiple variables in one credible set, it means they're correlated and the model can't distinguish them — useful domain info [1].

3. **Revisit daily structure**: Instead of site-level means, try mixed-effects LASSO (`glmmLasso`) to use the full daily time series with site as a random effect.

4. **Stratify by `Burn_Unburn`**: Run separate models for burned vs unburned sites to see if covariate importance differs.

5. **Log-transform responses**: DOC and NO3 concentrations are often log-normal — try `y_doc <- log(doc_dat