---
title: "Variance"
author: "JC & AMP"
date: "`r Sys.Date()`"
output: html_document
---
Compare effect size variances across all watersheds between normalized/non-normalized data

Chunk 2180 - 2439 is an attempt to do it myself. It looks like # HYPOTHESIS i SUPPORTED (overall): Area normalization significantly reduces effect-size variance. Watershed area explains a meaningful share of cross-watershed variability. Area normalization explains 62% [49%, 72%] of the raw effect-size variance. Anything over 50% tells us that area is the dominant source of cross-watershed variability. When I break it up by nitrate and DOC for looking at variance difference by response variable: both Variables are positive = area normalization captures watershed variance. When looking at variance RATIO by response variable (log scale: both variables >1 = normalization shrinks variance.

```{r Read in final data frame with non-normal and normalized}
rm(list=ls(all=T)) #this clears your Environment

library(pacman)
p_load(tidyverse,
       here,
       forecastML, 
       zoo,
       # hrbrthemes,
       viridis, 
       ggpmisc,
       RobustLinearReg,
       lme4,
       car)

# Read in effect size data frame for normalized and non-normalized data
ES_nn_normal <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_size_orig_nonnormal.csv"))

ES_nn_normal <- ES_nn_normal %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      Study_ID == "Abbott et al. 2021" & occurrence == 3 ~ "10", # Third value
      Study_ID == "Abbott et al. 2021" & occurrence == 4 ~ "11", # Fourth value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire)) %>% 
  dplyr::select(Study_ID:Effect_size_no_watershed)

# Make dataframe long format
ES_long <- ES_nn_normal %>%
  mutate(row_id = row_number()) %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed),
    names_to  = "type",
    values_to = "effect_size"
  ) %>%
  mutate(type = factor(type))

head(ES_long)

```
```{r prep var}
# Levene's Test of all data ####
cat("=== Standard Levene's Test ===\n")
levene_obs <- leveneTest(effect_size ~ type, data = ES_long, center = median)
print(levene_obs)
obs_F <- levene_obs$`F value`[1]

# Quick look at the observed variances
cat("\nVar(Effect_size NORMALIZED):", var(ES_nn_normal$Effect_size), "\n")
cat("Var(Effect_size NON-NORMALIZED):", var(ES_nn_normal$Effect_size_no_watershed), "\n")
cat("Variance ratio:", 
    var(ES_nn_normal$Effect_size) / var(ES_nn_normal$Effect_size_no_watershed), "\n\n")
```

```{r bootstrap}

# BOOTSTRAPPING ####
set.seed(2024)
n_boot <- 10000

# Pivot (back to) to wide so each row_id is one observation with both the raw and area-normalized effect size
ES_wide <- ES_long %>%
  pivot_wider(
    id_cols  = c(Study_ID, response_var, Climate, Biome,
                 Effect_size_pair, Time_Since_Fire, row_id),
    names_from  = type,
    values_from = effect_size
  ) %>%
  drop_na(Effect_size, Effect_size_no_watershed)

# paired bootstrap of variance difference & ratio Resamples row_ids (whole pairs) to keep the pairing intact
boot_var_compare <- function(df, n_boot = 10000) {
  n <- nrow(df)
  replicate(n_boot, {
    bs  <- slice_sample(df, n = n, replace = TRUE)
    v_r <- var(bs$Effect_size,              na.rm = TRUE)
    v_n <- var(bs$Effect_size_no_watershed, na.rm = TRUE)
    c(var_raw = v_r, var_norm = v_n,
      diff = v_r - v_n,              # + → normalization shrinks variance
      ratio = v_r / v_n)            # >1 → normalization shrinks variance
  }) %>% t() %>% as_tibble()
}

# OVERALL comparison (all response variables pooled)
boot_overall <- boot_var_compare(ES_wide, n_boot)

overall_summary <- boot_overall %>%
  pivot_longer(everything(), names_to = "metric") %>%
  group_by(metric) %>%
  summarise(
    mean    = mean(value),
    median  = median(value),
    ci_lo   = quantile(value, 0.025),
    ci_hi   = quantile(value, 0.975),
    .groups = "drop"
  )

overall_summary %>% filter(metric %in% c("diff", "ratio"))

# Quick programmatic check
if (overall_summary %>%
      filter(metric == "diff") %>%
      pull(ci_lo) > 0) {
  cat("✔ HYPOTHESIS i SUPPORTED (overall):
      Area normalization significantly reduces effect-size variance.
      Watershed area explains a meaningful share of cross-watershed variability.\n")
} else if (overall_summary %>%
             filter(metric == "diff") %>%
             pull(ci_hi) < 0) {
  cat("✔ HYPOTHESIS ii SUPPORTED (overall):
      Area normalization increases effect-size variance.
      Additional watershed characteristics likely modulate fire response.\n")
} else {
  cat("⚠ INCONCLUSIVE (overall):
      95% CI of the variance difference spans zero.\n")
}
```

```{r next}
# HYPOTHESIS i SUPPORTED (overall): Area normalization significantly reduces effect-size variance. Watershed area explains a meaningful share of cross-watershed variability.

# Proportion of raw variance "removed" by normalization
boot_overall <- boot_overall %>%
  mutate(prop_explained = (var_raw - var_norm) / var_raw)

prop_summary <- boot_overall %>%
  summarise(
    mean_prop   = mean(prop_explained),
    ci_lo       = quantile(prop_explained, 0.025),
    ci_hi       = quantile(prop_explained, 0.975)
  )

cat(sprintf(
  "\nArea normalization explains %.0f%% [%.0f%%, %.0f%%] of the raw effect-size variance.\n",
  prop_summary$mean_prop * 100,
  prop_summary$ci_lo     * 100,
  prop_summary$ci_hi     * 100
))

# Area normalization explains 62% [49%, 72%] of the raw effect-size variance.

# 0.50 -> Area is a dominant source of cross-watershed variability 
# 0.20 –> 0.50 -> Area is meaningful but other factors matter too
# 0.20 -> Area alone is a weak first-order correction 
# 0 -> Normalization actually inflates variance → hypothesis ii 

boot_by_response <- boot_by_response %>%
  mutate(
    direction = case_when(
      boot_diff_ci_lo > 0  ~ "i  – norm reduces variance",
      boot_diff_ci_hi < 0  ~ "ii – norm increases variance",
      TRUE                 ~ "inconclusive"
    ),
    prop_explained = obs_diff / obs_var_raw
  )

boot_by_response %>%
  select(response_var, n, obs_var_raw, obs_var_norm,
         obs_diff, prop_explained,
         boot_diff_ci_lo, boot_diff_ci_hi, direction) %>%
  arrange(desc(prop_explained)) %>%
  print(width = Inf)

# One-sided "p-value": proportion of resamples where raw > normalised
p_overall <- mean(boot_overall$diff > 0)

cat("\n===== OVERALL VARIANCE COMPARISON =====\n")
print(overall_summary)
cat(sprintf("\nP(var_raw > var_norm) = %.3f\n", p_overall))

# BY RESPONSE VARIABLE ####
boot_by_response <- ES_wide %>%
  group_by(response_var) %>%
  group_split() %>%
  map_dfr(function(df) {
    rv    <- unique(df$response_var)
    n_obs <- nrow(df)
    b     <- boot_var_compare(df, n_boot)

    tibble(
      response_var       = rv,
      n                  = n_obs,
      obs_var_raw        = var(df$Effect_size,              na.rm = TRUE),
      obs_var_norm       = var(df$Effect_size_no_watershed, na.rm = TRUE),
      obs_diff           = obs_var_raw - obs_var_norm,
      obs_ratio          = obs_var_raw / obs_var_norm,
      boot_diff_mean     = mean(b$diff),
      boot_diff_ci_lo    = quantile(b$diff, 0.025),
      boot_diff_ci_hi    = quantile(b$diff, 0.975),
      boot_ratio_mean    = mean(b$ratio),
      boot_ratio_ci_lo   = quantile(b$ratio, 0.025),
      boot_ratio_ci_hi   = quantile(b$ratio, 0.975),
      p_raw_gt_norm      = mean(b$diff > 0)
    )
  })

cat("\n===== BY RESPONSE VARIABLE =====\n")
print(boot_by_response, width = Inf)

# PLOTS
# 5a – Overlapping densities of bootstrapped variances (overall)
boot_overall %>%
  select(var_raw, var_norm) %>%
  pivot_longer(everything(),
               names_to  = "type",
               values_to = "variance") %>%
  mutate(type = recode(type,
    var_raw  = "Raw effect size",
    var_norm = "Area-normalised effect size")) %>%
  ggplot(aes(variance, fill = type)) +
  geom_density(alpha = 0.45, colour = "grey30") +
  labs(title = "Bootstrapped variance distributions (all responses pooled)",
       x = "Variance of effect sizes", y = "Density", fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")

# 5b – Density of variance DIFFERENCE with zero reference line
boot_overall %>%
  ggplot(aes(diff)) +
  geom_density(fill = "steelblue", alpha = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red", linewidth = 0.8) +
  annotate("text", x = 0, y = Inf, vjust = 2, hjust = -0.1,
           label = paste0("P(diff > 0) = ", round(p_overall, 3)),
           colour = "red", size = 4.5) +
  labs(title = "Bootstrapped variance difference (Raw − Normalised)",
       subtitle = "Positive -> normalization reduces variability  |  Negative -> normalization increases variability",
       x = "Var(raw) − Var(normalized)", y = "Density") +
  theme_minimal(base_size = 13)

# 5c – Forest-style plot: variance difference by response variable
boot_by_response %>%
  ggplot(aes(x = reorder(response_var, obs_diff), y = obs_diff)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "red") +
  geom_pointrange(aes(ymin = boot_diff_ci_lo, ymax = boot_diff_ci_hi),
                  size = 0.6) +
  coord_flip() +
  labs(title = "Variance difference by response variable",
       subtitle = "Positive = area normalization captures watershed variance\nNegative = additional modulators likely",
       x = NULL,
       y = "Var(raw) − Var(normalized)  [95 % bootstrap CI]") +
  theme_minimal(base_size = 13)

# Both Variables are Positive = area normalization captures watershed variance

# 5d – Variance RATIO by response variable (log scale)
boot_by_response %>%
  ggplot(aes(x = reorder(response_var, obs_ratio), y = obs_ratio)) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "red") +
  geom_pointrange(aes(ymin = boot_ratio_ci_lo, ymax = boot_ratio_ci_hi),
                  size = 0.6) +
  scale_y_log10() +
  coord_flip() +
  labs(title = "Variance ratio by response variable",
       subtitle = ">1 = normalisation shrinks variance  |  <1 = normalisation inflates variance",
       x = NULL,
       y = "Var(raw) / Var(normalised)  [95 % CI, log scale]") +
  theme_minimal(base_size = 13)

# Both variables >1 = normalization shrinks variance
```
