---
title: "Map"
output: html_document
date: "2023-10-19"
editor_options: 
  chunk_output_type: console
---
The purpose of this script is create the map figure 

Script Workflow:

Step 1) Load in the summary metadata sheet that include lat and long

Step 2) Load in the shape files of the burns

Step 3) Plot


# Status: in progress

# ==============================================================================
# Author: Jake Cavaiani 
# 19 October 2023
# ==============================================================================

## Load packages and set working directory
```{r Jake/Mac}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

library(ggplot2)
library(tidyverse)
library(here)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
library(broom)
library(usmap)
library(maps)
library(mapdata)
library(cowplot)
library(sf)
library(RColorBrewer)
library(patchwork)
```

This is the workflow to read in the meta analysis studies and the fires they are studying and then loading in a CONUS map and plotting our data on top of that
```{r load in fire names and associated study}
meta <- read_csv(here("inputs", "Studies_Summary", "Map_input.csv")) # reading in the finalized filtered meta-analysis papers. 

coords <- meta %>% 
  select(Fire_name, latitude, longitude, number_of_burn, State, group) # Picking only the columns I want 

coords_noNA <- coords %>% 
    drop_na(Fire_name) # eliminating rows that are NAs if they dont have an associated fire name with it. 
 
# Create an sf object from the data
coordinates_sf <- st_as_sf(coords_noNA, coords = c("longitude", "latitude"), crs = 4326)

```

```{r - US map with AK as an inset map}
# Download the US states shapefile from the Census Bureau
url_us_states <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url_us_states, "inputs/map_shape_files/us_states.zip", mode = "wb")
unzip("inputs/map_shape_files/us_states.zip", exdir = "inputs/map_shape_files/us_states")

# Download the Alaska shapefile
url_alaska <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url_alaska, "inputs/map_shape_files/alaska.zip", mode = "wb")
unzip("inputs/map_shape_files/alaska.zip", exdir = "inputs/map_shape_files/alaska")

# Read the US states shapefile
us_states <- st_read("inputs/map_shape_files/us_states/cb_2021_us_state_20m.shp")

# Read the Alaska shapefile
alaska <- st_read("inputs/map_shape_files/alaska/cb_2021_us_state_20m.shp")

ak <- ggplot() +
  geom_sf(data = alaska, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  xlim(-180, -127) +
  ylim(50, 72) + 
  theme_void() +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(),
        legend.position = "none")
  

print(ak)

# I want to include Canada # 
# Download shapefile for world countries
world <- ne_countries(scale = "medium", returnclass = "sf")

# Filter for Canada
canada <- world[world$iso_a2 == "CA", ]
canadian_provinces <- sf::st_as_sf(rnaturalearth::ne_states(country = "Canada", returnclass = "sf"))


# US and Canada plot with all the fires
us_ca <- ggplot() +
  geom_sf(data = canadian_provinces, fill = "white", color = "black") +
  geom_sf(data = us_states, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  scale_color_manual(values = c("#666666", "#A6CEE3", "#00AFBB", "#1B9E77",
                                "#7570B3", "#E7298A", "darkred", "black", 
                                "#1F78B4", "#E31A1C", "#33A02C", "black",
                                "darkblue", "darkgreen", "blue", "yellow", "red",
                                "green", "purple")) +
  xlim(-135, -68) +
  ylim(25, 62) +
  theme_void() +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(),
        legend.position = "none")

print(us_ca)

# add the AK insert #
us_ca + inset_element(ak, 0, -1.0, 0.3, 1.3, align_to = 'full')
  
ggsave("number_watersheds_map.pdf",
       path = here("initial_plots", "10_Map"),
       width = 6, height = 4,  units = "in")

```

```{r Chat GPT to get legends on top of each other}
set.seed(42)
df <- data.frame(
  x = rnorm(100),
  y = rnorm(100),
  group = rep(c("Group A", "Group B"), each = 50),
  color = rep(c("Red", "Blue"), each = 50),
  shape = rep(c("Circle", "Square"), each = 50)
)

p <- ggplot(df, aes(x, y)) +
  geom_point(aes(color = color, shape = shape)) +
  scale_color_manual(values = c("Red" = "red", "Blue" = "blue")) +
  scale_shape_manual(values = c("Circle" = 16, "Square" = 15)) +
  labs(title = "Example Plot") +
  theme(
    legend.position = c(0.9, 0.1),  # Adjust the position of the color legend
    legend.justification = c(1, 0),  # Right-justify the color legend
    legend.box.just = "right"  # Place the color legend outside the plot area
  )
 
print(p)

```






