---
title: "11_Random_forest"
output: html_document
date: "2024-02-07"
editor_options: 
  chunk_output_type: console
---

The purpose of this script is to run random forests for DOC and NO3 effect sizes using catchment characteristics

Script Workflow:

Step 1) Load in the geospatial csv that is generated from pulling in stream cat data

Step 2) Load in the fire name and year summary csv

Step 3) Load in the effect size sheet that comes from the All_Studies_Temporal_Normalization script that includes effect size for each watershed. 




# Status: in progress

# ==============================================================================
# Author: Peter Regier
#  February 2024
# ==============================================================================

# Library download # 
```{r Load packages}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

require(pacman)
library(tidyverse)
library(janitor)
library(PNWColors)
library(ranger)
library(here)

```

# random forest dataframes # 
```{r read in csv's}

## Read in effect datasets
no3 <- read_csv(here("Output_for_analysis", "08_Meta_summary_stats", "no3_meta_random_forest.csv")) %>% 
  clean_names() %>% 
  rename("dependent" = effect_size) %>% 
  select(-c(minelevraw, elevfixed))

doc <- read_csv(here("Output_for_analysis", "08_Meta_summary_stats", "doc_meta_random_forest.csv")) %>% 
  clean_names() %>% 
  rename("dependent" = effect_size) %>% 
  select(-c(minelevraw, elevfixed))

## Set ggplot theme
theme_set(theme_bw())

```

# 2. Random Forest models 
```{r function to calculate variable importance}
## Function to calculate variable importance for a default ranger() model
plot_fi <- function(data){
  
  ## Make model
  rf_model <- ranger(dependent ~ ., data = data, importance = "impurity")
  
  ## Visualize model output
  print(rf_model)
  
  ## Set vectors for renaming stuff
  var_names <- rf_model$variable.importance
  col_names <- c("predictor", "raw_fi")
  
  ## Convert feature importance to a tibble with variables as a column
  fi0 <- as.data.frame(var_names) %>% 
    tibble::rownames_to_column() %>% 
    as_tibble()
  
  ## Rename columns
  colnames(fi0) = col_names
  
  ## Output variable importance (or feature importance)
  fi0
}

```

# Plot FI for effect sizes
```{r plot Feature importance}
## Create a dataset of variable importance
df <- bind_rows(plot_fi(no3) %>% mutate(analyte = "no3"), 
          plot_fi(doc) %>% mutate(analyte = "doc")) %>% 
  group_by(analyte) %>% 
  mutate(fi = raw_fi / max(raw_fi))

write_csv(df, here("Output_for_analysis", "11_Random_forest", "Random_Forest_Effect_Size_Fig_data.csv"))


## Set up a color palette
var_colors <- PNWColors::pnw_palette("Bay", n = length(unique(df$predictor)))

## Make plots
# burn_percentage - Burn percentage 
# Climate - Climate classification 
# maxelevsmo - Maximum elevation (smoothed) in centimeters
# minelevsmo - Minimum elevation (smoothed) in centimeters
# slope - Slope of flowline (meters/meters) based on smoothed elevations
# time_since_fire - time since fire 
# totdasqkm -Total upstream catchment area from downstream end of flowline.

dose.labs <- c("DOC", "Nitrate")
names(dose.labs) <- c("doc", "no3")

ggplot(df, aes(fi, 
               reorder(predictor, fi), fill = predictor)) + 
  geom_col(alpha = 0.8, show.legend = F, width = 0.7) + 
  facet_wrap(~analyte, nrow = 1, labeller = labeller(analyte = dose.labs)) + 
  scale_fill_manual(values = var_colors) + 
  labs(x = "Feature Importance", 
       y = "", fill = "") +
  scale_y_discrete(labels=c("burn_percentage" = "Burn Percentage", 
                            "climate" = "Climate",
                            "maxelevsmo" = "Max Elevation",
                            "minelevsmo" = "Minimum Elevation",
                            "slope" = "Average Slope",
                            "time_since_fire" = "Time Since Fire",
                            "totdasqkm" = "Catchment Area"))
                            

ggsave("Effect_Size_Random_Forest.pdf",
       path = here("initial_plots", "11_Meta_random_forest"),
       width = 5.5, height = 3,  units = "in")


```







