---
title: "Wenas_Effect_Size_Models"
output:
  html_document: default
  word_document: default
  pdf_document: default
date: "2023-11-16"
editor_options:
  chunk_output_type: console
---

The purpose of this script is to perform models for effect size against burn characterizations 

Script Workflow:

Step 1) Load in the geospatial csv that is generated from pulling in stream cat data

Step 2) Load in the fire name and year summary csv

Step 3) Load in the effect size sheet that comes from the All_Studies_Temporal_Normalization script that includes effect size for each watershed. 

Our model:Effect_size ~ burn_percentage*TSF + (1|Site)


# Status: in progress

# ==============================================================================
# Author: Jake Cavaiani 
# 16 November 2023
# ==============================================================================

## Geospatial data
```{r Jake/Mac Load Packages}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment


library(tidyverse)
library(here)
library(forecastML)
library(zoo)
library(hrbrthemes)
library(viridis)
library(readxl)
library(lme4)
library(GGally)
library(merTools)
library(ggcorrplot)
library(ggpmisc)
library(RColorBrewer)
library(ggbreak)
library(lmerTest)
library(car)
```

```{r load in csv's}
options(max.print = 200)

# Geospatial csv from stream cat with only the sites and the total area + the burn percentages 
geospatial <- read_csv(here("R_scripts", "Geospatial_data_2023-11-13.csv")) %>% 
  dplyr::select(site, comid, totdasqkm, streamorde, MTBS_1986Ws, MTBS_1987Ws, MTBS_1988Ws, MTBS_1989Ws, MTBS_1990Ws, MTBS_1991Ws, MTBS_1992Ws, MTBS_1993Ws, MTBS_1994Ws, MTBS_1995Ws, MTBS_1996Ws, MTBS_1997Ws, MTBS_1998Ws, MTBS_1999Ws, MTBS_2000Ws, MTBS_2001Ws, MTBS_2002Ws, MTBS_2003Ws, MTBS_2004Ws, MTBS_2005Ws, MTBS_2006Ws, MTBS_2007Ws, MTBS_2008Ws, MTBS_2009Ws, MTBS_2010Ws, MTBS_2011Ws, MTBS_2012Ws, MTBS_2013Ws, MTBS_2014Ws, MTBS_2015Ws, MTBS_2016Ws, MTBS_2017Ws)

# Site summary of each study and its respective fire and the year of that fire
studies_data <- read_excel(here("inputs", "StudiesData_Table1.xlsx"),
    sheet = "Study_info_filtered_V3_Q") %>% 
  dplyr::select(Study, Fire_name, Fire_year)

# The effect sizes of each study's watershed
effect_size <- read_csv(here("Output_for_analysis", "Effect_Size.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, Effect_size, Site, Time_Since_Fire)

effect_size_doc <- effect_size %>% 
  filter(response_var == "DOC_Interp")

effect_size_no3 <- effect_size %>% 
  filter(response_var == "NO3_Interp")

# Read in the study and site csv 
study_site_data <- read_excel(here("inputs", "Studies_Summary", "Fire_name_Lat_Long.xlsx")) %>% 
  dplyr::select(Study, Site, Fire_year, Effect_size_site) %>% 
  rename(site = Site, 
         Study_ID = Study, 
         year = Fire_year, 
         Site = Effect_size_site) %>% 
  na.omit()

# pivoting to make the responses in one column
geospatial_long <- geospatial %>%
  pivot_longer(
    cols = MTBS_1986Ws:MTBS_2017Ws,
    names_to = "year",
    values_to = "burn_percentage",
    values_drop_na = TRUE
  )

```

### Data summary visualization 
```{r - histogram of data summary}
ggplot(geospatial, aes(x = streamorde)) +
  geom_histogram(binwidth = 1, color = "black", fill = "#56B4E9") +
  theme_bw()

ggsave("Effect_Size_stream_order.pdf",
       path = here("initial_plots", "03_Wenas_effect_size"),
       width = 10, height = 8, units = "in")

```

```{r - plotting MTBS fire years }
ggplot(geospatial_long, aes(x = seq_along(year), y = burn_percentage)) +
  geom_line() +
  geom_point() +
  facet_wrap(~site, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


```{r - manually check fire years vs. MTBS data}

fire_clip <- right_join(geospatial_long, study_site_data) # 64 watersheds

fire_clip <- fire_clip %>% 
  na.omit() 
# The na.omit function is removing sites that dont have MTBS data which is due to out of country/too old or too recent!
# 51 watersheds now. 

######################### I am gonna manually check each one! 
# Akokala_Creek: 13.8379 is correct 
# Arroyo Hondo: 16.4842 is correct 
# Benjamin Slough: doesnt have the correct burn year from the study to the MTBS year
# Bowman Creek: 5.7229 is correct 
# Camas Creek is from Tidemann paper which was in 1973 with a burn from 1970 so there is not MTBS data 
# Coal: 55.8481 is correct 
# Coal_Creek: 0 is correct; from Hauer & Spencer from a burn in 1988. 
# Cold Creek is from Hickenbottom at is from too recent of a burn to have MTBS data 
# Control: The paper lists 74 % but ours says 42.78. 
# Crane Creek is from the Neary & Currier paper which investigates a burn from 1978. No MTBS data 
# Crow creek: 9.1827 is correct. 
# Dry Creek: 0 is correct
# DS1: 16.3095 is correct 
# DS2: 27.1144 is correct 
# DS3: 27.1121 is correct 
# East Fork: doesnt exist in here because its in Canada
# Fish Creek: 72.8600.....is the only MTBS data and is correct
# Gaviota: 12.9690....is the only MTBS data and is correct 
# Grade creek is from Tiedemann paper which investigates a 1970 fire so there will be no MTBS data
# Hobble Creek Upper: 0 is correct; doesnt have the correct burn year from the study (2018) to the MTBS years given
# Hobble Creek Lower: 0 is correct; doesnt have the correct burn year from the study (2018) to the MTBS years given
# Jones creek: 96.1969 is correct
# Logging Creek: 1.5154 is correct
# Lower Mcdonald Creek: 31.3921 is correct
# Middle_Fork is from the Hickenbottom paper which investigates a burn from 2021. Dont think we have that updated MTBS data yet
# Mill Race: 0 is correct; doesnt have the correect burn year from the study (2018) to the MTBS year given 
# Mitsubishi Race: 0 is correct; doesnt have the correect burn year from the study (2018) to the MTBS year given
# North Fork is from the Hickenbottom paper which investigates a burn from 2021. Dont think we have that updated MTBS data yet
# Notawohka Creek: doesnt have MTBS data because the site is located within Canada
# Payson doesnt have the correect burn year from the study (2018) to the MTBS year given 
# PBR: 0 is correct 
# Pinchot: 0.2982....is the only MTBS data and it is correct 
# PNF: 16.7187 is correct
# Provo river: 0 is correct; doesnt have the correect burn year from the study (2018) to the MTBS year given 
# PSF: 1.5243...is the only MTBS data and it is correct 
# Quartz Creek: 27.0952....is the only MTBS data and it is correct 
# Rattlesnake is a control site so should be 0 *
# Red_Bench_Creek: 100....is the only MTBS data and it is correct
# Red_Meadow_Creek: 24.0870....is the only MTBS data and it is correct 
# Reference: Supposed to be 0....but it lists it as 70.0648 which is the same as the Wragg which is not correct *
# Rocky Fire: 12.8141 is correct 
# San Onofre: 98.9419 is correct
# Site 1: 76.4673 is correct 
# Site 2: 76.4673 is correct 
# Site 3: 6.7889 is correct
# Spanish Fork Lower: 0 is correct
# Spanish Fork Upper: 0 is correct 
# Trout Creek: In the Hickenbottom paper so its too new for MTBS data 
# US1: 0 is correct;
# US2: Doesnt have the correct MTBS data 
# Upper MacDonald Creek: 24.9105 is correct 
# Wally Creek: 93.5391 is correct 
# Wash Branch is in the Neary & Currier paper which investigates a 1978 fire so there is no MTBS data 
# Wragg Fire: 70.0648 is correct 


fire_clip <- fire_clip %>% 
  mutate(burn_percentage = case_when(site == 'Benjamin Slough' ~ 67,
                                     site == 'Payson' ~ 90,
                                     site == 'Spanish Fork Lower' ~ 24,
                                     site == 'Spanish Fork Upper' ~ 25,
                                     site == 'Reference' ~ 0,
                                     TRUE ~ burn_percentage))
# These papers report the burn percentages so I manually put them in because the Streamcat data did not pull the proper data. 

```

### Burn percentage plots 
```{r plot DOC}
doc_effect_size_burn_merged <- right_join(effect_size_doc, fire_clip) %>% 
  na.omit() 

give.n <- function(x){
  return(c(y = median(x)*1.05, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}

ggpairs(doc_effect_size_burn_merged, 
        columns = c("Effect_size", "burn_percentage", "Time_Since_Fire")) +
  theme_bw() +
  ggtitle("Effect Size - DOC")
  
  
brewer.pal(n = 9, name = "YlOrRd")
brewer.pal(n = 9, name = "YlGn")
brewer.pal(n = 8, name = "Dark2")

ggpairs(doc_effect_size_burn_merged,
        columns = c("Effect_size", "burn_percentage"), 
        ggplot2::aes(color = Time_Since_Fire, alpha = 0.5)) +
  scale_color_manual(values = c("#800026", "#FC4E2A",  "#FF7F00", "#A6761D", "#78C679", "#006837")) +
  scale_fill_manual(values = c("#800026", "#FC4E2A",  "#FF7F00", "#A6761D", "#78C679", "#006837")) +
  theme_bw() +
  ggtitle("Effect Size - DOC")

ggsave("Effect_Size_Burn_Percentage_DOC.pdf",
       path = ("~/GitHub/rc_sfa-rc-3-wenas-meta/FireMeta_Rproj/initial_plots/03_Wenas_effect_size"),
       width = 10, height = 8, units = "in")


# Burn percentage
ggplot(data = doc_effect_size_burn_merged, aes(x = burn_percentage, y = Effect_size)) +
  stat_poly_line() +
  stat_poly_eq() +
  geom_boxplot() +
  theme_bw()

ggplot(data = doc_effect_size_burn_merged, aes(x = burn_percentage, y = Effect_size, color = Time_Since_Fire)) +
  stat_poly_line() +
  stat_poly_eq() +
  geom_point() +
  scale_color_manual(values = c("#800026", "#FC4E2A",  "#FF7F00", "#A6761D", "#78C679", "#006837")) +
  theme_bw() +
  theme(legend.position = "none")





```

### Burn percentage models
```{r Effect size models-DOC}
# DOC #
# Effect_size ~ burn_percentage*TSF + (1|Site)

ggpairs(doc_effect_size_burn_merged,
        columns = c("burn_percentage", "Time_Since_Fire"))

doc_burn_mixed = lmer(Effect_size ~ burn_percentage + (1 | site), data = doc_effect_size_burn_merged)

doc_burn_mixed

summary(doc_burn_mixed)
fixef(doc_burn_mixed)
confint(doc_burn_mixed)

ranef(doc_burn_mixed)$site %>% 
  head(5)

coef(doc_burn_mixed)$site %>% 
  head(5)

predictInterval(doc_burn_mixed)

REsim(doc_burn_mixed) 

plotREsim(REsim(doc_burn_mixed))

# anova #
doc_aov = aov(Effect_size ~ burn_percentage, data = doc_effect_size_burn_merged)
plot(doc_aov)
summary(doc_aov)


```

```{r Effect size models-NO3}
# NO3 #
# Effect_size ~ burn_percentage*TSF + (1|Site)
no3_effect_size_burn_merged <- right_join(effect_size_no3, fire_clip) %>% 
  na.omit() 


ggpairs(no3_effect_size_burn_merged,
        columns = c("burn_percentage", "Time_Since_Fire"))

no3_burn_mixed = lmer(Effect_size ~ burn_percentage + (1 | site), data = no3_effect_size_burn_merged)

summary(no3_burn_mixed)
confint(no3_burn_mixed)

ranef(no3_burn_mixed)$site %>% 
  head(5)

coef(no3_burn_mixed)$site %>% 
  head(5)

predictInterval(no3_burn_mixed)

REsim(no3_burn_mixed) 

plotREsim(REsim(no3_burn_mixed))

ggpairs(no3_effect_size_burn_merged,
        columns = c("Effect_size", "burn_percentage"), 
        ggplot2::aes(color = Time_Since_Fire, alpha = 0.5)) +
  scale_color_manual(values = c("#800026", "#FC4E2A",  "#FF7F00", "#A6761D", "#78C679", "#006837")) +
  scale_fill_manual(values = c("#800026", "#FC4E2A",  "#FF7F00", "#A6761D", "#78C679", "#006837")) +
  theme_bw() +
  ggtitle("Effect Size - Nitrate")

ggsave("Effect_Size_Burn_Percentage_NO3.pdf",
       path = ("~/GitHub/rc_sfa-rc-3-wenas-meta/FireMeta_Rproj/initial_plots/03_Wenas_effect_size"),
       width = 10, height = 8, units = "in")




```


### Climate models
```{r DOC climate models}
doc_climate_mixed <-  lmer(Effect_size ~ Climate + (1 | site), data = doc_effect_size_burn_merged)

doc_climate_mixed

summary(doc_climate_mixed)
fixef(doc_climate_mixed)
confint(doc_climate_mixed)

predictInterval(doc_climate_mixed)

REsim(doc_climate_mixed) 

plotREsim(REsim(doc_climate_mixed))

```

```{r DOC levenes test}
doc_climate_levene <- effect_size_doc %>% 
  filter(Study_ID != "Abbott et al. 2021",
         Study_ID != "Rhea et al. 2021") # Eliminating studies that monitor sites that are greater than 10 years post-fire

# Null: the variance amoung the groups is equal. 
# conduct Levene's Test for equality of variances
leveneTest(Effect_size ~ Climate, data = doc_climate_levene) #F-value: 66.819, p-value: <2.2e-16, Df: 4, 44

# plot 

boxplot(Effect_size ~ Climate,
  data = doc_climate_levene,
  main = "Effect size Distribution by Climate - DOC",
  xlab = "Climate",
  ylab = "Effect Size",
  col = "steelblue",
  border = "black")
```
The p-value of the test is <2.2e-16, which is less than our significance level of 0.05. 
Thus, we reject the null hypothesis and conclude that the variance among the climate groups is NOT equal. 

From the plot, Csa (Hot-Mediterranean) is the most variable. Now we have to think about why?

```{r NO3 climate models}
no3_climate_mixed <-  lmer(Effect_size ~ Climate + (1 | site), data = no3_effect_size_burn_merged)

no3_climate_mixed

summary(no3_climate_mixed)
fixef(no3_climate_mixed)
confint(no3_climate_mixed)

predictInterval(no3_climate_mixed)

REsim(no3_climate_mixed) 

plotREsim(REsim(no3_climate_mixed))

```

```{r NO3 levenes test}
no3_climate_levene <- effect_size_no3 %>% 
  filter(Study_ID != "Abbott et al. 2021",
         Study_ID != "Rhea et al. 2021") # Eliminating studies that monitor sites that are greater than 10 years post-fire

# Null: the variance amoung the groups is equal. 
# conduct Levene's Test for equality of variances
leveneTest(Effect_size ~ Climate, data = no3_climate_levene) #F-value: 2.8984, p-value: 0.008966, Df: 7,88

# plot 

boxplot(Effect_size ~ Climate,
  data = no3_climate_levene,
  main = "Effect size Distribution by Climate - NO3",
  xlab = "Climate",
  ylab = "Effect Size",
  col = "steelblue",
  border = "black")

```
The p-value of the test is 0.008966, which is less than our significance level of 0.05. 
Thus, we reject the null hypothesis and conclude that the variance among the climate groups is NOT equal. 

"BSk" = "Cold semi-arid"
"Cfb" = "Subtropical highland",
"Csa" = "Hot-Mediterranean",
"Csb" = "Warm-Mediterranean",
"Dfb" = "Warm-humid",
"Dfc" = "Subarctic",

"Dsb" = "Mediterranean", 
"Cfa" = "Humid subtropical",


### TSF models
```{r DOC TSF models}
doc_tsf_mixed <-  lmer(Effect_size ~ Time_Since_Fire + (1 | site), data = doc_effect_size_burn_merged)

doc_tsf_mixed

summary(doc_tsf_mixed)
fixef(doc_tsf_mixed)
confint(doc_tsf_mixed)

predictInterval(doc_tsf_mixed)

REsim(doc_tsf_mixed) 

plotREsim(REsim(doc_tsf_mixed))

```

```{r DOC levenes test}
effect_size_doc$Time_Since_Fire <- factor(effect_size_doc$Time_Since_Fire, levels = c("0", "1", "2", "3", "4", "5", ">10")) # reordering to make sure that the >10 year bin is at the end of the plot to visualize through time. 

# Null: the variance amoung the groups is equal. 
# conduct Levene's Test for equality of variances
leveneTest(Effect_size ~ Time_Since_Fire, data = effect_size_doc) #F-value: 0.73 p-value: 0.6277, Df: 6, 48

# plot 

boxplot(Effect_size ~ Time_Since_Fire,
  data = effect_size_doc,
  main = "Effect size Distribution by Time Since Fire - DOC",
  xlab = "Time Since Fire",
  ylab = "Effect Size",
  col = "steelblue",
  border = "black") 

```
The p-value of the test is 0.73, which is greater than our significance level of 0.05. 
Thus, we do not reject the null hypothesis and conclude that the variance among the climate groups is equal. 


```{r NO3 TSF models}
no3_tsf_mixed <-  lmer(Effect_size ~ Time_Since_Fire + (1 | site), data = no3_effect_size_burn_merged)

no3_tsf_mixed

summary(no3_tsf_mixed)
fixef(no3_tsf_mixed)
confint(no3_tsf_mixed)

predictInterval(no3_tsf_mixed)

REsim(no3_tsf_mixed) 

plotREsim(REsim(no3_tsf_mixed))

```

```{r NO3 levenes test}
effect_size_no3$Time_Since_Fire <- factor(effect_size_no3$Time_Since_Fire, levels = c("0", "1", "2", "3", "4", "5", ">10")) # reordering to make sure that the >10 year bin is at the end of the plot to visualize through time. 

# Null: the variance amoung the groups is equal. 
# conduct Levene's Test for equality of variances
leveneTest(Effect_size ~ Time_Since_Fire, data = effect_size_no3) #F-value: 2.4422, p-value: 0.03067, Df: 6,95

# plot 

boxplot(Effect_size ~ Time_Since_Fire,
  data = effect_size_no3,
  main = "Effect size Distribution by Time Since Fire - NO3",
  xlab = "Time Since Fire",
  ylab = "Effect Size",
  col = "steelblue",
  border = "black")


```
The p-value of the test is 0.03067, which is less than our significance level of 0.05. 
Thus, we reject the null hypothesis and conclude that the variance among the time since fire groups is NOT equal. 


# Make the same TSF plot but with Burn percentage #

```{r}
effect_both_doc_no3 <- rbind(no3_effect_size_burn_merged, doc_effect_size_burn_merged)

# Test for grouping 
df <- effect_both_doc_no3 %>%
  mutate(grouped_percentage = cut(burn_percentage, breaks = seq(0, 100, by = 5), include.lowest = TRUE, labels = FALSE))

df <- df %>% 
  mutate(grouped_percentage = as.character(grouped_percentage))

effect_both_doc_no3 <- df %>% 
  mutate(burn_percentage_group = case_when(grouped_percentage == "1" ~ "0-5",
                                           grouped_percentage == "2" ~ "6-10",
                                           grouped_percentage == "3" ~ "11-15",
                                           grouped_percentage == "4" ~ "16-20",
                                           grouped_percentage == "5" ~ "21-25",
                                           grouped_percentage == "6" ~ "26-30",
                                           grouped_percentage == "12" ~ "56-60",
                                           grouped_percentage == "14" ~ "66-70",
                                           grouped_percentage == "15" ~ "71-75",
                                           grouped_percentage == "16" ~ "76-80",
                                           grouped_percentage == "18" ~ "86-90",
                                           grouped_percentage == "20" ~ "96-100"))
                                           

#

effect_both_doc_no3$burn_percentage_group <- factor(effect_both_doc_no3$burn_percentage_group, levels = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "56-60", "66-70", "71-75", "76-80", "86-90", "96-100"))


# calculating the mean percent_difference by climate
burn_percentage_effect <- effect_both_doc_no3 %>% 
  group_by(response_var, burn_percentage_group) %>% 
  summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>% 
   mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect) 


# Plot #
vn = expression(paste(""*N*O[3]^"-"))

ggplot(effect_both_doc_no3, aes(burn_percentage_group, Effect_size, color = response_var),
       position = position_dodge(width = -0.5)) +
  geom_jitter(position = position_jitter(0.1), alpha = 0.4, size = 3) +
  geom_pointrange(aes(ymin = lower_ci_effect, ymax = upper_ci_effect,
                      color = response_var),
                  position = position_dodge(width = -0.5), size = 1.5, data = burn_percentage_effect) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  ylab("Effect Size") +
  xlab("Burn Percentage") +
  scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Analyte"),
                     labels = c('DOC', vn)) +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
                  position = position_nudge(x = -0.4)) +
  scale_x_discrete(expand = c(0.1, 0.1), limits = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "31-35", "36-40",
                                                   "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76-80",                                                      "81-85", "86-90",
                                                    "91-95", "96-100")) +
  ylim(-5, 5) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 30, angle = -90),
        axis.text.y = element_text(size = 30),
        axis.title.x = element_text(size = 40),
        axis.title.y = element_text(size = 35),
        legend.text = element_text(size = 30),
        legend.position = c(0.5, 0.86))

ggsave("Effect_Size_Burn_Percentage.pdf",
       path = ("~/GitHub/rc_sfa-rc-3-wenas-meta/FireMeta_Rproj/initial_plots/03_Wenas_effect_size"),
       width = 10, height = 8, units = "in")






# Break this down line by line #

burn_percentage_effect <- effect_both_doc_no3 %>% 
  group_by(response_var, burn_percentage_group) %>% 
  summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>% 
   mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect) 


ggplot(effect_both_doc_no3, aes(burn_percentage_group, Effect_size, color = response_var),
       position = position_dodge(width = -0.5)) +
  geom_jitter(position = position_jitter(0.1), alpha = 0, size = 3) +
  theme(legend.position = "none") +

  geom_pointrange(aes(ymin = lower_ci_effect, ymax = upper_ci_effect,
                      color = response_var),
                  position = position_dodge(width = -0.5), size = 1.5, data = burn_percentage_effect) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  ylab("Effect Size") +
  xlab("Burn Percentage") +
  scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = ""),
                     labels = c('DOC', vn)) +
  scale_x_discrete(expand = c(0.1, 0.1), limits = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "31-35", "36-40",
                                                   "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76-80",                                                      "81-85", "86-90",
                                                    "91-95", "96-100")) +
  ylim(-5, 5) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 20, angle = -90),
        axis.text.y = element_text(size = 30),
        axis.title.x = element_text(size = 40),
        axis.title.y = element_text(size = 35),
        legend.text = element_text(size = 30),
        legend.position = c(0.5, 0.86))

ggsave("AGU_Effect_Size_burn_percentage.pdf",
       path = here("initial_plots", "AGU"),
       width = 10, height = 8,  units = "in")

# Fade DOC
test <- burn_percentage_effect[c(10:21, 1:9), ]

test2 <- effect_both_doc_no3[c(79:123, 1:78), ]

ggplot(test2, aes(burn_percentage_group, Effect_size, alpha = response_var, color = response_var),
       position = position_dodge(width = -0.5)) +
  geom_jitter(position = position_jitter(0.1), alpha = 0, size = 3) +
  theme(legend.position = "none") +
  geom_pointrange(aes(ymin = lower_ci_effect, ymax = upper_ci_effect,
                      color = response_var
                      ),
                  position = position_dodge(width = -0.5), size = 1.5, data = test) +
  scale_alpha_manual(values = c(0.25, 1)) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  ylab("Effect Size") +
  xlab("Burn Percentage") +
  scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = ""),
                     labels = c('DOC', vn)) +
  scale_x_discrete(expand = c(0.1, 0.1), limits = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "31-35", "36-40",
                                                   "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76-80",                                                      "81-85", "86-90",
                                                    "91-95", "96-100")) +
  ylim(-5, 5) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 20, angle = -90),
        axis.text.y = element_text(size = 30),
        axis.title.x = element_text(size = 40),
        axis.title.y = element_text(size = 35),
        legend.text = element_text(size = 30),
        legend.position = "none")

ggsave("Fade_DOC_AGU_Effect_Size_burn_percentage.pdf",
       path = here("initial_plots", "AGU"),
       width = 10, height = 8,  units = "in")


# Fade NO3
ggplot(test2, aes(burn_percentage_group, Effect_size, alpha = response_var, color = response_var),
       position = position_dodge(width = -0.5)) +
  geom_jitter(position = position_jitter(0.1), alpha = 0, size = 3) +
  theme(legend.position = "none") +
  geom_pointrange(aes(ymin = lower_ci_effect, ymax = upper_ci_effect,
                      color = response_var
                      ),
                  position = position_dodge(width = -0.5), size = 1.5, data = test) +
  scale_alpha_manual(values = c(1, 0.25)) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  ylab("Effect Size") +
  xlab("Burn Percentage") +
  scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = ""),
                     labels = c('DOC', vn)) +
  scale_x_discrete(expand = c(0.1, 0.1), limits = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "31-35", "36-40",
                                                   "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76-80",                                                      "81-85", "86-90",
                                                    "91-95", "96-100")) +
  ylim(-5, 5) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 20, angle = -90),
        axis.text.y = element_text(size = 30),
        axis.title.x = element_text(size = 40),
        axis.title.y = element_text(size = 35),
        legend.text = element_text(size = 30),
        legend.position = "none")

ggsave("Fade_NO3_AGU_Effect_Size_burn_percentage.pdf",
       path = here("initial_plots", "AGU"),
       width = 10, height = 8,  units = "in")






# nitrate only
no3_effect_size_burn_merged <- effect_both_doc_no3 %>% 
  filter(response_var == "NO3_Interp")

no3_burn_percentage <- burn_percentage_effect %>% 
  filter(response_var == "NO3_Interp")

ggplot(no3_effect_size_burn_merged, aes(burn_percentage_group, Effect_size, color = response_var),
       position = position_dodge(width = -0.5)) +
  geom_jitter(position = position_jitter(0.1), alpha = 0, size = 3) +
  geom_pointrange(aes(ymin = lower_ci_effect, ymax = upper_ci_effect,
                      color = response_var),
                  position = position_dodge(width = -0.5), size = 1.5, alpha = 0.25, data = no3_burn_percentage) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  ylab("Effect Size") +
  xlab("Burn Percentage") +
  scale_color_manual(values = c("#E7B800"),
                     guide = guide_legend(title = ""),
                     labels = c(vn)) +
  scale_x_discrete(expand = c(0.1, 0.1), limits = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "31-35", "36-40",
                                                   "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76-80",                                                      "81-85", "86-90",
                                                    "91-95", "96-100")) +
  ylim(-5, 5) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 30, angle = -90),
        axis.text.y = element_text(size = 30),
        axis.title.x = element_text(size = 40),
        axis.title.y = element_text(size = 35),
        legend.text = element_text(size = 30),
        legend.position = c(0.5, 0.86))

ggsave("NO3_Effect_Size_Burn_Percentage.pdf",
       path = ("~/GitHub/rc_sfa-rc-3-wenas-meta/FireMeta_Rproj/initial_plots/03_Wenas_effect_size"),
       width = 10, height = 8, units = "in")

# DOC only
doc_effect_size_burn_merged <- effect_both_doc_no3 %>% 
  filter(response_var == "DOC_Interp")

doc_burn_percentage <- burn_percentage_effect %>% 
  filter(response_var == "DOC_Interp")

ggplot(doc_effect_size_burn_merged, aes(burn_percentage_group, Effect_size, color = response_var),
       position = position_dodge(width = -0.5)) +
  geom_jitter(position = position_jitter(0.1), alpha = 0, size = 3) +
  geom_pointrange(aes(ymin = lower_ci_effect, ymax = upper_ci_effect,
                      color = response_var),
                  position = position_dodge(width = -0.5), size = 1.5, data = doc_burn_percentage) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  ylab("Effect Size") +
  xlab("Burn Percentage") +
  scale_color_manual(values = c("#00AFBB"),
                     guide = guide_legend(title = ""),
                     labels = c("DOC")) +
  scale_x_discrete(expand = c(0.1, 0.1), limits = c("0-5", "6-10", "11-15", "16-20", "21-25", "26-30", "31-35", "36-40",
                                                   "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76-80",                                                      "81-85", "86-90",
                                                    "91-95", "96-100")) +
  ylim(-5, 5) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 30, angle = -90),
        axis.text.y = element_text(size = 30),
        axis.title.x = element_text(size = 40),
        axis.title.y = element_text(size = 35),
        legend.text = element_text(size = 30),
        legend.position = c(0.5, 0.86))

ggsave("DOC_Effect_Size_Burn_Percentage.pdf",
       path = ("~/GitHub/rc_sfa-rc-3-wenas-meta/FireMeta_Rproj/initial_plots/03_Wenas_effect_size"),
       width = 10, height = 8, units = "in")

  

```


