---
title: "Variance"
author: "JC & AMP"
date: "`r Sys.Date()`"
output: html_document
---
Compare effect size variances across all watersheds between normalized/non-normalized data

Chunk 2180 - 2439 is an attempt to do it myself. It looks like # HYPOTHESIS i SUPPORTED (overall): Area normalization significantly reduces effect-size variance. Watershed area explains a meaningful share of cross-watershed variability. Area normalization explains 62% [49%, 72%] of the raw effect-size variance. Anything over 50% tells us that area is the dominant source of cross-watershed variability. When I break it up by nitrate and DOC for looking at variance difference by response variable: both Variables are positive = area normalization captures watershed variance. When looking at variance RATIO by response variable (log scale: both variables >1 = normalization shrinks variance.

```{r Read in final data frame with non-normal and normalized}
rm(list=ls(all=T)) #this clears your Environment

library(pacman)
p_load(tidyverse,
       here,
       zoo,
       # hrbrthemes,
       viridis, 
       ggpmisc,
       RobustLinearReg,
       lme4,
       car)

# Read in effect size data frame for normalized and non-normalized data
ES_nn_normal <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_size_orig_nonnormal.csv"))

ES_nn_normal <- ES_nn_normal %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      Study_ID == "Abbott et al. 2021" & occurrence == 3 ~ "10", # Third value
      Study_ID == "Abbott et al. 2021" & occurrence == 4 ~ "11", # Fourth value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire)) %>% 
  dplyr::select(Study_ID:Effect_size_no_watershed)

  head(ES_nn_normal)
```
```{r levene prep}
# Make dataframe long format
ES_long <- ES_nn_normal %>%
  mutate(row_id = row_number()) %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed),
    names_to  = "type",
    values_to = "effect_size"
  ) %>%
  mutate(type = factor(type))

head(ES_long)

```
```{r prep var}
# Levene's Test of all data ####
cat("=== Standard Levene's Test ===\n")
levene_obs <- leveneTest(effect_size ~ type, data = ES_long, center = median)
print(levene_obs)
obs_F <- levene_obs$`F value`[1]

# Quick look at the observed variances
cat("\nVar(Effect_size NORMALIZED):", var(ES_nn_normal$Effect_size), "\n")
cat("Var(Effect_size NON-NORMALIZED):", var(ES_nn_normal$Effect_size_no_watershed), "\n")
cat("Variance ratio:", 
    var(ES_nn_normal$Effect_size) / var(ES_nn_normal$Effect_size_no_watershed), "\n\n")
```

```{r paired outlier check}
# Check for unusually large within-row shifts between non-normalized and normalized effect sizes.
paired_outlier_df <- ES_nn_normal %>%
  mutate(
    row_id = row_number(),
    paired_diff = Effect_size - Effect_size_no_watershed,
    abs_paired_diff = abs(paired_diff)
  )

paired_q1 <- quantile(paired_outlier_df$abs_paired_diff, 0.25, na.rm = TRUE)
paired_q3 <- quantile(paired_outlier_df$abs_paired_diff, 0.75, na.rm = TRUE)
paired_iqr <- IQR(paired_outlier_df$abs_paired_diff, na.rm = TRUE)
paired_upper_fence <- paired_q3 + 1.5 * paired_iqr
paired_median <- median(paired_outlier_df$paired_diff, na.rm = TRUE)
paired_mad <- mad(paired_outlier_df$paired_diff, constant = 1, na.rm = TRUE)

paired_outlier_df <- paired_outlier_df %>%
  mutate(
    iqr_outlier = abs_paired_diff > paired_upper_fence,
    robust_z = if (paired_mad == 0) {
      rep(NA_real_, n())
    } else {
      0.6745 * (paired_diff - paired_median) / paired_mad
    },
    robust_outlier = abs(robust_z) > 3.5
  )

cat("\n===== PAIRED DIFFERENCE OUTLIER CHECK =====\n")
cat(sprintf("IQR upper fence on absolute paired difference: %.3f\n", paired_upper_fence))
cat(sprintf("Median paired difference: %.3f | MAD: %.3f\n\n", paired_median, paired_mad))

cat("IQR outliers with metadata:\n")
print(
  paired_outlier_df %>%
    filter(iqr_outlier) %>%
    arrange(desc(abs_paired_diff)) %>%
    select(
      row_id, Study_ID, response_var, Climate, Biome,
      Effect_size_pair, Time_Since_Fire,
      Effect_size_no_watershed, Effect_size,
      paired_diff, abs_paired_diff, robust_z
    ),
  width = Inf
)

cat("\nRobust-z outliers with metadata:\n")
print(
  paired_outlier_df %>%
    filter(robust_outlier) %>%
    arrange(desc(abs_paired_diff)) %>%
    select(
      row_id, Study_ID, response_var, Climate, Biome,
      Effect_size_pair, Time_Since_Fire,
      Effect_size_no_watershed, Effect_size,
      paired_diff, abs_paired_diff, robust_z
    ),
  width = Inf
)

cat("\nStudy counts among IQR outliers:\n")
print(
  paired_outlier_df %>%
    filter(iqr_outlier) %>%
    count(Study_ID, sort = TRUE),
  width = Inf
)
```

```{r sensitivity analysis}
# Sensitivity analysis: compare the variance result after excluding studies
# that contribute the largest within-pair shifts.
sensitivity_summary <- function(df, label) {
  tibble(
    dataset = label,
    n = nrow(df),
    studies = n_distinct(df$Study_ID),
    var_non_normalized = var(df$Effect_size_no_watershed, na.rm = TRUE),
    var_normalized = var(df$Effect_size, na.rm = TRUE),
    diff = var_non_normalized - var_normalized,
    log_ratio = log(var_non_normalized / var_normalized)
  )
}

robust_outlier_studies <- paired_outlier_df %>%
  filter(robust_outlier) %>%
  distinct(Study_ID) %>%
  pull(Study_ID)

iqr_outlier_studies <- paired_outlier_df %>%
  filter(iqr_outlier) %>%
  distinct(Study_ID) %>%
  pull(Study_ID)

sensitivity_results <- bind_rows(
  sensitivity_summary(ES_nn_normal, "Full dataset"),
  sensitivity_summary(
    ES_nn_normal %>% filter(!Study_ID %in% robust_outlier_studies),
    "Exclude robust-outlier studies"
  ),
  sensitivity_summary(
    ES_nn_normal %>% filter(!Study_ID %in% iqr_outlier_studies),
    "Exclude IQR-outlier studies"
  )
) %>%
  mutate(
    interpretation = case_when(
      diff > 0 ~ "normalization reduces variance",
      diff < 0 ~ "normalization increases variance",
      TRUE ~ "no change"
    )
  )

cat("\n===== SENSITIVITY ANALYSIS =====\n")
cat("Robust-outlier studies removed:\n")
print(robust_outlier_studies)
cat("\nIQR-outlier studies removed:\n")
print(iqr_outlier_studies)

cat("\nVariance comparison across sensitivity datasets:\n")
print(sensitivity_results, width = Inf)
```

```{r dispersion comparison}
# Companion analysis: compare multiple spread metrics before and after normalization.
# This provides a robustness check because variance can be strongly influenced by outliers.
spread_compare <- function(x_non_normalized, x_normalized) {
  tibble(
    metric = c("Variance", "SD", "MAD", "IQR"),
    non_normalized = c(
      var(x_non_normalized, na.rm = TRUE),
      sd(x_non_normalized, na.rm = TRUE),
      mad(x_non_normalized, constant = 1, na.rm = TRUE),
      IQR(x_non_normalized, na.rm = TRUE)
    ),
    normalized = c(
      var(x_normalized, na.rm = TRUE),
      sd(x_normalized, na.rm = TRUE),
      mad(x_normalized, constant = 1, na.rm = TRUE),
      IQR(x_normalized, na.rm = TRUE)
    )
  ) %>%
    mutate(
      abs_change = normalized - non_normalized,
      fold_change = normalized / non_normalized,
      log_ratio = log(normalized / non_normalized),
      interpretation = case_when(
        abs_change < 0 ~ "normalization reduces spread",
        abs_change > 0 ~ "normalization increases spread",
        TRUE ~ "no change"
      )
    )
}

spread_overall <- spread_compare(
  ES_nn_normal$Effect_size_no_watershed,
  ES_nn_normal$Effect_size
)

spread_by_response <- ES_nn_normal %>%
  group_by(response_var) %>%
  group_modify(~ spread_compare(.x$Effect_size_no_watershed, .x$Effect_size)) %>%
  ungroup()

cat("\n===== DISPERSION COMPARISON (OVERALL) =====\n")
print(spread_overall, width = Inf)

cat("\n===== DISPERSION COMPARISON (BY RESPONSE VARIABLE) =====\n")
print(spread_by_response, width = Inf)
```

```{r bootstrap}

# BOOTSTRAPPING ####
set.seed(2024)
n_boot <- 10000

# Audit missingness before bootstrapping so any dropped rows are visible.
boot_na_audit <- ES_nn_normal %>%
  summarise(
    n_rows = n(),
    n_missing_effect_size = sum(is.na(Effect_size)),
    n_missing_effect_size_no_watershed = sum(is.na(Effect_size_no_watershed)),
    n_missing_either = sum(is.na(Effect_size) | is.na(Effect_size_no_watershed))
  )

cat("\n===== BOOTSTRAP NA AUDIT =====\n")
print(boot_na_audit)

# Bootstrap from the original wide data to avoid an unnecessary long -> wide round trip.
ES_boot <- ES_nn_normal %>%
  mutate(row_id = row_number()) %>%
  drop_na(Effect_size, Effect_size_no_watershed)

if (nrow(ES_boot) != nrow(ES_nn_normal) - boot_na_audit$n_missing_either) {
  stop("Unexpected row count after dropping incomplete effect-size pairs.")
}

if (anyNA(ES_boot$Effect_size) || anyNA(ES_boot$Effect_size_no_watershed)) {
  stop("NA values remain in bootstrap effect-size columns after filtering.")
}

head(ES_boot)

# paired bootstrap of variance difference and log variance ratio
# Resamples row_ids (whole pairs) to keep the pairing intact
boot_var_compare <- function(df, n_boot = 10000) {
  n <- nrow(df)
  replicate(n_boot, {
    bs  <- slice_sample(df, n = n, replace = TRUE)
    v_normalized <- var(bs$Effect_size,              na.rm = TRUE)
    v_non_normalized <- var(bs$Effect_size_no_watershed, na.rm = TRUE)
    c(var_normalized = v_normalized,
      var_non_normalized = v_non_normalized,
      diff = v_non_normalized - v_normalized,              # + -> area normalization reduces variance
      log_ratio = log(v_non_normalized / v_normalized),    # >0 -> area normalization reduces variance
      prop_explained = (v_non_normalized - v_normalized) / v_non_normalized)
  }) %>% t() %>% as_tibble()
}

# OVERALL comparison (all response variables pooled)
boot_overall <- boot_var_compare(ES_boot, n_boot)

head(boot_overall)

overall_summary <- boot_overall %>%
  pivot_longer(everything(), names_to = "metric") %>%
  group_by(metric) %>%
  summarise(
    mean    = mean(value),
    median  = median(value),
    ci_lo   = quantile(value, 0.025),
    ci_hi   = quantile(value, 0.975),
    .groups = "drop"
  )

overall_summary %>% filter(metric %in% c("diff", "log_ratio"))
```
```{r check}
# Interpretation:
# diff > 0 and log_ratio > 0 indicate that area normalization reduces
# cross-watershed variance, so watershed area is an important first-order
# source of variability. diff < 0 and log_ratio < 0 indicate that normalization
# increases variance, suggesting area alone is not sufficient and that
# additional watershed characteristics may be modulating response.
if (overall_summary %>%
      filter(metric == "diff") %>%
      pull(ci_lo) > 0) {
  cat("✔ HYPOTHESIS i SUPPORTED (overall):
      Area normalization reduces effect-size variance.
      Watershed area appears to be an important first-order source of cross-watershed variability.\n")
} else if (overall_summary %>%
             filter(metric == "diff") %>%
             pull(ci_hi) < 0) {
  cat("✔ HYPOTHESIS ii SUPPORTED (overall):
      Area normalization increases effect-size variance.
      Area alone does not appear to capture the dominant source of cross-watershed variability, and additional watershed characteristics may modulate fire response.\n")
} else {
  cat("⚠ INCONCLUSIVE (overall):
      The 95% CI of the variance difference spans zero, so area normalization does not clearly reduce or increase cross-watershed variance.\n")
}
```

```{r next}
# Summarize the fraction of non-normalized variance removed by area normalization.
prop_summary <- boot_overall %>%
  summarise(
    mean_prop   = mean(prop_explained),
    ci_lo       = quantile(prop_explained, 0.025),
    ci_hi       = quantile(prop_explained, 0.975)
  )

cat("\n===== PROPORTION OF VARIANCE REDUCED =====\n")
print(prop_summary)

cat(sprintf(
  "\nArea normalization reduces variance by %.0f%% [%.0f%%, %.0f%%] relative to the non-normalized effect sizes.\n",
  prop_summary$mean_prop * 100,
  prop_summary$ci_lo     * 100,
  prop_summary$ci_hi     * 100
))

# Heuristic interpretation of prop_explained:
# > 0.50 -> area is a dominant first-order source of cross-watershed variability
# 0.20 to 0.50 -> area is meaningful, but other factors likely matter too
# 0 to 0.20 -> area alone is a weak first-order correction
# < 0 -> normalization increases variance, consistent with hypothesis ii
```
```{r by response}
# One-sided summary: proportion of resamples where non-normalized variance
# exceeds normalized variance.
p_overall <- mean(boot_overall$diff > 0)

cat("\n===== OVERALL VARIANCE COMPARISON =====\n")
print(overall_summary)
cat(sprintf("\nP(var_non_normalized > var_normalized) = %.3f\n", p_overall))

# BY RESPONSE VARIABLE ####
boot_by_response <- ES_boot %>%
  group_by(response_var) %>%
  group_split() %>%
  map_dfr(function(df) {
    rv    <- unique(df$response_var)
    n_obs <- nrow(df)
    b     <- boot_var_compare(df, n_boot)

    tibble(
      response_var       = rv,
      n                  = n_obs,
      obs_var_normalized = var(df$Effect_size,              na.rm = TRUE),
      obs_var_non_normalized = var(df$Effect_size_no_watershed, na.rm = TRUE),
      obs_diff           = obs_var_non_normalized - obs_var_normalized,
      obs_log_ratio      = log(obs_var_non_normalized / obs_var_normalized),
      obs_prop_explained = (obs_var_non_normalized - obs_var_normalized) / obs_var_non_normalized,
      boot_diff_mean     = mean(b$diff),
      boot_diff_ci_lo    = quantile(b$diff, 0.025),
      boot_diff_ci_hi    = quantile(b$diff, 0.975),
      boot_log_ratio_mean  = mean(b$log_ratio),
      boot_log_ratio_ci_lo = quantile(b$log_ratio, 0.025),
      boot_log_ratio_ci_hi = quantile(b$log_ratio, 0.975),
      boot_prop_mean     = mean(b$prop_explained),
      boot_prop_ci_lo    = quantile(b$prop_explained, 0.025),
      boot_prop_ci_hi    = quantile(b$prop_explained, 0.975),
      p_non_normalized_gt_normalized = mean(b$diff > 0)
    )
  }) %>%
  mutate(
    direction = case_when(
      boot_diff_ci_lo > 0  ~ "normalization reduces variance",
      boot_diff_ci_hi < 0  ~ "normalization increases variance",
      TRUE                 ~ "inconclusive"
    )
  )

cat("\n===== BY RESPONSE VARIABLE =====\n")
print(boot_by_response, width = Inf)

boot_by_response %>%
  select(response_var, n, obs_var_non_normalized, obs_var_normalized,
         obs_diff, obs_prop_explained,
         boot_diff_ci_lo, boot_diff_ci_hi,
         boot_prop_ci_lo, boot_prop_ci_hi, direction) %>%
  arrange(desc(obs_prop_explained)) %>%
  print(width = Inf)

cat("\n===== BOOTSTRAP DIAGNOSTICS =====\n")

diagnostics_overall <- tibble(
  metric = c("diff", "log_ratio", "prop_explained"),
  observed = c(
    var(ES_boot$Effect_size_no_watershed, na.rm = TRUE) - var(ES_boot$Effect_size, na.rm = TRUE),
    log(var(ES_boot$Effect_size_no_watershed, na.rm = TRUE) / var(ES_boot$Effect_size, na.rm = TRUE)),
    (var(ES_boot$Effect_size_no_watershed, na.rm = TRUE) - var(ES_boot$Effect_size, na.rm = TRUE)) /
      var(ES_boot$Effect_size_no_watershed, na.rm = TRUE)
  ),
  boot_median = c(
    median(boot_overall$diff),
    median(boot_overall$log_ratio),
    median(boot_overall$prop_explained)
  ),
  boot_min = c(
    min(boot_overall$diff),
    min(boot_overall$log_ratio),
    min(boot_overall$prop_explained)
  ),
  boot_max = c(
    max(boot_overall$diff),
    max(boot_overall$log_ratio),
    max(boot_overall$prop_explained)
  ),
  ci_lo = c(
    quantile(boot_overall$diff, 0.025),
    quantile(boot_overall$log_ratio, 0.025),
    quantile(boot_overall$prop_explained, 0.025)
  ),
  ci_hi = c(
    quantile(boot_overall$diff, 0.975),
    quantile(boot_overall$log_ratio, 0.975),
    quantile(boot_overall$prop_explained, 0.975)
  )
)

print(diagnostics_overall, width = Inf)

cat("\nGroup sizes used in the response-specific bootstrap:\n")
print(boot_by_response %>% select(response_var, n), width = Inf)

cat("\nObserved variance inputs by response variable:\n")
print(
  boot_by_response %>%
    select(response_var, obs_var_non_normalized, obs_var_normalized,
           obs_diff, obs_log_ratio, obs_prop_explained),
  width = Inf
)

cat("\nPotential instability checks:\n")
cat(sprintf("Minimum bootstrapped normalized variance: %.4f\n", min(boot_overall$var_normalized)))
cat(sprintf("Minimum bootstrapped non-normalized variance: %.4f\n", min(boot_overall$var_non_normalized)))
cat(sprintf("Any non-finite log ratios? %s\n", any(!is.finite(boot_overall$log_ratio))))
cat(sprintf("Any non-finite prop_explained values? %s\n", any(!is.finite(boot_overall$prop_explained))))

# PLOTS
# Presentation-focused figures: one overall takeaway and two simple follow-ups.

plot_theme <- theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title.position = "plot",
    legend.position = "top"
  )

# 5a – By response variable: log variance ratio
plot_by_response_log_ratio <- boot_by_response %>%
  ggplot(aes(x = obs_log_ratio, y = reorder(response_var, obs_log_ratio))) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey55") +
  geom_errorbarh(aes(xmin = boot_log_ratio_ci_lo, xmax = boot_log_ratio_ci_hi),
                 height = 0.18, linewidth = 0.9, colour = "grey35") +
  geom_point(size = 3.2, colour = "#b5403b") +
  labs(
    title = "Log variance ratio by response variable",
    subtitle = "Positive values mean normalization reduces variance; negative values mean variance increases",
    x = "log(Var(non-normalized) / Var(normalized))",
    y = NULL
  ) +
  plot_theme

plot_by_response_log_ratio

ggsave(
  filename = "plot_by_response_log_ratio.png",
  plot = plot_by_response_log_ratio,
  width = 8.5,
  height = 4.5,
  dpi = 300
)

# 5b – Observed variances before and after normalization
plot_observed_variance_comparison <- boot_by_response %>%
  select(response_var, obs_var_non_normalized, obs_var_normalized) %>%
  pivot_longer(
    cols = c(obs_var_non_normalized, obs_var_normalized),
    names_to = "type",
    values_to = "variance"
  ) %>%
  mutate(type = case_when(
    type == "obs_var_non_normalized" ~ "Non-normalized",
    type == "obs_var_normalized" ~ "Area-normalized",
    TRUE ~ type
  )) %>%
  ggplot(aes(x = variance, y = reorder(response_var, variance), colour = type)) +
  geom_line(aes(group = response_var), linewidth = 1, colour = "grey75") +
  geom_point(size = 3) +
  labs(
    title = "Observed variances before and after normalization",
    subtitle = "Grey connectors show the direction of change within each response variable",
    x = "Observed variance",
    y = NULL,
    colour = NULL
  ) +
  scale_colour_manual(values = c("Non-normalized" = "#b5403b", "Area-normalized" = "#1f4e79")) +
  plot_theme

plot_observed_variance_comparison

ggsave(
  filename = "plot_observed_variance_comparison.png",
  plot = plot_observed_variance_comparison,
  width = 8.5,
  height = 4.5,
  dpi = 300
)

# 5c – Paired effect sizes for each observation
plot_paired_effect_sizes <- ES_boot %>%
  mutate(obs_label = paste0(row_id, " | ", Study_ID)) %>%
  select(row_id, obs_label, response_var, Effect_size_no_watershed, Effect_size) %>%
  pivot_longer(
    cols = c(Effect_size_no_watershed, Effect_size),
    names_to = "type",
    values_to = "effect_size"
  ) %>%
  mutate(
    type = case_when(
      type == "Effect_size_no_watershed" ~ "Non-normalized",
      type == "Effect_size" ~ "Area-normalized",
      TRUE ~ type
    )
  ) %>%
  ggplot(aes(x = effect_size, y = factor(obs_label, levels = rev(unique(obs_label))), colour = type)) +
  geom_line(aes(group = row_id), linewidth = 0.35, colour = "grey80", alpha = 0.8) +
  geom_point(size = 1.6, alpha = 0.9) +
  facet_wrap(~ response_var, scales = "free_y") +
  labs(
    x = "Effect size",
    y = "Observation | Study_ID",
    colour = NULL
  ) +
  scale_colour_manual(values = c("Non-normalized" = "#b5403b", "Area-normalized" = "#1f4e79")) +
  plot_theme +
  theme(
    panel.grid.major.y = element_blank(),
    strip.text = element_text(face = "bold"),
    axis.text.y = element_text(size = 6)
  )

plot_paired_effect_sizes

ggsave(
  filename = "plot_paired_effect_sizes.png",
  plot = plot_paired_effect_sizes,
  width = 9,
  height = 8,
  dpi = 300
)
```
