---
title: "Map"
output: html_document
date: "2023-10-19"
editor_options: 
  chunk_output_type: console
---
The purpose of this script is create the map figure 

Script Workflow:

Step 1) Load in the summary metadata sheet that include lat and long

Step 2) Load in the shape files of the burns

Step 3) Plot


# Status: in progress

# ==============================================================================
# Author: Jake Cavaiani 
# 19 October 2023
# ==============================================================================

## Load packages and set working directory
```{r Jake/Mac}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment


library(tidyverse)
library(here)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
library(broom)
library(usmap)
library(maps)
library(mapdata)
library(cowplot)
library(sf)
library(RColorBrewer)
```

This is the workflow to read in the meta analysis studies and the fires they are studying and then loading in a CONUS map and plotting our data on top of that
```{r load in fire names and associated study}

meta <- read_csv(here("inputs", "Studies_Summary", "Studies_included_in_climate_figure.csv")) # reading in filtered meta data sheet

coords <- meta %>% 
  select(Fire_Name, Lat, Long, number_of_burn, State, group) # Picking only the columns I want 

coords_noNA <- coords %>% 
    drop_na(Fire_Name)
 
# Create an sf object from the data
coordinates_sf <- st_as_sf(coords_noNA, coords = c("Long", "Lat"), crs = 4326)

# Download the United States map shapefile from the US Census Bureau
# Replace "cb_2019_us_state_20m.zip" with the most recent shapefile available
url <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url, "us_states.zip", mode = "wb")
unzip("us_states.zip", exdir = "us_states")

# Read the shapefile
us_states <- st_read("us_states/cb_2021_us_state_20m.shp")

# Plot the map with points
ggplot() +
  geom_sf(data = us_states, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(color = Fire_Name)) +
  xlim(-125, -68) +
  scale_size_continuous(range = c(1, 10)) +
  labs(title = "Points on US Map with Conditional Variable") +
  theme_minimal()

# Color values #
"#666666"
"#00AFBB"
"#1B9E77"
"#7570B3"
"#E7298A"
"darkred"
"#A6CEE3"
"#1F78B4"
"#E31A1C"
"#33A02C"
"#FB9A99"
"black"
ggplot() +
  geom_sf(data = us_states, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_Name)) +
  scale_size_continuous(guide = guide_legend(title = "")) +
  scale_color_manual(values = c("#666666", "#A6CEE3", "#00AFBB", "#1B9E77",
                                "#7570B3", "#E7298A", "darkred", "black", 
                                "#1F78B4", "#E31A1C", "#33A02C", "#FB9A99",
                                "darkblue", "darkgreen"), 
                     guide = guide_legend(title = "Fire Name")) +
  xlim(-125, -68) +
  ylim(25, 52) +
  theme_map() +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(),
        )

ggsave("climate_figure_map.pdf",
       path = here("initial_plots", "Maps"),
       width = 6, height = 8,  units = "in")


```


### Xia map making ###
```{r workflow for a map Xia}
effect_size_fire <- read_csv(here("Output_for_analysis", "Effect_Size_Geospatial_Fire_Lat_Long.csv")) # effect size data frame with lat longs and geospatial data and fire names

# IF there are no geospatial data (comid, totdasqkm, streamorde, burn_percentage) its because the watersheds couldn't be pulled from the NHD+ database. This might be because:
    #1) Might be out of the country
    #2) fires are too old (pre-1984) or too recent (2021-present)
# There is a subset of papers that were used for the Climate analysis and a subset of papers that were used for the burn percentage analysis due to these factors. So unfortunately this is what we got to work with. 

# Create an sf object from the data
map_sf <- st_as_sf(effect_size_fire, coords = c("Long", "Lat"), crs = 4326)

# Download the United States map shapefile from the US Census Bureau
# Replace "cb_2019_us_state_20m.zip" with the most recent shapefile available
url <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url, "us_states.zip", mode = "wb")
unzip("us_states.zip", exdir = "us_states")

# Read the shapefile
us_states <- st_read("us_states/cb_2021_us_state_20m.shp")

# Plot the map with points
ggplot() +
  geom_sf(data = us_states, fill = "white", color = "black") +
  geom_sf(data = map_sf, aes(color = Fire_name)) +
  xlim(-125, -68) +
  scale_size_continuous(range = c(1, 10)) +
  labs(title = "Points on US Map with Conditional Variable") +
  theme_minimal()


```






### Dataframe for Xia to make a map ### - completed on 1/10/2024
```{r - merge effect size dataframe with the fires dataframe}
effect_size <- read_csv(here("Output_for_analysis", "effect_size_geospatial_fire.csv"))

meta <- read_csv(here("inputs", "Studies_Summary", "Studies_included_in_climate_figure.csv")) # reading in filtered meta data sheet

meta <- meta %>%
  rename(Study_ID = Study) # rename Study to Study_ID to match the effect size dataframe


coords <- meta %>%
  select(Study_ID, Fire_Name, Lat, Long, number_of_burn, State, group) %>%  # Picking only the columns I want
  rename(Fire_name = Fire_Name) # rename Fire_Name in the meta df to Fire_name to match the effect size df to merge

effect_size_fire <- full_join(effect_size, coords) # this merges the df with lat/longs

effect_size_fire <- effect_size_fire %>% 
   drop_na(Effect_size) # This is dropping any row that doesn't have an effect size calculated for a specific site which means these are the sites that have a pre/post study design that we arent analyzing for the final product 
  

# Anaktuvuk River wildfire coordinates: 68.932756,	-150.7
effect_size_fire <- effect_size_fire %>% 
  mutate(Lat = case_when(Study_ID == 'Abbott et al. 2021' ~ 68.932756,
                                     TRUE ~ Lat),
         Long = case_when(Study_ID == 'Abbott et al. 2021' ~ -150.7,
                                     TRUE ~ Long))

# Northwest Territories Fire coordinates: 61.4	-121.433
effect_size_fire <- effect_size_fire %>% 
  mutate(Lat = case_when(Study_ID == 'Burd et al 2018' ~ 61.4,
                                     TRUE ~ Lat),
         Long = case_when(Study_ID == 'Burd et al 2018' ~ -121.433,
                                     TRUE ~ Long))
# Caldor Fire: 38.88376	-119.978005
effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_name = case_when(Study_ID == 'Hickenbottom et al. 2023' &
                               Time_Since_Fire == 1 ~ 'Caldor Fire',
                                        TRUE ~ Fire_name))

effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_year = case_when(Study_ID == 'Hickenbottom et al. 2023' &
                               Fire_name == 'Caldor Fire' ~ '2021',
                                        TRUE ~ Fire_year))

effect_size_fire <- effect_size_fire %>% 
  mutate(Lat = case_when(Fire_name == 'Caldor Fire' ~ 38.88376,
                                     TRUE ~ Lat),
         Long = case_when(Fire_name == 'Caldor Fire' ~ -119.978005,
                                     TRUE ~ Long))
                          
# Mosquito Fire: 38.958701	-120.920501
effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_name = case_when(Study_ID == 'Hickenbottom et al. 2023' &
                               Time_Since_Fire == 0 ~ 'Mosquito Fire',
                                        TRUE ~ Fire_name))

effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_year = case_when(Study_ID == 'Hickenbottom et al. 2023' &
                               Fire_name == 'Mosquito Fire' ~ '2022',
                                        TRUE ~ Fire_year))

effect_size_fire <- effect_size_fire %>% 
  mutate(Lat = case_when(Fire_name == 'Mosquito Fire' ~ 38.958701,
                                     TRUE ~ Lat),
         Long = case_when(Fire_name == 'Mosquito Fire' ~ -120.920501,
                                     TRUE ~ Long))

# Rampage Fire Coordinates 48.418961	-113.696178: 
effect_size_fire <- effect_size_fire %>% 
  mutate(Lat = case_when(Study_ID == 'Mast & Clow, 2008' ~ 48.418961,
                                     TRUE ~ Lat),
         Long = case_when(Study_ID == 'Mast & Clow, 2008' ~ -113.696178,
                                     TRUE ~ Long))

effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_name = case_when(Study_ID == 'Mast & Clow, 2008' ~ 'Rampage Fire',
                                     TRUE ~ Fire_name))

# Fourmile canyon coordinates: 40.03333 -105.4167
effect_size_fire <- effect_size_fire %>% 
  mutate(Lat = case_when(Fire_name == 'Fourmile Canyon Fire' ~ 40.03333,
                                     TRUE ~ Lat),
         Long = case_when(Fire_name == 'Fourmile Canyon Fire' ~ -105.4167,
                                     TRUE ~ Long))

# Hayman Fire coordinates: 
effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_name = case_when(Study_ID == 'Rhea et al. 2021' ~ 'Hayman Fire',
                                     TRUE ~ Fire_name),
         Fire_year = case_when(Study_ID == 'Rhea et al. 2021' ~ '2002',
                                     TRUE ~ Fire_year),
         Lat = case_when(Fire_name == 'Hayman Fire' ~ 39.2,
                                     TRUE ~ Lat),
         Long = case_when(Fire_name == 'Hayman Fire' ~ -105.3,
                                     TRUE ~ Long))

# Tiedemann, 1997 Safety Harbor Fire 48.10267, -120.3523
effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_name = case_when(Study_ID == 'Tiedemann, 1997' ~ 'Safety Harbor Fire',
                                     TRUE ~ Fire_name),
         Fire_year = case_when(Study_ID == 'Tiedemann, 1997' ~ '1980',
                                     TRUE ~ Fire_year),
         Lat = case_when(Fire_name == 'Safety Harbor Fire' ~ 48.10267,
                                     TRUE ~ Lat),
         Long = case_when(Fire_name == 'Safety Harbor Fire' ~ -120.3523,
                                     TRUE ~ Long))

# Uzun Rocky Fire coordinates: 38.923564	-122.326478
effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_name = case_when(Study_ID == 'Uzun et al. 2020' &
                               site == 'Rocky Fire' ~ 'Rocky Fire',
                                     TRUE ~ Fire_name),
         Lat = case_when(Fire_name == 'Rocky Fire' ~ 38.923564,
                                     TRUE ~ Lat),
         Long = case_when(Fire_name == 'Rocky Fire' ~ -122.326478,
                                     TRUE ~ Long))
# Uzun Wragg Fire coordinates: 38.512031	-122.097228
effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_name = case_when(Study_ID == 'Uzun et al. 2020' &
                               site == 'Wragg Fire' ~ 'Wragg Fire',
                                     TRUE ~ Fire_name),
         Lat = case_when(Fire_name == 'Wragg Fire' ~ 38.512031,
                                     TRUE ~ Lat),
         Long = case_when(Fire_name == 'Wragg Fire' ~ -122.097228,
                                     TRUE ~ Long))

# Writer et al. 2014 # High Park Fire 40.71600 -105.2330
effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_name = case_when(Study_ID == 'Writer et al. 2014' ~ 'High Park Fire',
                                     TRUE ~ Fire_name),
         Fire_year = case_when(Study_ID == 'Writer et al. 2014' ~ '2012',
                                     TRUE ~ Fire_year),
         Lat = case_when(Fire_name == 'High Park Fire' ~ 40.71600,
                                     TRUE ~ Lat),
         Long = case_when(Fire_name == 'High Park Fire' ~ -105.2330,
                                     TRUE ~ Long))

# Neary fire year - 1978 
effect_size_fire <- effect_size_fire %>% 
  mutate(Fire_year = case_when(Study_ID == 'Neary & Currier, 1982' ~ '1978',
                                     TRUE ~ Fire_year))

write_csv(effect_size_fire, here("Output_for_analysis", "Effect_Size_Geospatial_Fire_Lat_Long.csv"))
```



```{r graveyard code }
# Plot the high park fire
# shapefile <- st_read("co4058910540420120609_20110912_20130901_burn_bndy.shp")
# 
# 
# ggplot() +
#   geom_sf(data = shapefile, aes(fill = "red")) +
#   theme_bw()
# 
# ggplot() +
#   geom_sf(data = shapefile, aes(fill = "red")) +
#   scale_fill_viridis_c(option = "A") +
#   labs(title = "Your Shapefile Plot", fill = "Legend Title")
# 
# # Plot multiple fires
# angora <- st_read(here("gis_data", "Fire_Perimeters", "Angora", "ca3888612004020070623_20060812_20080630_burn_bndy.shp"))
# 
# caldor <- st_read(here("gis_data", "Fire_Perimeters", "Caldor", "ca3858612053820210815_20210805_20220723_burn_bndy.shp"))
# 
# camp_branch <- st_read(here("gis_data", "Fire_Perimeters", "Camp_Branch", "nc3516508354120161123_20151219_20161221_burn_bndy.shp"))
# 
# clovermist <- st_read(here("gis_data", "Fire_Perimeters", "Clovermist", "wy4473710995419880709_19860818_19890802_burn_bndy.shp"))
# 
# fern_lake <- st_read(here("gis_data", "Fire_Perimeters", "Fern_Lake", "co4036110556320121009_20100925_20130917_burn_bndy.shp"))
# 
# four_mile <- st_read(here("gis_data", "Fire_Perimeters", "Fourmile_Canyon", "co4005110538520100906_20100621_20110624_burn_bndy.shp"))
# 
# gaviota <- st_read(here("gis_data", "Fire_Perimeters", "Gaviota", "ca3448712019620040605_20040425_20050412_burn_bndy.shp"))
#   
# high_park <- st_read(here("gis_data", "Fire_Perimeters", "High_Park", "co4058910540420120609_20110912_20130901_burn_bndy.shp"))
# 
# hayman <- st_read(here("gis_data", "Fire_Perimeters", "Hayman", "co3922010528720020608_20010824_20030814_burn_bndy.shp"))
# 
# rampage <- st_read(here("gis_data", "Fire_Perimeters", "Rampage_Complex", "mt4839411359020030819_20010707_20040715_burn_bndy.shp"))
# 
# red_bench <- st_read(here("gis_data", "Fire_Perimeters", "Red_Bench", "mt4878711426219880906_19880819_19890729_burn_bndy.shp"))
# 
# rocky <- st_read(here("gis_data", "Fire_Perimeters", "Rocky", "ca3891212249320150729_20140910_20150904_burn_bndy.shp"))
# 
# tellico <- st_read(here("gis_data", "Fire_Perimeters", "Tellico", "nc3527408356020161103_20151219_20161221_burn_bndy.shp"))
# 
# thompson_ridge <- st_read(here("gis_data", "Fire_Perimeters", "Thompson_Ridge", "nm3589210662020130531_20110608_20140616_burn_bndy.shp"))
# 
# wragg <- st_read(here("gis_data", "Fire_Perimeters", "Wragg", "ca3851412213120150722_20150711_20150812_burn_bndy.shp"))
# 
# # 15 fires 
# 
# 
# 
# combined_shapefile <- rbind(angora, caldor, camp_branch, clovermist,
#                             fern_lake, four_mile, gaviota, high_park, 
#                             hayman, rampage, red_bench, rocky,
#                             tellico, thompson_ridge, wragg)
# 
# ggplot() +
#   geom_sf(data = combined_shapefile, aes(fill = "red"))
# 
# ggplot() +
#   geom_sf(data = wragg, aes(fill = "red"))
# 
# 
# ggplot(data = world) +
#     geom_sf() +
#     geom_point(data = coords, aes(x = Long, y = Lat), size = 4, 
#         shape = 23, fill = "darkred") +
#     geom_sf(data = combined_shapefile, aes(fill = "red")) +
#     coord_sf(xlim = c(-125, -70), ylim = c(24, 55), expand = FALSE) +
#   theme_bw()
# 
# ggplot() +
#   geom_sf(data = combined_shapefile, aes(fill = "red")) +
#     coord_sf(xlim = c(-125, -70), ylim = c(24, 55), expand = FALSE)
#   



```











