---
title: "02a_geospatial_comid_extraction"
output: html_document
date: "2026-04-30"
---


# ==============================================================================
# Author: Kathryn Willi & Matthew R. V. Ross: https://doi.org/10.5281/zenodo.8140272

# Adapted by Jake Cavaiani for the manuscript: Catchment characteristics modulate the influence of wildfires on nitrate and dissolved organic carbon across space and time: A meta-analysis; jake.cavaiani@pnnl.gov

# 30 October 2023
# ==============================================================================


```{r setup, include=TRUE, echo = T, warning = F, comment = F, message = FALSE}
rm(list = ls())
p_load(sf, 
       tidyverse, 
       terra, 
       nhdplusTools, 
       mapview, 
       dataRetrieval,
       lubridate, 
       prism, 
       ggspatial, 
       nngeo, 
       stars, 
       here,
       webshot, 
       readxl, 
       StreamCatTools)

# this gives you an error, but it can be ignored:
try(plyr::ldply(list.files(path = here("R_scripts", "data", "src"),
                           pattern="*.R",
                           full.names=TRUE),
                source))
# Rmarkdown options
knitr::opts_chunk$set(echo = T, warning = F, comment = F, message = F)

# mapview options
mapviewOptions(basemaps.color.shuffle=FALSE,basemaps='OpenTopoMap')
```

### Setting up your site data set.

For this code to run properly, your site data must be configured as follows:

1)  Each site is identified with a unique site name. In the data set, this column must be called `site`.
2)  Each site has coordinates, with column names `longitude` and `latitude`. Knowledge of coordinate projection required. **OR:**
    Each site has their known COMID, with column name `comid`. 
4)  Site data table is a CSV, and stored in the `data/` folder. 

I have included an example data set called `placeholder.csv`, along with all of the additional data sets necessary for the code to run, stored in the `data` folder. 

#### Downloading necessary data sets

Currently, this workflow requires downloading several data sets locally for much speedier run times. This includes: PRISM climate & aridity rasters, NHD flow direction data, and CONUS-wide NHD catchments. All data sets are found in the shared `data` folder.

### Site data assumptions.

This analysis is only appropriate for locations along adequately-sized streams. Some streams are too small to be captured by NHDPlusV2; it is also common for coordinates to fall in the wrong catchment (especially for big rivers). For that reason, review each site and make sure that the NHD feature attributes were appropriately captured.

# National Hydrodraphy Dataset (NHD) data extraction

Identify each sample's NHD COMID. This COMID will allow site linkages with all datasets in this workflow. If COMID is already listed in the CSV, make `site_type = "comid"`.

```{r}
sf_use_s2(FALSE)

site_type = "xy" # OR site_type = "comid"

sites <- read_csv(here("Output_for_analysis", "01_read_and_gapfill", "01_daily_time_series_paired.csv"), na = c('-9999', 'N/A')) %>%
  distinct(Site, .keep_all = TRUE) %>%
  dplyr::select('Site','latitude','longitude') %>% 
  rename(site = Site) %>% 
  drop_na()


```

Additional steps for sites with coordinates, and no COMID:

```{r}
if(site_type == "xy"){
  sites <- sites %>%
    dplyr::select(site, latitude, longitude) %>% 
    sf::st_as_sf(coords = c("longitude","latitude"), crs = 4269) # 4269 = NAD83 CRS
  
  if(sf::st_crs(sites) != sf::st_crs(4269)){
    sites <- sites %>% st_transform(., crs = 4269)
  }
  
  mapview(sites)
}
```

Pull all meta data associated with each site's COMID. 

```{r}

sites$comid <- NA
for(i in 1:nrow(sites)){
  result <- try(discover_nhdplus_id(sites[i,]), silent = TRUE)
  if(!inherits(result, "try-error") && length(result) > 0){
    sites$comid[i] <- result
  } else {
    message("No COMID for site: ", sites$site[i])
  }
}

sites <- sites |> 
  mutate(comid = case_when(site == "Gaviota" ~ 17594763, 
                           TRUE ~ comid))

# NEW: extract coordinates into plain columns and drop geometry
sites_1 <- sites %>%
  mutate(
    longitude = sf::st_coordinates(.)[,1],
    latitude  = sf::st_coordinates(.)[,2]
  ) %>%
  sf::st_drop_geometry()


```

### Setting up your site data set.

For this code to run properly, your site data must be configured as follows:

1)  Each site is identified with a unique site name. In the data set, this column must be called `site`.
2)  Each site has their known COMID, with column name `comid`. 
4)  Site data table is a CSV, and stored in the `data/` folder. 


#### Downloading necessary data sets

Currently, this workflow requires downloading several data sets locally for much speedier run times. This includes: PRISM climate & aridity rasters, NHD flow direction data, and CONUS-wide NHD catchments. All data sets are found in the shared `data` folder.


# National Hydrodraphy Dataset (NHD) data extraction

Use COMID to allow site linkages with all datasets in this workflow. 
Pull all meta data associated with each site's COMID. 
```{r}
# site_type = "comid" # This steps assumes you have checked your COMIDs
# 
# if(site_type == "comid"){
#   sites <- getNHDcomid(df = dplyr::select(sites, site, comid))
# }

site_type = "comid"

if(site_type == "comid"){
  
  # 1. Save the full site list (with lat/long preserved)
  sites_all <- sites_1
  
  # 2. Identify sites without COMIDs
  sites_no_comid <- sites_all %>% dplyr::filter(is.na(comid))
  
  if(nrow(sites_no_comid) > 0){
    message("Skipping NHD extraction for ", nrow(sites_no_comid),
            " site(s) without COMIDs: ",
            paste(sites_no_comid$site, collapse = ", "))
  }
  
  # 3. Run getNHDcomid() only on sites WITH valid COMIDs
  sites_with_comid <- sites_all %>%
    dplyr::filter(!is.na(comid)) %>%
    dplyr::select(site, comid)
  
  sites_nhd <- getNHDcomid(df = sites_with_comid)
  
  # 4. Drop sf geometry so we can join as a regular data frame
  sites_nhd_df <- sites_nhd %>% 
    sf::st_drop_geometry() %>%
    dplyr::select(-any_of(c("latitude", "longitude")))   # drop NHD-snapped coords
  
  # 5. Use sites_all as the spine (keeps original lat/long for ALL sites),
  #    then left_join the NHD metadata onto it
  sites <- sites_all %>%
    dplyr::left_join(sites_nhd_df, by = c("site", "comid"))
  
  # 6. Sanity checks
  message("Total rows: ", nrow(sites),
          " | With NHD data: ", sum(!is.na(sites$comid)),
          " | Without (NA): ", sum(is.na(sites$comid)))
}

# Remove data frames that are not needed
rm(sites_all)
rm(sites_nhd)
rm(sites_nhd_df)
rm(sites_no_comid)
rm(sites_with_comid)

```


# STREAMCAT BURN PERCENT for each COMID 
```{r}
# ---- 1. Read fire-year metadata from Sites_meta_data.csv ---------
sites_meta <- read_csv(here("inputs", "Studies_Summary", "Sites_meta_data.csv"),
                       show_col_types = FALSE) %>%
  dplyr::select(Study_ID, Fire_year) %>%
  distinct()

# Expand multi-year fire strings (e.g., "2021_2022") into separate rows
fire_years_long <- sites_meta %>%
  mutate(Fire_year = str_split(as.character(Fire_year), "_")) %>%
  unnest(Fire_year) %>%
  mutate(Fire_year = as.integer(Fire_year)) %>%
  filter(!is.na(Fire_year))

# ---- 2. Link each Site to its Study_ID and fire year(s) ----------
daily_ts <- read_csv(here("Output_for_analysis", "01_read_and_gapfill",
                          "01_daily_time_series_paired.csv"),
                     show_col_types = FALSE)

site_fire_key <- daily_ts %>%
  distinct(Study_ID, Site) %>%
  left_join(fire_years_long, by = "Study_ID") %>%
  left_join(sites_1 %>% dplyr::select(site, comid) %>%
              rename(Site = site),
            by = "Site") %>%
  filter(!is.na(comid), !is.na(Fire_year))

# ---- 3. Build the metric list from the unique fire years ---------
# StreamCat MTBS coverage: 1984–2018
needed_years  <- unique(site_fire_key$Fire_year)
covered_years <- needed_years[needed_years >= 1984 & needed_years <= 2018]
uncovered <- setdiff(needed_years, covered_years)

if(length(uncovered) > 0){
  message("Fire years outside StreamCat MTBS coverage (1984–2018): ",
          paste(uncovered, collapse = ", "),
          " — these sites will have NA burn percent.")
}

# Metric naming: MTBS_{year} (the API will return the Ws version because
# we pass aoi = "watershed"). The returned column will be named like
# "mtbs_{year}ws" after lowercasing.
mtbs_metrics <- paste0("pctburnarea", covered_years)
message("Requesting StreamCat metrics: ", paste(mtbs_metrics, collapse = ", "))

# ---- 4. Pull burn percentages from StreamCat API -----------------
unique_comids <- unique(site_fire_key$comid)

burn_data <- sc_get_data(
  metric = paste(mtbs_metrics, collapse = ","),
  aoi    = "watershed",
  comid  = paste(unique_comids, collapse = ",")
)

# Standardize column names to lowercase (API returns mixed case)
names(burn_data) <- tolower(names(burn_data))

# Quick sanity check
message("burn_data columns: ", paste(names(burn_data), collapse = ", "))

# ---- 5. Pivot to long: one row per (comid, year) -----------------
burn_long <- burn_data %>%
  pivot_longer(cols = starts_with("pctburnarea"),
               names_to  = "metric",
               values_to = "burn_percent") %>%
  mutate(Fire_year = as.integer(str_extract(metric, "\\d{4}"))) %>%
  dplyr::select(comid, Fire_year, burn_percent)

# ---- 6. Match each site to its fire-year burn percentage ---------
site_burn <- site_fire_key %>%
  left_join(burn_long, by = c("comid", "Fire_year"))

# For multi-year fires (e.g., 2021_2022), collapse per site
site_burn_summary <- site_burn %>%
  group_by(Study_ID, Site, comid) %>%
  summarise(
    burn_percent_fire_year = sum(burn_percent, na.rm = TRUE),   # or max()
    fire_years_used        = paste(sort(unique(Fire_year)), collapse = "_"),
    .groups = "drop"
  )

# ---- 7. Join onto sites (rename Site -> site to match sites' lowercase key) -----------
sites_bp_geo <- sites %>%
  left_join(
    site_burn_summary %>% 
      dplyr::select(-Study_ID) %>%
      rename(site = Site),
    by = c("site", "comid")
  )


```

# MTBS BURN SEVERITY percent low, moderate, high for each COMID 
```{r}
# ---- 1. Build severity metric list from the covered fire years ---
# We already have `site_fire_key` and `covered_years` from the burn_percent pull
sev_metrics <- c(
  paste0("pcthighsev", covered_years),
  paste0("pctmodsev",  covered_years),
  paste0("pctlowsev",  covered_years)
)
message("Requesting severity metrics: ", paste(sev_metrics, collapse = ", "))

# ---- 2. Pull from StreamCat API ---------------------------------
sev_data <- sc_get_data(
  metric = paste(sev_metrics, collapse = ","),
  aoi    = "watershed",
  comid  = paste(unique_comids, collapse = ",")
)

# Standardize column names
names(sev_data) <- tolower(names(sev_data))
message("sev_data columns: ", paste(names(sev_data), collapse = ", "))

# ---- 3. Pivot to long: one row per (comid, year, severity class) -
sev_long <- sev_data %>%
  pivot_longer(
    cols = matches("^pct(high|mod|low)sev\\d{4}ws$"),
    names_to = "metric",
    values_to = "severity_percent"
  ) %>%
  mutate(
    Fire_year = as.integer(str_extract(metric, "\\d{4}")),
    severity_class = str_extract(metric, "(high|mod|low)")
  ) %>%
  dplyr::select(comid, Fire_year, severity_class, severity_percent)

# ---- 4. Match each site to its fire-year severity values ---------
site_sev <- site_fire_key %>%
  left_join(sev_long, by = c("comid", "Fire_year"))

# For multi-year fires, collapse by summing (or max)
site_sev_summary <- site_sev %>%
  group_by(Study_ID, Site, comid, severity_class) %>%
  summarise(
    severity_percent = sum(severity_percent, na.rm = TRUE),   # or max()
    fire_years_used  = paste(sort(unique(Fire_year)), collapse = "_"),
    .groups = "drop"
  )

# ---- 5. Pivot back to wide: one column per severity class --------
site_sev_wide <- site_sev_summary %>%
  dplyr::select(-fire_years_used) %>%    # already captured in burn_percent join
  pivot_wider(
    names_from = severity_class,
    values_from = severity_percent,
    names_prefix = "burn_sev_"           # -> burn_sev_high, burn_sev_mod, burn_sev_low
  )

# ---- 6. Join onto the sites data frame ---------------------------
sites_sev_bp_geo <- sites_bp_geo %>%
  left_join(
    site_sev_wide %>%
      dplyr::select(-Study_ID) %>%
      rename(site = Site),
    by = c("site", "comid")
  )

# ---- 7. Based on comid_fact_check.Rmd we need to manually change some bp and severity values -------
# Define site groups once to keep it readable
unburn_sites <- c("Logging_Creek", "Wally_Creek", "Pinchot", "Site_1",
                  "Crow Creek", "U3",
                  "Reference First Rainy Season",
                  "Reference Second Rainy Season",
                  "Reference Subsequent Flush")

na_sites_all <- c("East Fork Kootenai", "Middle Fork Kootenai",
                  "Middle_Fork_American", "Trout_Creek",
                  "Wash Branch", "Camas Creek")

na_sites_sev_only <- c("East Fork Kootenai", "Middle Fork Kootenai")

sites_sev_bp_geo <- sites_sev_bp_geo %>%
  mutate(
    # --- burn_percent_fire_year ---
    burn_percent_fire_year = case_when(
      site %in% unburn_sites  ~ 0,
      site %in% na_sites_all  ~ NA_real_,
      TRUE                    ~ burn_percent_fire_year
    ),
    # --- all burn_sev_* columns ---
    across(starts_with("burn_sev"),
           ~ case_when(
               site %in% unburn_sites      ~ 0,
               site %in% na_sites_sev_only ~ NA_real_,
               TRUE                        ~ .x
             ))
  )



```

# StreamCat pulls for the rest of the variables of interest
```{r}
unique_comids <- unique(sites_sev_bp_geo$comid)

# ---- Define metrics by category --------------------------------
# See full list: sc_get_params(param = "name")

metrics_list <- c(
  # NLCD 2019 (most recent) - land cover %
  "pcturbop2019", "pcturblo2019", "pcturbmd2019", "pcturbhi2019",
  "pctdecid2019", "pctconif2019", "pctmxfst2019",
  "pctshrb2019", "pctgrs2019", "pcthay2019", "pctcrop2019",
  "pctwdwet2019", "pcthbwet2019", "pctbl2019", "pctice2019", "pctow2019",
  
  # BFI - base flow index
  "bfi",
  
  # NADP - atmospheric deposition (2008)
  "inorgnwetdep_2008", "sn_2008", "no3_2008", "nh4_2008",
  
  # Reference stream temperature
  "mast_2014", "mwst_2014", "msst_2014",   # mean annual/winter/summer stream temp
  
  # Runoff
  "runoff",
  
  # STATSGO soils (Set 1)
  "clay", "sand", "wtdep", "om", "perm", "rckdep",
  
  # STATSGO soils (Set 2)
  "hydrlcond", "kffact",
  
  # Lithology
  "pctalkintruvol", "pctsilicic", "pctcarbresid", "pctnoncarbresid",
  "pctsallake", "pctalluvcoast", "pctcolluvsed", "pctglactilclay",
  "pctglactilcrs", "pctglactilloam", "pctglaclakecrs", "pctglaclakefine",
  "pctglacoutcrs", "pctglacoutfine", "pcthydric", "pcteolcrs", "pcteolfine",
  "pctsaline", "pctwater", "pctextruvol", "pctintruvol", "pctmetvol",
  "pctcoastcrs", "pctwatr",
  
  # PRISM climate (1981-2010 normals)
  "precip8110", "tmean8110", "tmin8110", "tmax8110",
  
  # Surface water (sw) - may refer to SurfaceWater or WetIndex
  "wetindexws"
)

# Pull all at once (watershed-scale)
sc_data <- sc_get_data(
  metric = paste(metrics_list, collapse = ","),
  aoi    = "watershed",
  comid  = paste(unique_comids, collapse = ",")
)

names(sc_data) <- tolower(names(sc_data))
message("Pulled ", ncol(sc_data) - 1, " StreamCat variables for ",
        nrow(sc_data), " COMIDs")

# ---- Join onto sites -------------------------------------------
sites_sc <- sites_sev_bp_geo %>%
  left_join(sc_data, by = "comid")

```


# EXPORT FILE 
```{r}
# ---- Deduplicate: one row per site ---------------------------------
# Some sites get expanded when multiple input sites share a COMID.
# Collapse to a single row per site.
n_before <- nrow(sites_sc)

sites_sc <- sites_sc %>%
  distinct(site, .keep_all = TRUE)

n_after <- nrow(sites_sc)

message("Deduplicated: ", n_before, " rows -> ", n_after, " rows (",
        n_before - n_after, " duplicates removed)")

# Verify no remaining duplicates
dup_check <- sites_sc %>% count(site) %>% filter(n > 1)
stopifnot(nrow(dup_check) == 0)
message("Confirmed: one row per site in final geospatial data frame.")

# ---- Export --------------------------------------------------------
out_dir <- here("Output_for_analysis/02_geospatial_comid_extraction")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
write_csv(sites_sc,
          file.path(out_dir, "geospatial_variables_bp_severity_pull.csv"))


```



# GAVIOTA TEST 
```{r}
library(nhdplusTools)
library(sf)
library(ggplot2)
library(maps)

# Define your COMID of interest
my_comid <- 17594969

# Set up the NLDI feature list
nldi_feature <- list(featureSource = "comid", featureID = as.integer(my_comid))

# Retrieve the watershed basin polygon
watershed_basin <- get_nldi_basin(nldi_feature = nldi_feature)

# Quick base R plot
plot(st_geometry(watershed_basin), col = "lightblue", border = "darkblue")

# Or plot using ggplot2
ggplot(watershed_basin) +
  geom_sf(fill = "lightblue", color = "darkblue", alpha = 0.5) +
  theme_minimal() +
  labs(title = paste("Watershed Basin for COMID:", my_comid))

# 1. Fetch the watershed basin polygon
my_comid <- 17594969 # Replace with your target COMID
nldi_feature <- list(featureSource = "comid", featureID = as.integer(my_comid))
watershed_basin <- get_nldi_basin(nldi_feature = nldi_feature)

# 2. Get the USA state boundaries and convert to an sf object
usa_states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))

# 3. Match Coordinate Reference Systems (CRS) to ensure perfect overlay
# Albers Equal Area (EPSG: 5070) is ideal for nationwide USA mapping
target_crs <- 5070
usa_states <- st_transform(usa_states, crs = target_crs)
watershed_basin <- st_transform(watershed_basin, crs = target_crs)

# 4. Generate the plot
ggplot() +
  # Draw the USA base map
  geom_sf(data = usa_states, fill = "gray95", color = "gray80")

ggplot() +
  # Draw the USA base map
  geom_sf(data = usa_states, fill = "gray95", color = "gray80") +
  # Overlay the watershed basin (bright color makes it visible at a national scale)
  geom_sf(data = watershed_basin, fill = "red", color = "darkred", size = 1.5) +
  theme_minimal() +
  labs(
    title = paste("Location of Watershed COMID", my_comid, "in the USA"),
    x = "Longitude", 
    y = "Latitude"
  )

```

```{r}
library(nhdplusTools)
library(sf)
library(ggplot2)

# Start with your current (incorrect/too far downstream) COMID
wrong_comid <- 17594969 

# 1. Trace all upstream tributaries (this finds all nested stream segments)
upstream_network <- navigate_nldi(
  nldi_feature = list(featureSource = "comid", featureID = as.integer(wrong_comid)),
  mode = "upstreamTributaries", # Traces every single branch upstream
  data_source = "flowlines"
)

# 2. View a list of all potential nested COMIDs
nested_comids <- upstream_network$UT_flowlines$nhdplus_comid
print(head(nested_comids))

# 3. Plot them interactively to visually locate your target stream segment
# Click on a segment in the map viewer to see its true nhdplus_comid
library(mapview)
mapview(upstream_network$UT_flowlines, zcol = "nhdplus_comid")



```

```{r}
# Define your COMID of interest
my_comid <- 17594763

# Set up the NLDI feature list
nldi_feature <- list(featureSource = "comid", featureID = as.integer(my_comid))

# Retrieve the watershed basin polygon
watershed_basin <- get_nldi_basin(nldi_feature = nldi_feature)

# Quick base R plot
plot(st_geometry(watershed_basin), col = "lightblue", border = "darkblue")

# Or plot using ggplot2
ggplot(watershed_basin) +
  geom_sf(fill = "lightblue", color = "darkblue", alpha = 0.5) +
  theme_minimal() +
  labs(title = paste("Watershed Basin for COMID:", my_comid))



```


