---
title: "AGU_2023_Burn_Unburn_comp"
output: html_document
date: "2023-12-01"
editor_options: 
  chunk_output_type: console
---

The purpose of this script is to read in the studies from the final meta data list and to normalize daily stream concentrations by time and plot burn vs. unburned concentrations for a figure for my presentation

# Status: in progress

# ==============================================================================
# Author: Jake Cavaiani; jake.cavaiani@pnnl.gov 
# 01 December 2023
# ==============================================================================

## Temporal Normalization 
```{r Jake/Mac Load Packages and}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment


library(tidyverse)
library(here)
library(forecastML)
library(zoo)
# library(hrbrthemes)
library(viridis)
library(car)
library(lme4)
library(merTools)


```

```{r load in publications}
studies_file_list <- list.files(path = "inputs/Studies/meta_final/", 
                                  recursive=F, 
                                  pattern=".csv", 
                                  full.names=TRUE)

storm_list_beta<-do.call("list", lapply(studies_file_list, 
                                        read.csv, 
                                        stringsAsFactors=FALSE, 
                                        header=T))

meta_df <-do.call("rbind", lapply(studies_file_list, 
                                     read.csv,
                                     na = c('-9999', 'N/A'),
                                     check.names = FALSE,
                                     stringsAsFactors=FALSE, 
                                     header=T, blank.lines.skip = TRUE, fill=TRUE)) 

```


```{r format date and units}
options(max.print = 40)
# Filtering by only the control vs.impact studies to make sure we have no paper that is included that is any other study design. 

# Crandall is filtered out because the burned and reference sites are not paired so I do that is a separate script and will read that csv in later to merge to this dataframe to get the more accurate effect size calculations 

# The studies that monitor streams more than 5 years following a wildfire are also excluded for this initial code and will be added in later.  

control_impact <- meta_df %>% 
  filter(Design_Control_Burn_Pre_Post == "Control_Reference_vs_Impact",
         Study_ID != "Crandall et al. 2021",
         Mean_Median_or_IndividualSample == "Individual",
         as.numeric(Time_Since_Fire) < 6)

# separate out the time columns 
time_format <- control_impact %>% 
  select(Study_ID, Sampling_Date) 

# merge this dataframe with the metadata df to make sure all the dates are the same format and there aren't any NAs
control_impact <- control_impact %>% 
  mutate(Sampling_Date = time_format$Sampling_Date)


# changing the structure of the concentrations
control_impact <- control_impact %>% 
  mutate(DOC = as.numeric(DOC),
         NO3 = as.numeric(NO3), 
         Sampling_Date = ymd(Sampling_Date))

# Changing the units to make sure everything is consistent in mg_N_L or mg_C_L
control_impact_units <- control_impact %>% 
  mutate(Area_watershed_km = case_when(Area_unit == 'ha' ~ Area_watershed * .01,
                                       Area_unit == 'km' ~ print(Area_watershed)),
         
         DOC_mg_C_L = case_when(DOC_unit == 'mg_C_L' ~ print(DOC),
                                DOC_unit == 'mg_L' ~ print(DOC),
                                DOC_unit == 'um' ~ DOC * 0.01201),
         
         NO3_mg_N_L = case_when(NO3_unit == 'um' ~ NO3 * 0.014007000,
                                NO3_unit == 'umol_L' ~ NO3 * 0.014007000,
                                NO3_unit == 'umol_NO2_NO3_L' ~ NO3 * 0.014007000,
                                NO3_unit == 'mg_N_L' ~ print(NO3),
                                NO3_unit == 'ug_N_L' ~ NO3 * .001, 
                                NO3_unit == 'mg_L' ~ NO3 * 0.225904780),
         
         DOC_uM_C = case_when(DOC_unit == 'um' ~ print(DOC),
                              DOC_unit == 'mg_C_L' ~ DOC * 83.2639467,
                              DOC_unit == 'mg_L' ~ DOC * 83.2639467),

         NO3_uM_N = case_when(NO3_unit == 'um' ~ print(NO3),
                              NO3_unit == 'umol_L' ~ print(NO3),
                              NO3_unit == 'umol_NO2_NO3_L' ~ print(NO3),
                              NO3_unit == 'mg_L' ~ NO3 * 16.127729,
                              NO3_unit == 'mg_N_L' ~ NO3 * 3.64145101,
                              NO3_unit == 'ug_N_L' ~ NO3 * 0.22594948))

# Creating a site characteristics data frame that I will use to merge later to add in the proper data 
site_data <- control_impact_units %>% 
  select(Study_ID, Pair, Site, latitude, longitude, Area_watershed_km, Climate, Burn_Unburn)

site_data_unique <- site_data %>% 
  group_by(Study_ID, Pair, Area_watershed_km) %>% 
  distinct(Area_watershed_km, .keep_all = TRUE)

```


```{r - gap fill}
# filling in daily time series and interpolating by study and site
control_impact_fill <-  control_impact_units %>% 
  select("Study_ID", "latitude", "longitude", "Area_watershed_km", "Pair", "Climate", "Site", "Burn_Unburn", "Sampling_Date", "DOC_mg_C_L", "NO3_mg_N_L") %>%
  group_by(Study_ID, Site) %>%
  complete(Sampling_Date = seq(min(Sampling_Date), max(Sampling_Date), by = "1 day")) %>% 
  fill(Study_ID:Burn_Unburn) %>%
  fill(Study_ID, Site) %>%
  ungroup() %>%
  group_by(Site) %>%
  mutate(NO3_Interp = na.approx(NO3_mg_N_L, na.rm = FALSE),
         DOC_Interp = na.approx(DOC_mg_C_L, na.rm = FALSE)) %>%
  fill(Study_ID:Burn_Unburn)


control_impact_fill <- control_impact_fill %>% 
  mutate(year = year(Sampling_Date),
         month = month(Sampling_Date),
         day = day(Sampling_Date))
  

```


```{r calculating the temporal normalized data}
# pivoting to make the responses in one column
metadata_long <- control_impact_fill %>%
  pivot_longer(
    cols = NO3_Interp:DOC_Interp,
    names_to = "response_var",
    values_to = "concentration",
    values_drop_na = TRUE
  )


# aggregate by year 
yearly_mean_site <-  metadata_long %>%
  group_by(Study_ID, response_var, Pair, year) %>% 
  summarize(meanConc = mean(concentration)) %>% 
  left_join(site_data) %>% 
  group_by(year) %>% 
  distinct(meanConc, .keep_all = TRUE) %>% 
  mutate(TempNorm = meanConc / Area_watershed_km) 

yearly_mean_site$TempNorm <- format(yearly_mean_site$TempNorm, scientific = F)

yearly_mean_site <-  yearly_mean_site %>%
  mutate(TempNorm = as.numeric(TempNorm))


```


# Add in corrected Crandall and papers that are greater than 5 years to this list 
```{r}
# combine the conc. fill of Crandall, Abbott, and Rhea with the rest of it 

crandall_conc_fill <- read_csv(here("Output_for_analysis", "01_Crandall_effect_size", "crandall_temp_norm.csv"), na = c('-9999', 'N/A')) 

abbott_conc_fill <- read_csv(here("Output_for_analysis", "02_Abbott_effect_size", "Abbott_temp_norm.csv"), na = c('-9999', 'N/A')) 

rhea_conc_fill <- read_csv(here("Output_for_analysis", "03_Rhea_effect_size", "Rhea_temp_norm.csv"), na = c('-9999', 'N/A')) 


yearly_mean_site <- rbind(yearly_mean_site, crandall_conc_fill, abbott_conc_fill, rhea_conc_fill)

write_csv(yearly_mean_site, here("Output_for_analysis", "09_Temp_Norm_comp", "non-normalization_effect_size.csv"))


```



```{r Calculate CV}
metadata_cv <- yearly_mean_site %>% 
  group_by(Study_ID, response_var, Burn_Unburn) %>% 
  summarise(meanConc = mean(TempNorm, na.rm = TRUE),
            sdconc = sd(TempNorm, na.rm = TRUE),
            CVconc = sdconc/meanConc) # generating CVs

doc_cv <- metadata_cv %>% 
  filter(response_var == "DOC_Interp")

# DOC 
doc_yearly <- yearly_mean_site %>% 
  filter(response_var == "DOC_Interp")

# conduct anova test to see if the mean concentrations were different between burn and unbunr
one.way.doc <- aov(TempNorm ~ Burn_Unburn, data = doc_yearly)

plot(one.way.doc)
# QQ plot is not great. Looks like I should log transform. Lets try that. 

doc_yearly <- doc_yearly %>% 
  mutate(logTempNorm = log(TempNorm))

log.one.way.doc <- aov(logTempNorm ~ Burn_Unburn, data = doc_yearly)
plot(log.one.way.doc)

summary(log.one.way.doc) # F-value: 1.113, p-value: 0.294

# conduct Levene's Test for equality of variances
leveneTest(TempNorm ~ Burn_Unburn, data = doc_yearly) #F-value: 0.6377, p-value: 0.4266,


# NO3 
no3_yearly <- yearly_mean_site %>% 
  filter(response_var == "NO3_Interp")

no3_cv <- metadata_cv %>% 
  filter(response_var == "NO3_Interp")


# conduct anova test to see if the mean concentrations were different between burn and unbunr
one.way.no3 <- aov(TempNorm ~ Burn_Unburn, data = no3_yearly)

plot(one.way.no3)

# QQ plot is not great. Looks like I should log transform. Lets try that. 

no3_yearly <- no3_yearly %>% 
  mutate(logTempNorm = log(TempNorm))

log.one.way.no3 <- aov(logTempNorm ~ Burn_Unburn, data = no3_yearly)
plot(log.one.way.no3)

summary(log.one.way.no3) # df: 1,181, F-value: 0.152, p-value: 0.697



# conduct Levene's Test for equality of variances
leveneTest(TempNorm ~ Burn_Unburn, data = no3_yearly) # df: 1,181, F-value: 0.7465,  p-value: 0.38 

# Write csv's
doc_yearly_out <- doc_yearly %>%
  mutate(across(where(is.character), ~ifelse(is.na(.), "N/A", .))) %>%
  mutate(across(where(is.numeric), ~ifelse(is.na(.), -9999, .)))

write_csv(doc_yearly_out, here("Output_for_analysis", "09_Temp_Norm_comp", "DOC_temp_norm_fig_data.csv"))

doc_cv_out <- doc_cv %>%
  mutate(across(where(is.character), ~ifelse(is.na(.), "N/A", .))) %>%
  mutate(across(where(is.numeric), ~ifelse(is.na(.), -9999, .)))

write_csv(doc_cv_out, here("Output_for_analysis", "09_Temp_Norm_comp", "DOC_CV_temp_norm_fig_data.csv"))

no3_yearly_out <- no3_yearly %>%
  mutate(across(where(is.character), ~ifelse(is.na(.), "N/A", .))) %>%
  mutate(across(where(is.numeric), ~ifelse(is.na(.), -9999, .)))

write_csv(no3_yearly_out, here("Output_for_analysis", "09_Temp_Norm_comp", "NO3_temp_norm_fig_data.csv"))

no3_cv_out <- no3_cv %>%
  mutate(across(where(is.character), ~ifelse(is.na(.), "N/A", .))) %>%
  mutate(across(where(is.numeric), ~ifelse(is.na(.), -9999, .)))

write_csv(no3_cv_out, here("Output_for_analysis", "09_Temp_Norm_comp", "NO3_CV_temp_norm_fig_data.csv"))

```

### PLOT Burn vs. Unburn ###

```{r}
# DOC 
# simple box plot 
ggplot(doc_yearly, aes(Burn_Unburn, TempNorm, fill = Burn_Unburn)) +
  geom_boxplot() +
  theme_bw()
  
# Boxplot with density plot   
ggplot(doc_yearly, aes(x = Burn_Unburn, y = TempNorm, fill = Burn_Unburn)) +
  scale_fill_manual(values = c("darkred", "#69b3a2")) +
  labs(x =  " ", y = expression(paste("mg  ", DOC ," ", L^-1, km^-2, ""))) +
  ggdist::stat_halfeye(adjust = 0.5, 
                       justification = -.3,
                       .width = 0,
                       point_colour = NA,
                       scale = 0.5) +
  geom_boxplot(width = .25,
                 outlier.colour = NA,
                 alpha = 0.5) +
  scale_y_continuous(trans='log10') +
  ggtitle("") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 25),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 25))

ggsave("DOC_yield.pdf",
       path = here("initial_plots", "09_Temp_Norm_comp"),
       width = 6, height = 6,  units = "in")

# Box plot with CV
ggplot(doc_cv, aes(x = Burn_Unburn, y = CVconc, fill = Burn_Unburn)) +
  scale_fill_manual(values = c("darkred", "#69b3a2")) +
  labs(y = "CV", x = "") +
  ggdist::stat_halfeye(adjust = 0.5, 
                       justification = -.2,
                       .width = 0,
                       point_colour = NA) +
  geom_boxplot(width = .25,
                 outlier.colour = NA,
                 alpha = 0.5) +
  ggtitle("Coefficient of Variation for DOC") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 25),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25))

ggsave("DOC_CV_yield.pdf",
       path = here("initial_plots", "09_Temp_Norm_comp"),
       width = 6, height = 6,  units = "in")        

# NO3
vn = expression(paste("mg " *N*O[3]^"-",  L^-1, km^-2, ""))

# Boxplot with density plot   
ggplot(no3_yearly, aes(x = Burn_Unburn, y = TempNorm, fill = Burn_Unburn)) +
  scale_fill_manual(values = c("darkred", "#69b3a2")) +
  ggdist::stat_halfeye(adjust = 0.5, 
                       justification = -.3,
                       .width = 0,
                       point_colour = NA,
                       scale = 0.5) +
  geom_boxplot(width = .25,
                 outlier.colour = NA,
                 alpha = 0.5) +
  scale_y_continuous(trans='log10') +
  ylab(vn) +
  xlab("") +
  ggtitle("") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 25),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 25))

ggsave("NO3_yield.pdf",
       path = here("initial_plots", "09_Temp_Norm_comp"),
       width = 6, height = 6,  units = "in")

# Box plot with CV
ggplot(no3_cv, aes(x = Burn_Unburn, y = CVconc, fill = Burn_Unburn)) +
  scale_fill_manual(values = c("darkred", "#69b3a2")) +
  labs(y = "CV", x = "") +
  ggdist::stat_halfeye(adjust = 0.5, 
                       justification = -.2,
                       .width = 0,
                       point_colour = NA) +
  geom_boxplot(width = .25,
                 outlier.colour = NA,
                 alpha = 0.5) +
  ggtitle("Coefficient of Variation for Nitrate") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 25),
        axis.text.y = element_text(size = 25),
        axis.title.y = element_text(size = 25))

ggsave("NO3_CV_yield.pdf",
       path = here("initial_plots", "09_Temp_Norm_comp"),
       width = 6, height = 6,  units = "in")





```


```{r - report summary stats}
# DOC
doc_summary <- doc_yearly %>%
  group_by(Burn_Unburn) %>% 
  summarise(meanDOC = mean(TempNorm),
            sdDOC = sd(TempNorm),
            range = range(TempNorm))
# Burn mean: 4.661536
# Burn sd: 11.02459
# Burn range: [0.010632809:79.805081141]

# Unburn mean: 6.511373
# Unburn sd: 12.59621
# Unburn range: [0.005674765:52.736711286]

# NO3
no3_summary <- no3_yearly %>%
  group_by(Burn_Unburn) %>% 
  summarise(meanDOC = mean(TempNorm),
            sdDOC = sd(TempNorm),
            range = range(TempNorm, scientific = F))

# Burn mean: 2.527080
# Burn sd: 10.383352
# Burn range: [0:87.40485]

# Unburn mean: 1.339846
# Unburn sd: 3.505517
# Unburn range: [0:15.14391]


```

# ==============================================================================
```{r TSF df TEST}
TSF_ES_Fig_data <- read_csv("Output_for_analysis/04_Meta_merge_all_studies_effect_size/TSF_ES_Fig_data.csv")

sort(unique(TSF_ES_Fig_data$Study_ID)) # 17 studies This matches the df down below but there are only 139 observations here and 237 down there -WHY?

ounted_data <- TSF_ES_Fig_data %>%
  group_by(Study_ID, response_var) %>%
  summarise(row_count = n(), .groups = 'drop')

counted_data <- yearly_mean_site %>%
  group_by(Study_ID, response_var) %>%
  summarise(row_count = n(), .groups = 'drop')

merged_data <- cbind(ounted_data, counted_data)

```

# 12/19/2025 - resubmission stats
```{r - report summary stats - non-normalized}
# This chunk is comparing the meanConc (non-normalized) of DOC and nitrate between burned and unburned watersheds (These values are NOT the effect size, its just the raw conc. averaged per site/year/study_ID)

# Step 1: Run above code until 192 and then remove all the data frames you don't need anymore
rm(abbott_conc_fill)
rm(control_impact)
rm(control_impact_fill)
rm(control_impact_units)
rm(crandall_conc_fill)
rm(meta_df)
rm(metadata_long)
rm(rhea_conc_fill)
rm(site_data)
rm(site_data_unique)
rm(storm_list_beta)
rm(time_format)

View(yearly_mean_site)
yearly_mean_site <- yearly_mean_site %>% 
  filter(Study_ID !="Murphy et al. 2015")
sort(unique(yearly_mean_site$Study_ID)) # 17 studies

# meanConc is the non-normalized mean concentrations of DOC and nitrate for each site and year so I should be able to do the same triage but the response variable is meanConc rather than TempNorm

# group by study_ID, response_var, Burn_Unburn
metadata_cv <- yearly_mean_site %>% 
  group_by(Study_ID, response_var, Burn_Unburn) %>% 
  summarise(meanConc = mean(TempNorm, na.rm = TRUE),
            sdconc = sd(TempNorm, na.rm = TRUE),
            CVconc = sdconc/meanConc) # generating CVs

# Separate out DOC 
doc_cv <- metadata_cv %>% 
  filter(response_var == "DOC_Interp")

# DOC 
doc_yearly <- yearly_mean_site %>% 
  filter(response_var == "DOC_Interp")

# conduct anova test to see if the mean concentrations were different between burn and unburned for the non-normalized data (meanConc)
one.way.doc <- aov(meanConc ~ Burn_Unburn, data = doc_yearly)

plot(one.way.doc)

# QQ plot is not great. Looks like I should log transform. Lets try that. 
doc_yearly <- doc_yearly %>% 
  mutate(logmeanConc = log(meanConc))

log.one.way.doc <- aov(logmeanConc ~ Burn_Unburn, data = doc_yearly)
plot(log.one.way.doc)

summary(log.one.way.doc) # F-value: 0.757, p-value: 0.387


# DOC
doc_summary <- doc_yearly %>%
  group_by(Burn_Unburn) %>% 
  reframe(meanDOC = mean(meanConc),
            sdDOC = sd(meanConc),
            range = range(meanConc))
# Burn mean: 5.464236
# Burn sd: 7.028624
# Burn range: [0.8471739:28.6974038]

# Unburn mean: 5.192652
# Unburn sd: 6.565883
# Unburn range: [0.4378601:24.5090775]

# The mean NON-NORMALIZED (meanConc) concentrations for DOC (5.46 (mg DOC-C)) for burned and 5.19 (mg DOC-C) for unburned were not significantly different from each other (one-way ANOVA: F-value = 0.757; df = 1,72; p-value = 0.387; Fig. S7).

# NO3
no3_yearly <- yearly_mean_site %>% 
  filter(response_var == "NO3_Interp")

# conduct anova test to see if the mean concentrations were different between burn and unburned for the non-normalized data (meanConc)
one.way.n03 <- aov(meanConc ~ Burn_Unburn, data = no3_yearly)

plot(one.way.n03)

# QQ plot is not great. Looks like I should log transform. Lets try that. 
no3_yearly <- no3_yearly %>% 
  mutate(logmeanConc = log(meanConc))

log.one.way.no3 <- aov(logmeanConc ~ Burn_Unburn, data = no3_yearly)
plot(log.one.way.no3)

summary(log.one.way.no3) # F-value: 7.823, p-value: 0.00579 - Significant?


no3_summary <- no3_yearly %>%
  group_by(Burn_Unburn) %>% 
  reframe(meanNO3 = mean(TempNorm),
            sdNO3 = sd(TempNorm),
            range = range(TempNorm, scientific = F))

# Burn mean: 0.11068395
# Burn sd: 0.5803534
# Burn range: [0.000000:4.870437]

# Unburn mean: 0.06131053
# Unburn sd: 0.2377624
# Unburn range: [0:1.499988]

```

```{r normalized/non-normalized DOC and nitrate effect sizes across whole dataset -- ANOVA }
Effect_size_orig_nonnormal <- read_csv("Output_for_analysis/04_Meta_merge_all_studies_effect_size/Effect_size_orig_nonnormal.csv")

sort(unique(Effect_size_orig_nonnormal$Study_ID)) # 17 studies - this is all of them

# Pivot into long format
long_data <- Effect_size_orig_nonnormal %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_data$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_data %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# The results revealed significant differences between the effect types for `DOC_Interp`, while no significant differences were observed for `NO3_Interp`. For the response variable `DOC_Interp`, the effect type (`Effect_size` vs. `Effect_size_no_watershed`) showed a significant effect on the response (F(1, 86) = 21.23, *p* < 0.001). The mean sum of squares for the effect type was considerably larger (11.15) compared to the residual variance (0.525). These results indicate that the adjustment for watershed markedly influences effect sizes for this response variable.

# In contrast, for `NO3_Interp`, the ANOVA results showed no significant difference between effect types (F(1, 168) = 2.053, *p* = 0.154). The mean sum of squares for the effect type was considerably smaller (3.49) relative to the residual variance (1.701). This suggests that watershed adjustments do not have a statistically meaningful impact on effect sizes for this response variable.

```

```{r Non-normalized effect size through time}
#### DOC NON-NORMAL ####
# Filter data for DOC and nitrate separately
DOC_data <- Effect_size_orig_nonnormal %>% filter(response_var == "DOC_Interp")

# Read in site data to merge the sites to the DOC_data data frame 
study_site_data <- read_csv(here("inputs", "catchment_characteristics", "Fire_name_Lat_Long.csv"), na = c('-9999', 'N/A')) %>% 
  dplyr::select(Study_ID, Site, Fire_year, Effect_size_pair) %>% 
  rename(year = Fire_year) %>% 
  na.omit()

unique(sort(study_site_data$Study_ID)) # should be all 17 studies. 

DOC_model <- full_join(DOC_data, study_site_data) %>% 
  filter(!is.na(Effect_size_no_watershed))# joining the TSF dataframe used to make the figure and study site data to match sites because we will be including site as a random effect

# Want to make the TSF continuous so I am gonna put actual years to them 
DOC_model <- DOC_model %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire),
    Site = case_when(Study_ID == "Abbott et al. 2021" ~ "Burn",
                                     TRUE ~ Site)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire))
  

# perform linear mixed effect model for non-normalized effect size # Site as a random effect 
doc_tsf_mixed_nonnormal <-  lmer(Effect_size_no_watershed ~ Time_Since_Fire + (1 | Site), data = DOC_model)


# Check for equal variance
performance::check_heteroscedasticity(doc_tsf_mixed_nonnormal) #Variance error

# Check for normality
residuals <- resid(doc_tsf_mixed_nonnormal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.97841, p-value = 0.5004: p-value is > 0.05
# when p > 0.05 can assume normality 


doc_tsf_mixed_nonnormal

summary(doc_tsf_mixed_nonnormal) 
fixef(doc_tsf_mixed_nonnormal)
confint(doc_tsf_mixed_nonnormal)

# The fixed effects revealed a significant negative relationship between Time_Since_Fire and Effect_size_no_watershed (t = -5.238, p < 0.05). Specifically, the fixed-effect estimate indicates that for each unit increase in Time_Since_Fire, Effect_size_no_watershed decreases by 0.05603 (± 0.01070 standard error). Additionally, the intercept suggests an initial estimated effect size of 0.40913 (± 0.07692 standard error).

# The random effects analysis showed that variability among sites contributed a variance of 0.06683, with a standard deviation of 0.2585. The residual variance was 0.02049, with a standard deviation of 0.1432. This suggests that site-level differences were modest but contributed meaningfully to the overall model structure.

# Overall, these results demonstrate that Time_Since_Fire has a significant and consistent negative effect on Effect_size_no_watershed, with site-specific variability included as a random effect in the model framework

#### DOC NORMAL (orig) ####
# perform linear mixed effect model for non-normalized effect size # Site as a random effect 
doc_tsf_mixed_normal <-  lmer(Effect_size ~ Time_Since_Fire + (1 | Site), data = DOC_model)

# Check for equal variance
performance::check_heteroscedasticity(doc_tsf_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.987).

# Check for normality
residuals <- resid(doc_tsf_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.97094, p-value = 0.2639: p-value is > 0.05
# when p > 0.05 can assume normality 

doc_tsf_mixed_normal

summary(doc_tsf_mixed_normal) 

# The fixed effects showed a negative relationship between Time_Since_Fire and Effect_size, but the effect was not statistically significant (t = -1.420, p > 0.05). The fixed-effect estimate suggests that for each unit increase in Time_Since_Fire, Effect_size decreases by 0.02677 (± 0.01886 standard error). The intercept value of -0.33942 (± 0.27074 standard error) suggests a slightly negative starting point, but it is also non-significant (t = -1.254).

# The random effects analysis revealed notable variability between sites, with a variance of 1.27387 and a standard deviation of 1.1287. This indicates that site-level differences contribute substantially to the overall variation in Effect_size. In comparison, the residual variance was much smaller at 0.01956, with a standard deviation of 0.1399, suggesting that most of the variability is captured by the random site effect.

# In summary, the analysis did not detect a significant relationship between Time_Since_Fire and Effect_size. However, the random effects highlight substantial variability across study sites, indicating that site-specific factors play a critical role in modulating the observed effect sizes. This underscores the importance of accounting for site-level differences in the analysis.

# DOC model comparison #
# The nonnormal model (Effect_size_no_watershed) identifies a significant negative effect of Time_Since_Fire on the response variable, while the normal model (Effect_size) does not detect a significant relationship.

# The normal model attributes much greater variability to site-level differences, suggesting that Effect_size varies substantially across sites. The nonnormal model captures less site-level variability, suggesting that Effect_size_no_watershed may better account for systematic site differences. Using Effect_size_no_watershed (nonnormal) emphasizes a significant negative relationship with Time_Since_Fire, whereas using Effect_size (normal) dilutes this relationship due to substantial variability across sites.

#### NO3 NON-NORMAL ####
Nitrate_data <- Effect_size_orig_nonnormal %>% 
  filter(response_var == "NO3_Interp")

NO3_model <- full_join(Nitrate_data, study_site_data) %>% 
  filter(!is.na(Effect_size_no_watershed))# joining the TSF dataframe used to make the figure and study site data to match sites because we will be including site as a random effect

# Want to make the TSF continuous so I am gonna put actual years to them 
NO3_model <- NO3_model %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire),
    Site = case_when(Study_ID == "Abbott et al. 2021" ~ "Burn",
                                     TRUE ~ Site)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire))
  

# perform linear mixed effect model for non-normalized effect size # Site as a random effect 
no3_tsf_mixed_nonnormal <-  lmer(Effect_size_no_watershed ~ Time_Since_Fire + (1 | Site), data = NO3_model)

# Check for equal variance
performance::check_heteroscedasticity(no3_tsf_mixed_nonnormal) # OK: Error variance appears to be homoscedastic (p = 0.287).

# Check for normality
residuals <- resid(no3_tsf_mixed_nonnormal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.93737, p-value = 0.0003269: p-value is < 0.05
# when p > 0.05 can assume normality 
# NOT NORMAL?

no3_tsf_mixed_nonnormal

summary(no3_tsf_mixed_nonnormal) 
fixef(no3_tsf_mixed_nonnormal)
confint(no3_tsf_mixed_nonnormal)

# The fixed effects revealed a significant positive relationship between Time_Since_Fire and Effect_size_no_watershed (t = 2.694, p < 0.05). Specifically, the fixed-effect estimate indicates that for each unit increase in Time_Since_Fire, Effect_size_no_watershed increases by 0.09152 (± 0.03397 standard error). The intercept indicates an initial estimated effect size of 0.70308 (± 0.17360 standard error).

# The random effects analysis showed that site-level variability contributed a variance of 0.4472, with a standard deviation of 0.6687. The residual variance was higher at 0.5603, with a standard deviation of 0.7485, suggesting that variability within sites was substantial relative to between-site differences.

# In summary, these results demonstrate that Time_Since_Fire has a significant positive effect on Effect_size_no_watershed, with site-specific variability included as a random effect. This positive relationship implies that as time since fire increases, the influence on Effect_size_no_watershed becomes increasingly pronounced.

#### NO3 NORMAL (orig) ####
# perform linear mixed effect model for non-normalized effect size # Site as a random effect 
no3_tsf_mixed_normal <-  lmer(Effect_size ~ Time_Since_Fire + (1 | Site), data = NO3_model)

# Check for equal variance
performance::check_heteroscedasticity(no3_tsf_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.923).

# Check for normality
residuals <- resid(no3_tsf_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.95725, p-value = 0.005135: p-value is < 0.05
# when p > 0.05 can assume normality 
# NOT NORMAL?

no3_tsf_mixed_normal

summary(no3_tsf_mixed_normal) 
# The fixed effects showed a positive relationship between Time_Since_Fire and Effect_size, although the effect was marginally non-significant (t = 1.827, p > 0.05). The fixed-effect estimate suggests that for each unit increase in Time_Since_Fire, Effect_size increases by 0.08813 (± 0.04824 standard error). The intercept indicates an initial estimated effect size of 0.31320 (± 0.30045 standard error).

# The random effects analysis demonstrated notable variability across sites, with a random-effect variance of 2.1417 and a standard deviation of 1.4634. Residual variance accounted for less variability, with a value of 0.5006 and a standard deviation of 0.7076. This indicates that site-level differences contribute substantially to the variance in Effect_size overall.

# In summary, while the positive relationship between Time_Since_Fire and Effect_size was detectable, the results were marginally non-significant. Site-specific variability was substantial, underscoring the importance of accounting for between-site differences in the analysis.

# * Comparison between normal and non-normal* 
#The non-normal model (Effect_size_no_watershed) detected a significant positive relationship between Time_Since_Fire and the response variable, while the normal model (Effect_size) did not. This difference suggests that using the non-normalized data may obscure the significance of the relationship, whereas the transformation or adjustments applied to Effect_size_no_watershed make this relationship clearer.

```

```{r Non-normalized effect size - Climate}
#### DOC NON-NORMAL ####
# perform linear mixed effect model for non-normalized effect size # Site as a random effect 
doc_climate_mixed_nonnormal <-  lmer(Effect_size_no_watershed ~ as.factor(Climate) + (1 | Site), data = DOC_model)

# Check for equal variance
performance::check_heteroscedasticity(doc_climate_mixed_nonnormal) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_climate_mixed_nonnormal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.97404, p-value = 0.3477: p-value is > 0.05
# when p > 0.05 can assume normality 

summary(doc_climate_mixed_nonnormal) 
summary_output <- summary(doc_climate_mixed_nonnormal) 
summary_df <- as.data.frame(summary_output$doc_climate_mixed_nonnormal)

# post hoc test 
emmeans::lsmeans(doc_climate_mixed_nonnormal, pairwise ~ Climate, adjust = "tukey") # This is assessing the pair wise comparisons for the rest of the climate classifications. In this case there were no pairs that were significantly different from each other 

# conduct Levene's Test for equality of variances
# Null: the variance amoung the groups is equal. 
leveneTest(Effect_size_no_watershed ~ as.factor(Climate), data = DOC_model) #F-value: 2.616, p-value: 0.03757, Df: 5,43 
# Variability is significantly different 

# plot 
boxplot(Effect_size_no_watershed ~ Climate,
  data = DOC_model,
  main = "Effect size - non-normal Distribution by Climate - DOC",
  xlab = "Climate",
  ylab = "Effect Size",
  col = "#00AFBB",
  border = "black")


#### DOC NORMAL (orig) ####
doc_climate_mixed_normal <-  lmer(Effect_size ~ as.factor(Climate) + (1 | Site), data = DOC_model)

# Check for equal variance
performance::check_heteroscedasticity(doc_climate_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.988).

# Check for normality
residuals <- resid(doc_climate_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.97153, p-value = 0.2784: p-value is > 0.05
# when p > 0.05 can assume normality 

summary(doc_climate_mixed_normal) 

# post hoc test 
emmeans::lsmeans(doc_climate_mixed_normal, pairwise ~ Climate, adjust = "tukey") # This is assessing the pair wise comparisons for the rest of the climate classifications. In this case there were no pairs that were significantly different from each other 

# conduct Levene's Test for equality of variances
# Null: the variance amoung the groups is equal. 
leveneTest(Effect_size ~ as.factor(Climate), data = DOC_model) #F-value: 40.417, p-value: 3.479e-15, Df: 5,43
# Variability is significantly different 

# plot 
boxplot(Effect_size ~ Climate,
  data = DOC_model,
  main = "Effect size - Distribution by Climate - DOC",
  xlab = "Climate",
  ylab = "Effect Size",
  col = "#00AFBB",
  border = "black")

#### NO3 NON-NORMAL ####
no3_climate_mixed_nonnormal <-  lmer(Effect_size_no_watershed ~ as.factor(Climate) + (1 | Site), data = NO3_model)

# Check for equal variance
performance::check_heteroscedasticity(no3_climate_mixed_nonnormal) # OK: Error variance appears to be homoscedastic (p = 0.583).

# Check for normality
residuals <- resid(no3_climate_mixed_nonnormal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.94414, p-value = 0.0007995: p-value is < 0.05
# when p > 0.05 can assume normality 
# NON-NORMAL?

summary(no3_climate_mixed_nonnormal) 

# post hoc test 
emmeans::lsmeans(no3_climate_mixed_nonnormal, pairwise ~ Climate, adjust = "tukey") # This is assessing the pair wise comparisons for the rest of the climate classifications. In this case there were no pairs that were significantly different from each other 

# conduct Levene's Test for equality of variances
# Null: the variance amoung the groups is equal. 
leveneTest(Effect_size_no_watershed ~ as.factor(Climate), data = NO3_model) #F-value: 0.8502, p-value: 0.5618, Df: 8,80

# plot 
boxplot(Effect_size_no_watershed ~ Climate,
  data = NO3_model,
  main = "Effect size - non-normal Distribution by Climate - NO3",
  xlab = "Climate",
  ylab = "Effect Size",
  col = "#E7B800",
  border = "black")

#### NO3 NORMAL ####
no3_climate_mixed_normal <-  lmer(Effect_size ~ as.factor(Climate) + (1 | Site), data = NO3_model)

# Check for equal variance
performance::check_heteroscedasticity(no3_climate_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.969).

# Check for normality
residuals <- resid(no3_climate_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.94782, p-value = 0.001324: p-value is < 0.05
# when p > 0.05 can assume normality 
# NON-NORMAL?

summary(no3_climate_mixed_normal) 

# post hoc test 
emmeans::lsmeans(no3_climate_mixed_normal, pairwise ~ Climate, adjust = "tukey") # This is assessing the pair wise comparisons for the rest of the climate classifications. In this case there were two pairs that were significantly different: Csb - Dfc, Csa - Dfc

# conduct Levene's Test for equality of variances
# Null: the variance amoung the groups is equal. 
leveneTest(Effect_size ~ as.factor(Climate), data = NO3_model) #F-value: 2.2233, p-value: 0.03401, Df: 8,80

# plot 
boxplot(Effect_size ~ Climate,
  data = NO3_model,
  main = "Effect size - normal Distribution by Climate - NO3",
  xlab = "Climate",
  ylab = "Effect Size",
  col = "#E7B800",
  border = "black")

```











