---
title: "Map"
output: html_document
date: "2023-10-19"
editor_options: 
  chunk_output_type: console
---
The purpose of this script is create the map figure 

Script Workflow:

Step 1) Load in the summary metadata sheet that include lat and long

Step 2) Load in the shape files of the burns

Step 3) Plot


# Status: in progress

# ==============================================================================
# Author: Jake Cavaiani; jake.cavaiani@pnnl.gov 
# 19 October 2023
# ==============================================================================

# Figure in the manuscript (3/7/2024)
## Load packages and set working directory
```{r Jake/Mac}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

library(ggplot2)
library(tidyverse)
library(here)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
library(broom)
library(usmap)
library(maps)
library(mapdata)
library(cowplot)
library(sf)
library(RColorBrewer)
library(patchwork)
library(ggspatial)
library(raster)
```

This is the workflow to read in the meta analysis studies and the fires they are studying and then loading in a CONUS map and plotting our data on top of that
```{r load in fire names and associated study}
meta <- read_csv(here("inputs", "Studies_Summary", "Map_input.csv"), na = c('-9999', 'N/A')) # reading in the finalized filtered meta-analysis papers. 

coords <- meta %>% 
 dplyr::select(Fire_name, latitude, longitude, number_of_burn, State, group) # Picking only the columns I want 

# Projection # 
laea_proj <- "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
 
# Create an sf object from the data
coordinates_sf <- st_as_sf(coords, coords = c("longitude", "latitude"), crs = 4269)

```

```{r - US map with AK as an inset map}
# Download the US states shapefile from the Census Bureau
url_us_states <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url_us_states, "inputs/map_shape_files/us_states.zip", mode = "wb")
unzip("inputs/map_shape_files/us_states.zip", exdir = "inputs/map_shape_files/us_states")

# Read the US states shapefile
us_states <- read_sf("inputs/map_shape_files/us_states/cb_2021_us_state_20m.shp") %>% 
  st_transform(laea_proj) %>% 
  filter(NAME != 'Hawaii' & NAME != 'Puerto Rico')

# Check the plot 
ggplot() +
  geom_sf(data = us_states, fill = "white", color = "black")

# Read in Canadian provinces 
canadian_provinces <- sf::st_as_sf(rnaturalearth::ne_states(country = "Canada", returnclass = "sf")) %>% 
  st_transform(laea_proj) 

# Check the plot
ggplot() +
  geom_sf(data = canadian_provinces, fill = "white", color = "black")

# Merge USA and Canada plots
us_ca <- bind_rows(us_states, canadian_provinces)

# Check the plot 
ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black")

# Manuscript plot with fire coordinates with no legend
ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  scale_color_manual(values = c("#E31A1C", "#A6CEE3", "#00AFBB", "#1B9E77",
                                "#7570B3", "#E7298A", "darkred", "black", 
                                "#1F78B4", "#E31A1C", "#33A02C", "black",
                                "darkblue", "darkgreen", "blue", "yellow", "red",
                                "green", "purple")) +
  ggspatial::annotation_scale(
    location = "br",
    pad_x = unit(0.5, "in"), 
    bar_cols = c("black", "white")) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(1.1, "in"), pad_y = unit(0.5, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("black", "white"),
      line_col = "grey20")) +
  theme_minimal() +
  theme(legend.position = "none")
  
ggsave("number_watersheds_map_no_legend.pdf",
       path = "~/Documents/RC_3/Wenas_Watershed_project/Manuscript",
       width = 8, height = 6,  units = "in")

# Manuscript plot with fire coordinates with legend
ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  scale_color_manual(values = c("#E31A1C", "#A6CEE3", "#00AFBB", "#1B9E77",
                                "#7570B3", "#E7298A", "darkred", "black", 
                                "#1F78B4", "#E31A1C", "#33A02C", "black",
                                "darkblue", "darkgreen", "blue", "yellow", "red",
                                "green", "purple"),
                     guide = guide_legend(title = "Fire Name")) +
  labs(size = "Number of Burned Watersheds") +
  ggspatial::annotation_scale(
    location = "br",
    pad_x = unit(0.5, "in"), 
    bar_cols = c("black", "white")) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(1.1, "in"), pad_y = unit(0.5, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("black", "white"),
      line_col = "grey20")) +
  theme_minimal() +
  theme(legend.position = "left")
  
ggsave("number_watersheds_map_legend.pdf",
       path = "~/Documents/RC_3/Wenas_Watershed_project/Manuscript",
       width = 10, height = 8,  units = "in")




```



## Load packages and set working directory
```{r Jake/Mac}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

library(ggplot2)
library(tidyverse)
library(here)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
library(broom)
library(usmap)
library(maps)
library(mapdata)
library(cowplot)
library(sf)
library(RColorBrewer)
library(patchwork)
library(ggspatial)
library(raster)
```

This is the workflow to read in the meta analysis studies and the fires they are studying and then loading in a CONUS map and plotting our data on top of that
```{r load in fire names and associated study}
meta <- read_csv(here("inputs", "Studies_Summary", "Map_input.csv"), na = c('-9999', 'N/A')) # reading in the finalized filtered meta-analysis papers. 

coords <- meta %>% 
 dplyr::select(Fire_name, latitude, longitude, number_of_burn, State, group) # Picking only the columns I want 

coords_noNA <- coords %>% 
    drop_na(Fire_name) # eliminating rows that are NAs if they dont have an associated fire name with it. 

# Projection # 
laea_proj <- "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
 
# Create an sf object from the data
coordinates_sf <- st_as_sf(coords_noNA, coords = c("longitude", "latitude"), crs = 4269)

# common_crs = 4326 - World Geodetic System 1984
# common_crs = 4269 - North American Datum 1983
# common_crs = 3857 - Google maps
# common_crs = 4087 - World Equidistant Cylindrical
common_crs = 5070 # = USA Contiguous Albers Equal Area Conic
# common_crs = 3338 # - Alaska equal

# common_crs = 2264 # - State Plane North Carolina Feet. 

```

```{r - US map with AK as an inset map}
# Download the US states shapefile from the Census Bureau
url_us_states <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url_us_states, "inputs/map_shape_files/us_states.zip", mode = "wb")
unzip("inputs/map_shape_files/us_states.zip", exdir = "inputs/map_shape_files/us_states")

# Download the Alaska shapefile
url_alaska <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url_alaska, "inputs/map_shape_files/alaska.zip", mode = "wb")
unzip("inputs/map_shape_files/alaska.zip", exdir = "inputs/map_shape_files/alaska")

# Read the US states shapefile
us_states <- read_sf("inputs/map_shape_files/us_states/cb_2021_us_state_20m.shp") %>% 
  st_transform(common_crs) %>% 
  filter(NAME != 'Hawaii' & NAME != 'Puerto Rico')
# %>%
#   st_crop(xmin = -180, xmax = -55, ymin = 12, ymax = 62)

ggplot() +
  geom_sf(data = us_states, fill = "white", color = "black")

# Read the Alaska shapefile
alaska <- st_read("inputs/map_shape_files/alaska/cb_2021_us_state_20m.shp")

ak <- ggplot() +
  geom_sf(data = alaska, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  # annotation_north_arrow(location = "tr", which_north = "true", scale = 0.2) +
  annotation_scale(location = "bl", width_hint = 0.1) +
  # xlim(-180, -127) +
  # ylim(50, 72) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid = element_blank(), 
        axis.text.x = element_blank())

print(ak)

# I want to include Canada # 
# Download shapefile for world countries
world <- ne_countries(scale = "medium", returnclass = "sf")

# Filter for Canada
canada <- world[world$iso_a2 == "CA", ]
canadian_provinces <- sf::st_as_sf(rnaturalearth::ne_states(country = "Canada", returnclass = "sf")) %>% 
  st_transform(common_crs) 

ggplot() +
  geom_sf(data = canadian_provinces, fill = "white", color = "black")


# US and Canada plot with all the fires
us_ca <- ggplot() +
  geom_sf(data = canadian_provinces, fill = "white", color = "black") +
  geom_sf(data = us_states, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  annotation_north_arrow(location = "tr", which_north = "true") +
  annotation_scale(location = "br", width_hint = 0.1) +
  scale_color_manual(values = c("#666666", "#A6CEE3", "#00AFBB", "#1B9E77",
                                "#7570B3", "#E7298A", "darkred", "black", 
                                "#1F78B4", "#E31A1C", "#33A02C", "black",
                                "darkblue", "darkgreen", "blue", "yellow", "red",
                                "green", "purple")) +
  xlim(-135, -68) +
  ylim(25, 62) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid = element_blank())

print(us_ca)

# add the AK insert #
us_ca + inset_element(ak, 0.05, -0.75, 0.35, 1.2, align_to = 'full')
  
ggsave("number_watersheds_map.pdf",
       path = here("initial_plots", "10_Map"),
       width = 6, height = 4,  units = "in")


# Bind US and canadian provinces # 
us_ca <- bind_rows(us_states, canadian_provinces)

ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black")

# Set up the Lambert Azimuthal Equal Area projection
laea_proj <- "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  # xlim(-185, -60) +
  # ylim(25, 90) +
  # coord_sf(crs = laea_proj) +
  theme_void() +
  theme(legend.position = "none")

# us_ca_transformed <- st_transform(us_ca, crs = laea_proj)

# ggplot() +
#   geom_sf(data = us_ca_transformed, fill = "white", color = "black") +
#   theme_void()

```

```{r koppen climate classification layer}
# Peter code to plot tif #

## Read in koppen climate classification - only needs to be read in once when plotting and then can be deleted after the kppen_reproject
koppen_raw <-  raster::raster("~/Downloads/1984_2013/kc_1984_2013.tif")

## Re-project koppen map
koppen_reproject <- raster::projectRaster(koppen_raw, crs = 5070)

## Mask (crop in essence) severity to fire boundary
koppen_crop <- raster::mask(koppen_reproject, us_ca)

## Convert to dataframe
koppen <- as.data.frame(koppen_crop, xy = T) %>% 
  as_tibble() %>% 
  rename("long" = x, 
         "lat" = y, 
         "koppen" = kc_1984_2013) %>% 
  filter(!is.na(koppen)) %>% 
  mutate(f_koppen = as.factor(round(koppen, 0)),
         f2_koppen = case_when(f_koppen < 200 ~ "Tropical",
                               f_koppen > 200 & f_koppen < 300 ~ "Arid",
                               f_koppen > 300 & f_koppen < 400 ~ "Temperate",
                               f_koppen > 400 & f_koppen < 500 ~ "Boreal",
                               f_koppen > 500 ~ "Polar"))

# Plot raster 
ggplot() + 
  geom_raster(data = koppen, 
              aes(long, lat, fill = f_koppen), alpha = 0.85, show.legend = FALSE) +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn), color = "black", show.legend = FALSE) +
  annotation_north_arrow(location = "bl", which_north = "true") +
  annotation_scale(location = "br", width_hint = 0.1) +
  theme_map()
  # geom_sf(data = boundary, fill = NA, color = "red", lwd = 0.5) + 
  scale_fill_manual(values = fire_colors) + 
  theme_map() + 
  labs(x = "", y = "", fill = "Severity \n (dNBR)")
  
#
ggsave("koppen_map.pdf",
       path = here("initial_plots", "10_Map"),
       width = 20, height = 20,  units = "in") 
#

#
boundary <- read_sf("~/GitHub/rc_sfa-rc-3-wenas-modeling/data/mtbs/wa4685412079920200831/wa4685412079920200831_20200830_20200909_burn_bndy.shp") %>% 
  st_transform(common_crs)

w10 <- tibble(lat = 46.86752, long = -120.7744, site_id = "W10") %>% 
  st_as_sf(coords = c("long", "lat"), crs = common_crs)

ggplot() + 
  geom_sf(data = boundary) + 
  geom_sf(data = w10)

```

```{r Chat GPT code}
library(raster)
library(sf)

raster_data <- raster("~/Downloads/1984_2013/kc_1984_2013.tif")

# Convert raster to SpatialPolygonsDataFrame
polygons <- rasterToPolygons(raster_data)

# Convert SpatialPolygonsDataFrame to sf object
sf_object <- st_as_sf(polygons)

# Plot the sf object
ggplot() +
  geom_sf(data = sf_object) +
  theme_void()



```


