---
title: "Map"
output: html_document
date: "2023-10-19"
editor_options: 
  chunk_output_type: console
---
The purpose of this script is create the map figure 

Script Workflow:

Step 1) Load in the summary metadata sheet that include lat and long

Step 2) Load in the shape files of the burns

Step 3) Plot


# Status: in progress

# ==============================================================================
# Author: Jake Cavaiani; jake.cavaiani@pnnl.gov 
# 19 October 2023
# ==============================================================================

# Figure in the manuscript (3/7/2024)
## Load packages and set working directory
```{r Jake/Mac}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

library(ggplot2)
library(tidyverse)
library(here)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
library(broom)
library(usmap)
library(maps)
library(mapdata)
library(cowplot)
library(sf)
library(RColorBrewer)
library(patchwork)
library(ggspatial)
library(raster)
```

This is the workflow to read in the meta analysis studies and the fires they are studying and then loading in a CONUS map and plotting our data on top of that
```{r load in fire names and associated study}
meta <- read_csv(here("inputs", "Studies_Summary", "Map_input.csv"), na = c('-9999', 'N/A')) # reading in the finalized filtered meta-analysis papers. 

coords <- meta %>% 
 dplyr::select(Fire_name, latitude, longitude, number_of_burn, State, group) # Picking only the columns I want 

# Projection # 
laea_proj <- "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
 
# Create an sf object from the data
coordinates_sf <- st_as_sf(coords, coords = c("longitude", "latitude"), crs = 4269)

```

```{r - US map with AK as an inset map}
# Download the US states shapefile from the Census Bureau
url_us_states <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url_us_states, "inputs/map_shape_files/us_states.zip", mode = "wb")
unzip("inputs/map_shape_files/us_states.zip", exdir = "inputs/map_shape_files/us_states")

# Read the US states shapefile
us_states <- read_sf("inputs/map_shape_files/us_states/cb_2021_us_state_20m.shp") %>% 
  st_transform(laea_proj) %>% 
  filter(NAME != 'Hawaii' & NAME != 'Puerto Rico')

# Check the plot 
ggplot() +
  geom_sf(data = us_states, fill = "white", color = "black")

# Read in Canadian provinces 
canadian_provinces <- sf::st_as_sf(rnaturalearth::ne_states(country = "Canada", returnclass = "sf")) %>% 
  st_transform(laea_proj) 

# Check the plot
ggplot() +
  geom_sf(data = canadian_provinces, fill = "white", color = "black")

# Merge USA and Canada plots
us_ca <- bind_rows(us_states, canadian_provinces)

# Check the plot 
ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black")

# Manuscript plot with fire coordinates with no legend
ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  scale_color_manual(values = c("#E31A1C", "#A6CEE3", "#00AFBB", "#1B9E77",
                                "#7570B3", "#E7298A", "darkred", "black", 
                                "#1F78B4", "#E31A1C", "#33A02C", "black",
                                "darkblue", "darkgreen", "blue", "yellow", "red",
                                "green", "purple")) +
  ggspatial::annotation_scale(
    location = "br",
    pad_x = unit(0.5, "in"), 
    bar_cols = c("black", "white")) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(1.1, "in"), pad_y = unit(0.5, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("black", "white"),
      line_col = "grey20")) +
  theme_minimal() +
  theme(legend.position = "none")
  
ggsave("number_watersheds_map_no_legend.pdf",
       path = "~/Documents/RC_3/Wenas_Watershed_project/Manuscript",
       width = 8, height = 6,  units = "in")

# Manuscript plot with fire coordinates with legend
ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  scale_color_manual(values = c("#E31A1C", "#A6CEE3", "#00AFBB", "#1B9E77",
                                "#7570B3", "#E7298A", "darkred", "black", 
                                "#1F78B4", "#E31A1C", "#33A02C", "black",
                                "darkblue", "darkgreen", "blue", "yellow", "red",
                                "green", "purple"),
                     guide = guide_legend(title = "Fire Name")) +
  labs(size = "Number of Burned Watersheds") +
  ggspatial::annotation_scale(
    location = "br",
    pad_x = unit(0.5, "in"), 
    bar_cols = c("black", "white")) +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(1.1, "in"), pad_y = unit(0.5, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("black", "white"),
      line_col = "grey20")) +
  theme_minimal() +
  theme(legend.position = "bottom")
  
ggsave("number_watersheds_map_legend.pdf",
       path = "~/Documents/RC_3/Wenas_Watershed_project/Manuscript",
       width = 14, height = 8,  units = "in")




```

#### Addressing comment regarding jittering points ####
```{r}
# Convert the sf object to a data frame for jittering 
coordinates_df <- cbind(st_coordinates(coordinates_sf), st_drop_geometry(coordinates_sf))

ggplot() +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name, alpha = 0.5)) +
  geom_jitter(data = coordinates_df, aes(x = X, y = Y), width = 0.5, height = 0.2)


ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name, alpha = 0.5)) +
  geom_jitter(data = coordinates_df, aes(x = X, y = Y), width = 0.2, height = 0.2) + # Jitter points
  scale_color_manual(values = c("#E31A1C", "#A6CEE3", "#00AFBB", "#1B9E77",
                                "#7570B3", "#E7298A", "darkred", "black", 
                                "#1F78B4", "#E31A1C", "#33A02C", "black",
                                "darkblue", "darkgreen", "blue", "yellow", "red",
                                "green", "purple")) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

```

#### Supplemental map with Koppen biomes #####


## Load packages and set working directory
```{r Jake/Mac}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

library(ggplot2)
library(tidyverse)
library(here)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
library(broom)
library(usmap)
library(maps)
library(mapdata)
library(cowplot)
library(sf)
library(RColorBrewer)
library(patchwork)
library(ggspatial)
library(raster)
```

This is the workflow to read in the meta analysis studies and the fires they are studying and then loading in a CONUS map and plotting our data on top of that
```{r load in fire names and associated study}
meta <- read_csv(here("inputs", "Studies_Summary", "Map_input.csv"), na = c('-9999', 'N/A')) # reading in the finalized filtered meta-analysis papers. 

coords <- meta %>% 
 dplyr::select(Fire_name, latitude, longitude, number_of_burn, State, group) # Picking only the columns I want 

# Create an sf object from the data
coordinates_sf <- st_as_sf(coords, coords = c("longitude", "latitude"), crs = 4269)

common_crs = 4269
```

```{r - US map with AK as an inset map}
# Download the US states shapefile from the Census Bureau
url_us_states <- "https://www2.census.gov/geo/tiger/GENZ2021/shp/cb_2021_us_state_20m.zip"
download.file(url_us_states, "inputs/map_shape_files/us_states.zip", mode = "wb")
unzip("inputs/map_shape_files/us_states.zip", exdir = "inputs/map_shape_files/us_states")


# Read the US states shapefile
us_states <- read_sf("inputs/map_shape_files/us_states/cb_2021_us_state_20m.shp") %>% 
  st_transform(common_crs) %>% 
  filter(NAME != 'Hawaii' & NAME != 'Puerto Rico') %>%
  st_crop(xmin = -180, xmax = -55, ymin = 12, ymax = 62)

ggplot() +
  geom_sf(data = us_states, fill = "white", color = "black")


# Canadian provinces
canadian_provinces <- sf::st_as_sf(rnaturalearth::ne_states(country = "Canada", returnclass = "sf")) %>% 
  st_transform(common_crs) 

ggplot() +
  geom_sf(data = canadian_provinces, fill = "white", color = "black")


# Bind US and canadian provinces # 
us_ca <- bind_rows(us_states, canadian_provinces)

ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black")

ggplot() +
  geom_sf(data = us_ca, fill = "white", color = "black") +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn, color = Fire_name)) +
  theme_void() +
  theme(legend.position = "none")

```

```{r koppen climate classification layer}
# Peter code to plot tif #

## Read in koppen climate classification - only needs to be read in once when plotting and then can be deleted after the kppen_reproject
koppen_raw <-  raster::raster("~/Downloads/1984_2013/kc_1984_2013.tif")

## Re-project koppen map
koppen_reproject <- raster::projectRaster(koppen_raw, crs = 4269)

## Mask (crop in essence) severity to fire boundary
koppen_crop <- raster::mask(koppen_reproject, us_ca)

## Convert to dataframe
koppen <- as.data.frame(koppen_crop, xy = T) %>% 
  as_tibble() %>% 
  rename("long" = x, 
         "lat" = y, 
         "koppen" = kc_1984_2013) %>% 
  filter(!is.na(koppen)) %>% 
  mutate(f_koppen = as.numeric(round(koppen, 0)),
         f2_koppen = case_when(f_koppen < 200 ~ "Tropical",
                               f_koppen > 200 & f_koppen < 300 ~ "Arid",
                               f_koppen > 300 & f_koppen < 400 ~ "Temperate",
                               f_koppen > 400 & f_koppen < 500 ~ "Boreal",
                               f_koppen > 500 ~ "Polar"))
koppen <- koppen %>% 
  na.omit(f2_koppen)

# Plot raster 
ggplot() + 
  geom_raster(data = koppen, 
              aes(long, lat, fill = f2_koppen), alpha = 0.85) +
  geom_sf(data = coordinates_sf, aes(size = number_of_burn), color = "black") +
  labs(fill = "Biome") +
  labs(size = "Number of Burned Watersheds") +
  ggtitle("Koppen-Geiger Classification") +
  ggspatial::annotation_scale(
    location = "br",
    pad_x = unit(0.5, "in"), 
    bar_cols = c("black", "white")) +
  ggspatial::annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(1.1, "in"), pad_y = unit(0.5, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("black", "white"),
      line_col = "grey20")) +
  theme_minimal() +
  theme(legend.position = "left", 
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
 
  
#
ggsave("koppen_map_burn_label.pdf",
       path = "~/Documents/RC_3/Wenas_Watershed_project/Manuscript",
       width = 10, height = 8,  units = "in") 


```




