---
title: "11_Random_forest"
output: html_document
date: "2024-02-07"
editor_options: 
  chunk_output_type: console
---

The purpose of this script is to run random forests for DOC and NO3 effect sizes using catchment characteristics

Script Workflow:

Step 1) Load in the geospatial csv that is generated from pulling in stream cat data

Step 2) Load in the fire name and year summary csv

Step 3) Load in the effect size sheet that comes from the All_Studies_Temporal_Normalization script that includes effect size for each watershed. 




# Status: in progress

# ==============================================================================
# Author: Peter Regier

# Adapted by Jake Cavaiani; jake.cavaiani@pnnl.gov

# 7 February 2024
# ==============================================================================

# Library download # 
```{r Load packages}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

require(pacman)
library(tidyverse)
library(janitor)
library(tidymodels)
library(PNWColors)
library(ranger)
library(here)

```

# random forest dataframes # 
```{r read in csvs}
read_data <- function(filename){
  read_csv(here("Output_for_analysis", "08_Meta_summary_stats", filename), na = c('-9999', 'N/A')) %>% 
  clean_names() %>%
  rename("dependent" = effect_size) %>%
  dplyr::select(dependent, climate, totdasqkm, maxelevsmo, slope, tmean9120ws, om_ws, bfi_ws, perm_ws, rck_dep_ws,
                burn_percentage, mean_dnbr, time_since_fire) %>%
  mutate(mean_dnbr = as.numeric(mean_dnbr),
         white_noise = rnorm(1:n(), mean = 0, sd = 1)) %>%
  na.omit()
}

## Read in normalized effect datasets
no3 <- read_data("no3_meta_random_forest.csv")
doc <- read_data("doc_meta_random_forest.csv")
  
## Add non-normalized data
no3_nn <- read_data("no3_meta_random_forest_nonnormal.csv")
doc_nn<- read_data("doc_meta_random_forest_nonnormal.csv")

## Set ggplot theme
theme_set(theme_bw())

```

# 2. Random Forest models 
```{r function to calculate variable importance}
## Function to calculate variable importance for a default ranger() model
plot_fi <- function(data){
  
  model_recipe <- data %>% 
    recipe(dependent ~ .) %>% 
    step_integer(climate) %>% 
    # step_corr(all_predictors()) %>% # looking at correlations between the variables and if two things are strongly correlated. Runoff is strongly correlated with another variable within DOC. Random forest do a good job handling multi-colinearity
    step_normalize(all_predictors(), -all_outcomes()) %>% 
    recipes::prep()
  
  df <- model_recipe %>% 
    bake(data)
  
  ## Make model
  rf_model <- ranger(dependent ~ ., data = df, importance = "impurity")
  
  ## Visualize model output
  print(rf_model)
  
  ## Set vectors for renaming stuff
  var_names <- rf_model$variable.importance
  col_names <- c("predictor", "raw_fi")
  
  ## Convert feature importance to a tibble with variables as a column
  fi0 <- as.data.frame(var_names) %>% 
    tibble::rownames_to_column() %>% 
    as_tibble()
  
  ## Rename columns
  colnames(fi0) = col_names
  
  ## Output variable importance (or feature importance)
  fi0
}


```


# Plot FI for effect sizes
```{r plot Feature importance}
## Create a dataset of variable importance
# Old code with no white noise included. 
# df <- bind_rows(plot_fi(no3) %>% mutate(analyte = "no3"), 
#           plot_fi(doc) %>% mutate(analyte = "doc")) %>% 
#   group_by(analyte) %>% 
#   mutate(fi = raw_fi / sum(raw_fi),
#          fi100 = fi * 100)
# 
# write_csv(df, here("Output_for_analysis", "11_Random_forest", "Random_Forest_Effect_Size_Fig_data.csv"))

replicate_plot_fi <- function(parameter, analyte, nn){
  
  #set.seed(100)
  
  replicate(200, plot_fi(parameter), simplify = FALSE) %>% 
    bind_rows() %>% 
    group_by(predictor) %>% 
    summarize(mean_fi = mean(raw_fi, na.rm = T), 
              min_fi = min(raw_fi, na.rm = T), 
              max_fi = max(raw_fi, na.rm = T)) %>% 
    ungroup() %>%
    mutate(analyte = analyte) %>% 
    mutate(normalization = nn)
}


# JSC Testing here 
# @PETER this is what i've been messing with
doc_eb <- replicate_plot_fi(doc, "doc", "Normalized")
no3_eb <- replicate_plot_fi(no3, "no3", "Normalized")
doc_eb_nn <- replicate_plot_fi(doc_nn, "doc", "Not normalized")
no3_eb_nn <- replicate_plot_fi(no3_nn, "no3", "Not normalized")

# Mutate each column at a time
df <- bind_rows(doc_eb, no3_eb, doc_eb_nn, no3_eb_nn)

df_test <- df %>% 
  group_by(analyte) %>% 
  mutate(across(contains("_fi"), ~ . / sum(mean_fi, na.rm = TRUE)))
```


```{r white noise}
# BACK TO NORMAL WORKFLOW
# pr <- bind_rows(replicate_plot_fi(no3, "no3"),
#                     replicate_plot_fi(doc, "doc")) %>%  
#   group_by(analyte) %>%
#   mutate(across(contains("_fi"), ~ . / sum(.)))

## Set up a color palette
var_colors <- PNWColors::pnw_palette("Bay", n = length(unique(df$predictor)))

## Make plots
# burn_percentage - Burn percentage 
# Climate - Climate classification 
# maxelevsmo - Maximum elevation (smoothed) in centimeters
# minelevsmo - Minimum elevation (smoothed) in centimeters
# slope - Slope of flowline (meters/meters) based on smoothed elevations
# time_since_fire - time since fire 
# Mean Bedrock depth - RckDep
# totdasqkm -Total upstream catchment area from downstream end of flowline.


dose.labs <- c("DOC", "Nitrate")
names(dose.labs) <- c("doc", "no3")

ggplot(df_test, aes(mean_fi * 100, 
               reorder(predictor, mean_fi), fill = predictor)) + 
  geom_col(alpha = 0.8, show.legend = F, width = 0.7) +
  geom_errorbar(aes(xmin = min_fi * 100, xmax = max_fi * 100, width = 0.2)) +
  #facet_wrap(~analyte, nrow = 1, labeller = labeller(analyte = dose.labs)) + 
  facet_grid(normalization~analyte, labeller = labeller(analyte = dose.labs)) + 
  scale_fill_manual(values = var_colors) + 
  labs(x = "Feature Importance (%)", 
       y = "", fill = "") +
  scale_y_discrete(labels=c("bfi_ws" = "Baseflow Index",
                            "runoff_ws" = "Runoff",
                            "perm_ws" = "Soil Permeability",
                            "tmean9120ws" = "Mean Annual Air Temperature",
                            "mean_dnbr" = "Burn Severity",
                            "om_ws" = "Organic Matter",
                            "maxelevsmo" = "Max Watershed Elevation",
                            "burn_percentage" = "Burn Percentage", 
                            "climate" = "Koppen Climate Classification",
                            "rck_dep_ws" = "Mean Bedrock Depth",
                            "white_noise" = "White Noise",
                            "slope" = "Catchment Slope",
                            "time_since_fire" = "Time Since Fire",
                            "totdasqkm" = "Catchment Area")) +
  theme_bw()
                            

ggsave("Effect_Size_Random_Forest_white_noise.pdf",
       path = here("initial_plots", "11_Meta_random_forest"),
       width = 5.5, height = 6,  units = "in")

# including white noise. You put in a random variable. You feed that in as a predictor. If white noise has higher feature importance than any of these variables. The concern there is, we dont want to add anything that doesn't address the comment. 

# white noise is greater than climate, tmean, rck_dep, burn percentage, perm, om, watershed area, bfi for doc
  # white noise is greater than rck_dep, om, bp, perm, watershed area, bfi

# white noise is greater than slope and watershed area for no3


# Plot only with variables that are greater than white noise.  
# Calculate the white noise mean_fi for each analyte
white_noise_values <- df_test %>% 
  filter(predictor == "white_noise") %>% 
  select(analyte, normalization, mean_fi)

# Filter the data to include only variables with mean_fi greater than white noise value for each analyte
filtered_df <- df_test %>%
  inner_join(white_noise_values, by = c("analyte", "normalization"), suffix = c("", "_white_noise")) %>%
  filter(mean_fi > mean_fi_white_noise)
```



```{r make plots}
# Plotting
# CHAT 1.0 
# Define the blank spaces
# blank_spaces <- data.frame(
#   analyte = c(rep("doc", 5), "no3"),
#   predictor = c("perm_ws", "tmean9120ws", "om_ws", 
#                 "bfi_ws", "burn_percentage", "slope")
# )

# Filtered dataset assumed to be 'filtered_df'
# Original plot with X annotations
p <- ggplot(filtered_df, aes(mean_fi * 100, reorder(predictor, mean_fi), fill = predictor)) +
  geom_col(alpha = 0.8, show.legend = F, width = 0.7) +
  geom_errorbar(aes(xmin = min_fi * 100, xmax = max_fi * 100, width = 0.2)) +
  facet_grid(normalization~analyte, labeller = labeller(analyte = dose.labs)) + 
  scale_fill_manual(values = var_colors) +
  labs(x = "Feature Importance (%)", y = "", fill = "") +
  scale_y_discrete(labels = c(
    "bfi_ws" = "Baseflow Index",
    "runoff_ws" = "Runoff",
    "perm_ws" = "Soil Permeability",
    "tmean9120ws" = "Mean Annual Air Temperature",
    "mean_dnbr" = "Burn Severity",
    "om_ws" = "Organic Matter",
    "maxelevsmo" = "Max Watershed Elevation",
    "burn_percentage" = "Burn Percentage", 
    "climate" = "Koppen Climate Classification",
    "rck_dep_ws" = "Mean Bedrock Depth",
    "white_noise" = "White Noise",
    "slope" = "Catchment Slope",
    "time_since_fire" = "Time Since Fire",
    "totdasqkm" = "Catchment Area"
  )) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))

blank_spaces <- df_test %>%
  inner_join(white_noise_values, by = c("analyte", "normalization"), suffix = c("", "_white_noise")) %>%
  filter(mean_fi <= mean_fi_white_noise) %>% 
  filter(predictor != "white_noise")


# Manually create a data frame for the red X's for missing predictors in the "doc" panel
# blank_spaces <- data.frame(
#   analyte = c(rep("doc", 5), "no3"),
#   predictor = c("perm_ws", "tmean9120ws", "om_ws", 
#                 "bfi_ws", "burn_percentage", "slope")
# )
# missing_doc <- data.frame(
#   analyte = c(rep("doc", 5), "no3"),
#   predictor = c("perm_ws", "tmean9120ws", "bfi_ws", "om_ws", "burn_percentage", "slope"),
#   mean_fi = c(0, 0, 0, 0, 0, 0)  # Set mean_fi to 0 to position the X's
# )

# A single frame for the y-values
# predictor_y_mapping <- data.frame(
#   predictor = c("perm_ws", "tmean9120ws", "bfi_ws", "om_ws", "burn_percentage", "slope"),
#   y_position = c(11, 10, 9, 6, 2, 8) # Adjust this multiplyer to move X upwards or downwards
# )

# Add y_position from predictor_y_mapping to missing_doc
# missing_doc <- merge(missing_doc, predictor_y_mapping, by="predictor")

# Add red X for cases where predictors are specified as blank for "doc"
p + 
  geom_text(data = blank_spaces, aes(x = mean_fi + 5, label = "X"),
                   color = "red", size = 5, vjust = 0.5)


ggsave("Effect_Size_Random_Forest_variables_less_than_white_noise.pdf",
       path = here("initial_plots", "11_Meta_random_forest"),
       width = 5.5, height = 3,  units = "in")

```






