---
title: "12_Meta_Clean_MTBS_files"
output: html_document
date: "2024-06-13"
---

#' Clean MTBS wildfire files 
#' 
#' MTBS downloads come in multiple zipped folders and the file names are long and meaningless. This will pull all files in a single folder and rename files to be more user friendly. 

The purpose of this script is to clean MTBS wildfire file names. The MTBS downloads come in multiple zipped folders and the file names are long and meaningless. This will pull all files in a single folder and rename files to be more user friendly. 

# Status: in progress

# ==============================================================================
# Author: Katie Wampler
#     Adapted by Jake Cavaiani; jake.cavaiani@pnnl.gov
# 13 June 2024
# ==============================================================================

```{r load libraries}
library(here)

```
### Functions
```{r}
# Function 1
str_split2 <- function(string, pattern, n=Inf, simplify=F){
  val <- str_split(string, pattern=pattern, n=n, simplify = simplify)
  if(length(val) == 1){
    val <- unlist(val)
  }
  val
}


# Define the function to clean MTBS files
clean_mtbs <- function(wd){
  # Get files of fires 
  files <- list.files(wd, recursive = TRUE)  
  # files <- files[stringr::str_detect(files, ".zip")] # Uncomment if you have zipped files
  
  lapply(1:length(files), function(f){
    
    x <- files[f]
    
    # Get place to put zip files
    exdir <- paste(wd, unlist(stringr::str_split(x, "/"))[1], sep="/")
    
    # Unzip file
    unzip(paste(wd, x, sep="/"), exdir=exdir)
    
    # Find shape file to extract fire name
    zip_files <- list.files(exdir)
    shp_file <- zip_files[stringr::str_detect(zip_files, "burn_bndy.shp")]
    
    # Read file to get fire name 
    name <- sf::st_read(paste(exdir, shp_file, sep="/"))
    filename <- name$Incid_Name
    
    # Format name for file 
    filename <- tolower(filename)
    filename <- stringr::str_replace_all(filename, " ", "_")
    filename <- gsub("[][()]", "", filename) 
    
    # Rename files 
    # Get dates of data 
    dates <- str_split2(shp_file, "_")[2:3]
    
    # Get header, remove files from other fires 
    header <- str_split2(shp_file, "_")[1]
    zip_files <- zip_files[str_detect(zip_files, regex(header, ignore_case=TRUE))]
    
    # If both, remove completely 
    new_files <- gsub(paste("_", paste(dates, collapse ="_"), sep=""), "", zip_files)
    
    # If just one use pre or post 
    new_files <- gsub(dates[1], "pre", new_files)
    new_files <- gsub(dates[2], "post", new_files)
    
    # Replace long name with fire name 
    new_files <- gsub(header, filename, new_files)
    
    # Replace metadata name 
    new_files <- gsub(toupper(header), filename, new_files)
    
    # Copy files with new names instead of renaming
    file.copy(from = paste(exdir, zip_files, sep="/"), 
              to = paste(exdir, new_files, sep="/"))
  })
}
```

# Loop through all the fires we have for the analysis
```{r}
# Define the base directory where the folders are located
base_dir <- here("gis_data", "Fire_Perimeters")

folder_names <- c("angora", "caldor", "clovermist", "fourmile_canyon", "gaviota", "hayman", "high_park", "mosquito", "pole_creek", "rampage_complex_double_mountain_2","red_bench", "wragg") # 13 fires


# Define a named vector or list of folder names and their corresponding years
folders_and_years <- list(
  "angora" = "2007",
  "caldor" = "2021",
  "clovermist" = "1988",
  "fourmile_canyon" = "2010",
  "gaviota" = "2004",
  "hayman" = "2002",
  "hewlett" = "2012",
  "high_park" = "2012",
  "mosquito" = "2022",
  "pole_creek" = "2018",
  "rampage_complex" = "2003",
  "red_bench" = "1988",
  "wragg" = "2015")

# folder_name <- "Clovermist_1988"
# folder_name <- "Fourmile_Canyon_2010"
# folder_name <- "Gaviota_2004"
# folder_name <- "Hayman_2002"
# folder_name <- "High_Park_2012"
# folder_name <- "Mosquito_2022"
# folder_name <- "Pole_Creek_2018"
# folder_name <- "Rampage_Complex_2003"
# folder_name <- "Red_Bench_1988"


# Iterate over each folder and process the files using a for loop
for (folder_name in names(folders_and_years)) {
  year <- folders_and_years[[folder_name]]
  wd <- here("gis_data", "Fire_Perimeters", folder_name, "mtbs", year)
  clean_mtbs(wd)
}

```



# ============================ Hewlett Gulch ============================================ # 
```{r}
# Define the function to clean MTBS files
clean_mtbs <- function(wd){
  # Get files of fires 
  files <- list.files(wd, recursive = TRUE)  
  # files <- files[stringr::str_detect(files, ".zip")] # Uncomment if you have zipped files

  lapply(1:length(files), function(f){
    
    x <- files[f]
  
    # Get place to put zip files
    exdir <- paste(wd, unlist(stringr::str_split(x, "/"))[1], sep="/")
    
    # Unzip file
    unzip(paste(wd, x, sep="/"), exdir=exdir)
    
    # Find shape file to extract fire name
    zip_files <- list.files(exdir)
    shp_file <- zip_files[stringr::str_detect(zip_files, "burn_bndy_alb.shp")]
    
    # Read file to get fire name 
    name <- sf::st_read(paste(exdir, shp_file, sep="/"))
    filename <- name$Incid_Name
    
    # Format name for file 
    filename <- tolower(filename)
    filename <- stringr::str_replace_all(filename, " ", "_")
    filename <- gsub("[][()]", "", filename) 
    
    # Rename files 
    # Get dates of data 
    dates <- str_split2(shp_file, "_")[2:3]
    
    # Get header, remove files from other fires 
    header <- str_split2(shp_file, "_")[1]
    zip_files <- zip_files[str_detect(zip_files, regex(header, ignore_case=TRUE))]
    
    # If both, remove completely 
    new_files <- gsub(paste("_", paste(dates, collapse ="_"), sep=""), "", zip_files)
    
    # If just one use pre or post 
    new_files <- gsub(dates[1], "pre", new_files)
    new_files <- gsub(dates[2], "post", new_files)
    
    # Replace long name with fire name 
    new_files <- gsub(header, filename, new_files)
    
    # Replace metadata name 
    new_files <- gsub(toupper(header), filename, new_files)
    
    # Copy files with new names instead of renaming
    file.copy(from = paste(exdir, zip_files, sep="/"), 
              to = paste(exdir, new_files, sep="/"))
  })
}
```

```{r Hewlett Gulch}
# Define the base directory where the folders are located
base_dir <- here("gis_data", "Fire_Perimeters")

folder_names <- c("hewlett") # 1 fire

# Define a named vector or list of folder names and their corresponding years
folders_and_years <- list(
  "hewlett" = "2012")

# Iterate over each folder and process the files using a for loop
for (folder_name in names(folders_and_years)) {
  year <- folders_and_years[[folder_name]]
  wd <- here("gis_data", "Fire_Perimeters", folder_name, "mtbs", year)
  clean_mtbs(wd)
}

```



