---
title: "Temporal_normalization_vs_nonnormal"
output: html_document
date: "2026-02-12"
editor_options: 
  chunk_output_type: console
---

## Read in libraries 
```{r Jake/Mac Load Packages and}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

library(pacman)
p_load(tidyverse,
       here,
       forecastML, 
       zoo,
       # hrbrthemes,
       viridis, 
       ggpmisc,
       RobustLinearReg,
       lme4,
       car)

```

### Build Model data frame for all variables of interest - BP, BS, TSF, CLIMATE 
```{r Build model data frames}
# Read in effect size data frame for normalized and non-normalized data
ES_nn_normal <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_size_orig_nonnormal.csv"))

ES_nn_normal <- ES_nn_normal %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      Study_ID == "Abbott et al. 2021" & occurrence == 3 ~ "10", # Third value
      Study_ID == "Abbott et al. 2021" & occurrence == 4 ~ "11", # Fourth value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire)) %>% 
  dplyr::select(Study_ID:Effect_size_no_watershed)

# Read in the study and site csv 
study_site_data <- read_csv(here("inputs", "catchment_characteristics", "Fire_name_Lat_Long.csv"), na = c('-9999', 'N/A')) %>% 
  dplyr::select(Study_ID, Site, Fire_year, Effect_size_pair, burn_percentage) %>% 
  rename(year = Fire_year) %>% 
  filter(!Study_ID == "Murphy et al. 2015") %>% 
  na.omit()

unique(study_site_data$Study_ID) # Should be 13 studies because we are missing Abbott.
# Abbott because of merging burned and unburned sites. 
# Murphy et al. 2015 was already accounted for in the Murphy et al. 2018 


### BURN PERCENTAGE MODEL DATAFRAME ####
bp_clip <- study_site_data %>% 
  dplyr::select(Study_ID, Site, year, Effect_size_pair, burn_percentage)

BP_ES <- full_join(bp_clip, ES_nn_normal) %>% 
  na.omit() %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Effect_size_no_watershed, burn_percentage) # joining the BP dataframe used to make the figure and study site data to match sites because we will be including site as a random effect

sort(unique(BP_ES$Study_ID)) # 13 studies

# Merge to include site
BP_model <- full_join(BP_ES, study_site_data) %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Site, Effect_size, Effect_size_no_watershed, burn_percentage)

# filter DOC and no3 out 
doc_es_bp <- BP_model %>% 
  filter(response_var == "DOC_Interp")

no3_es_bp <- BP_model %>% 
  filter(response_var == "NO3_Interp")

#### DNBR BURN SEVERITY MODEL DATAFRAME ##### 
DNBR_df <- read_csv(here("Output_for_analysis", "14_Meta_calculate_burn_severity", "Burn_severity_ES_fig_data.csv")) %>% 
  dplyr::select(Study_ID, response_var, fire_name, Effect_size_pair, mean_dnbr, burn_severity_category, burn_percentage, Time_Since_Fire)

DNBR_model <- left_join(DNBR_df, ES_nn_normal, by = c("Study_ID", "Effect_size_pair", "response_var", "Time_Since_Fire")) %>% 
  na.omit()

# Filter by response 
doc_es_dnbr <- DNBR_model %>% 
  filter(response_var == "DOC_Interp")

no3_es_dnbr <- DNBR_model %>% 
  filter(response_var == "NO3_Interp")

#### TSF MODEL DATAFRAME ##### 
doc_es_tsf <- ES_nn_normal %>% 
  filter(response_var == "DOC_Interp") %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Effect_size_no_watershed, 
                Time_Since_Fire)

no3_es_tsf <- ES_nn_normal %>% 
  filter(response_var == "NO3_Interp") %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Effect_size_no_watershed, 
                Time_Since_Fire)

#### CLIMATE MODEL DATAFRAME ##### 
# Filter out studies that are greater than 10 years 
doc_es_climate <- ES_nn_normal %>% 
  filter(response_var == "DOC_Interp") %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, Effect_size_pair, Effect_size, Effect_size_no_watershed) %>% 
  filter(Study_ID != "Abbott et al. 2021") %>% 
  filter(Study_ID != "Rhea et al. 2021")

no3_es_climate <- ES_nn_normal %>% 
  filter(response_var == "NO3_Interp") %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, Effect_size_pair, Effect_size, Effect_size_no_watershed) %>% 
  filter(Study_ID != "Abbott et al. 2021") %>% 
  filter(Study_ID != "Rhea et al. 2021")

```

# TSF NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r TSF - nn vs. normal anova}
### DOC ####
# Pivot into long format
long_doc_nn_normal <- doc_es_tsf %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_doc_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_doc_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We conducted a one-way ANOVA to compare dissolved organic carbon (DOC) effect sizes calculated with and without watershed-level controls. The analysis included 88 observations and revealed a statistically significant difference between the two effect size calculations (F₁,₈₆ = 21.23, p < 0.001).

# The effect type (with vs. without watershed controls) explained a substantial portion of the variance in effect sizes, with an effect size sum of squares of 11.15 compared to residual sum of squares of 45.17. This significant difference indicates that the inclusion or exclusion of watershed-level controls meaningfully impacts the calculated effect sizes for fire impacts on DOC concentrations.

### Nitrate ####
# Pivot into long format
long_no3_nn_normal <- no3_es_tsf %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_no3_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_no3_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We examined whether the inclusion of watershed-level covariates significantly altered nitrate (NO3_Interp) effect size estimates using a one-way ANOVA comparing effect sizes calculated with and without watershed variables. The analysis included 170 observations (df = 168 for residuals, plus 1 df for the effect type factor).

# The ANOVA revealed no statistically significant difference between effect sizes calculated with watershed covariates versus those calculated without watershed covariates (F₁,₁₆₈ = 2.053, p = 0.154). This suggests that the inclusion of watershed-level variables does not substantially alter the magnitude of nitrate effect size estimates in our models.

# The residual mean square error was 1.701, indicating the typical within-group variability in effect sizes after accounting for the effect type classification.

```

# TSF MODEL 
```{r TSF - Normalized & non-normal models}
### TSF DOC normalized  ####
# Original workflow 
# data frame: doc_es_tsf 
doc_es_tsf <- doc_es_tsf %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_tsf_mixed_normal <-  lmer(Effect_size ~ Time_Since_Fire + (1 | Site), data = doc_es_tsf)


# Check for equal variance
performance::check_heteroscedasticity(doc_tsf_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.993).

# Check for normality
residuals <- resid(doc_tsf_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96893, p-value = 0.2772
# when p > 0.05 can assume normality 

doc_tsf_mixed_normal

summary(doc_tsf_mixed_normal) 

# P-value 
Anova(doc_tsf_mixed_normal, type = "III")

# We analyzed the relationship between time since fire and dissolved organic carbon (DOC) effect sizes using a linear mixed-effects model with site as a random effect to account for multiple observations from the same study locations. The model included 44 observations from 21 sites.

# The mixed-effects model revealed a marginally significant negative relationship between time since fire and DOC effect size (β = -0.034, SE = 0.018, t = -1.876, p = 0.071). This suggests that the magnitude of fire effects on DOC may decrease over time, with effect sizes becoming less negative (closer to zero) as time progresses following fire disturbance.

# The model intercept was -0.321 (SE = 0.246, t = -1.305, p = 0.205), indicating that at time zero (immediately post-fire), the predicted effect size was negative but not significantly different from zero. The random effect of site accounted for substantial variation in the data (variance = 1.165, SD = 1.079), while residual variance was relatively low (variance = 0.016, SD = 0.125), suggesting that site-level differences were an important source of variation in fire effects on DOC.

# A Type III Wald chi-square test confirmed the marginal significance of the time since fire effect (χ² = 3.519, df = 1, p = 0.061), indicating a trend toward recovery of DOC levels over time following fire disturbance.


### TSF DOC non-normalized ####
# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_tsf_mixed_nn <-  lmer(Effect_size_no_watershed ~ Time_Since_Fire + (1 | Site), data = doc_es_tsf)


# Check for equal variance
performance::check_heteroscedasticity(doc_tsf_mixed_nn) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_tsf_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.97221, p-value = 0.3613
# when p > 0.05 can assume normality 

doc_tsf_mixed_nn

summary(doc_tsf_mixed_nn) 

# P-value 
Anova(doc_tsf_mixed_nn, type = "III")

# We analyzed the relationship between time since fire and dissolved organic carbon (DOC) effect sizes calculated without watershed controls using a linear mixed-effects model with site as a random effect. The model included 44 observations from 21 sites.

#The mixed-effects model revealed a highly significant negative relationship between time since fire and DOC effect size without watershed controls (β = -0.057, SE = 0.010, t = -5.454, p < 0.001). This indicates that fire effects on DOC decrease substantially over time, with effect sizes becoming smaller as time progresses following fire disturbance.

#The model intercept was significantly positive (β = 0.391, SE = 0.072, t = 5.402, p < 0.001), indicating that immediately post-fire (time zero), there was a strong positive effect of fire on DOC concentrations. The random effect of site showed moderate variation (variance = 0.069, SD = 0.263), while residual variance remained low (variance = 0.016, SD = 0.127), suggesting that while site-level differences contributed to variation, the temporal pattern was consistent across sites.

# Type III Wald chi-square tests confirmed the highly significant effects of both the intercept (χ² = 29.18, df = 1, p < 0.001) and time since fire (χ² = 29.74, df = 1, p < 0.001). These results demonstrate a clear pattern of DOC recovery over time, with initially elevated post-fire concentrations declining significantly as ecosystems recover from fire disturbance.

# Note: The strong significance in this model contrasts with the marginal effects found when watershed controls were included, suggesting that watershed-level factors may mask or confound the temporal recovery patterns in DOC concentrations following fire.

### TSF NITRATE normalized ####
# Original workflow 
# data frame: doc_es_tsf 
no3_es_tsf <- no3_es_tsf %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
no3_tsf_mixed_normal <-  lmer(Effect_size ~ Time_Since_Fire + (1 | Site), data = no3_es_tsf)

# Check for equal variance
performance::check_heteroscedasticity(no3_tsf_mixed_normal) # OOK: Error variance appears to be homoscedastic (p = 0.910).

# Check for normality
residuals <- resid(no3_tsf_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96477, p-value = 0.1067
# when p > 0.05 can assume normality 

no3_tsf_mixed_normal

summary(no3_tsf_mixed_normal) 

# P-value 
Anova(no3_tsf_mixed_normal, type = "III")

# We analyzed the relationship between nitrate effect size and time since fire using a linear mixed-effects model with site as a random effect. The model included 85 observations across 34 sites.

# Time since fire had a significant positive effect on nitrate effect size (β = 0.099, SE = 0.049, χ² = 4.05, p = 0.044). For each additional year since fire, the nitrate effect size increased by approximately 0.10 units. The intercept was not significantly different from zero (β = 0.283, SE = 0.300, χ² = 0.89, p = 0.347).

# The random effect of site accounted for substantial variation in the data, with a standard deviation of 1.44 for the site-level intercepts compared to a residual standard deviation of 0.73. This indicates considerable between-site heterogeneity in baseline nitrate effect sizes. The correlation between the intercept and time since fire slope was moderate and negative (r = -0.48).

# These results suggest that the impact of fire on nitrate dynamics becomes more pronounced with increasing time since the fire event, though there is substantial variation among sites in their baseline responses.

### TSF NITRATE non-normalized ####
# data frame: no3_es_tsf 

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
no3_tsf_mixed_non_normal <-  lmer(Effect_size_no_watershed ~ Time_Since_Fire + (1 | Site), data = no3_es_tsf)

# Check for equal variance
performance::check_heteroscedasticity(no3_tsf_mixed_non_normal) # OK: Error variance appears to be homoscedastic (p = 0.259).

# Check for normality
residuals <- resid(no3_tsf_mixed_non_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.93518, p-value = 0.0003474
# when p > 0.05 can assume normality 
# NON-NORMAL?

no3_tsf_mixed_non_normal

summary(no3_tsf_mixed_non_normal) 

# P-value 
Anova(no3_tsf_mixed_non_normal, type = "III")

# We analyzed the relationship between nitrate effect size (excluding watershed studies) and time since fire using a linear mixed-effects model with site as a random effect. The model included 85 observations across 34 sites.

# Time since fire had a significant positive effect on nitrate effect size (β = 0.096, SE = 0.034, χ² = 7.79, p = 0.005). For each additional year since fire, the nitrate effect size increased by approximately 0.10 units. The intercept was significantly greater than zero (β = 0.698, SE = 0.172, χ² = 16.38, p < 0.001), indicating a positive baseline effect of fire on nitrate dynamics.

# The random effect of site showed moderate variation, with a standard deviation of 0.64 for the site-level intercepts compared to a residual standard deviation of 0.78. This suggests modest between-site heterogeneity in baseline nitrate effect sizes. The correlation between the intercept and time since fire slope was moderate and negative (r = -0.56).

# These results demonstrate that fire has a consistently positive effect on nitrate dynamics that becomes increasingly pronounced over time since the fire event. The exclusion of watershed-scale studies appears to have strengthened the relationship compared to the full dataset, with both the baseline effect and the temporal trend being statistically significant.

# COMPARE
# We fitted two linear mixed-effects models to examine the relationship between nitrate effect size and time since fire: one using the complete dataset and another excluding watershed-scale studies. Both models included site as a random effect and analyzed 85 observations across 34 sites.

# Effect of Time Since Fire: Both models showed a significant positive relationship between time since fire and nitrate effect size, with nearly identical slopes (β = 0.099 vs. 0.096). However, the statistical significance was stronger for non-normalized effect size (p = 0.005 vs. p = 0.044), suggesting that watershed-scale studies may introduce additional variability that weakens the temporal signal.

```

### TSF PLOT NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r TSF plot}
tsf_nn_normal_long <-  doc_es_tsf %>%
  pivot_longer(
    cols= c(Effect_size, Effect_size_no_watershed),
    names_to = "normalization",
    values_to = "Effect_size"
  )

vn = expression(paste(""*N*O[3]^"-"))
dose.labs <- as_labeller(c("DOC" = "DOC", "NO3" = "Nitrate"))

# Means and SD
tsf_nn_normal_plot = tsf_nn_normal_long %>%
  mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
   group_by(Time_Since_Fire, response_var, normalization) %>%
  summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>%
  mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect)
View(tsf_nn_normal_plot)

tsf_plot = tsf_nn_normal_plot %>% 
  ggplot(aes(Time_Since_Fire, Effect_size, color = normalization))+
  geom_pointrange(aes(ymin = mean_effect - se_effect, ymax = mean_effect + se_effect,
                      color = normalization)) +
  facet_grid(~response_var, labeller = dose.labs) +
     scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Normalization"),
                     labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
  xlab("Time Since Fire") +
  ylab("Effect Size") +
    geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
      theme_bw()

tsf_plot

```


# CLIMATE MODEL 
```{r Climate - nn vs. normal anova}
### DOC ####
# Pivot into long format
long_doc_nn_normal <- doc_es_climate %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_doc_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_doc_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We conducted a one-way ANOVA to compare dissolved organic carbon (DOC) effect sizes calculated with and without watershed-level controls. The analysis included 76 observations and revealed a statistically significant difference between the two effect size calculations (F₁,74 = 26.5, p < 0.001).

# The residual mean square error was 0.531, indicating relatively low within-group variability compared to the between-group differences in effect sizes.

### Nitrate ####
# Pivot into long format
long_no3_nn_normal <- no3_es_climate %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_no3_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_no3_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We examined whether the inclusion of watershed-level covariates significantly altered nitrate (NO3_Interp) effect size estimates using a one-way ANOVA comparing effect sizes calculated with and without watershed variables. The analysis included 158 observations (df = 156 for residuals, plus 1 df for the effect type factor).

# The ANOVA revealed no statistically significant difference between effect sizes calculated with watershed covariates versus those calculated without watershed covariates (F₁,₁₅₆ = 2.561, p = 0.112). This suggests that the inclusion of watershed-level variables does not substantially alter the magnitude of nitrate effect size estimates in our models.

# The residual mean square error was 1.640, indicating the typical within-group variability in effect sizes after accounting for the effect type classification.
```

```{r CLIMATE - Normalized & non-normal models}
### CLIMATE DOC normalized  ####
# Original workflow 
# data frame: doc_es_climate
doc_es_climate <- doc_es_climate %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_climate_mixed_normal <-  lmer(Effect_size ~ Climate + (1 | Site), data = doc_es_climate)

# Check for equal variance
performance::check_heteroscedasticity(doc_climate_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.994)

# Check for normality
residuals <- resid(doc_climate_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96704, p-value = 0.237
# when p > 0.05 can assume normality 

doc_climate_mixed_normal

summary(doc_climate_mixed_normal) 

# P-value 
Anova(doc_climate_mixed_normal, type = "III")

# We analyzed the relationship between dissolved organic carbon (DOC) effect size and climate type using a linear mixed-effects model with site as a random effect. The model included 38 observations across 17 sites and five climate categories.

# Climate type did not have a significant effect on DOC effect size (χ² = 0.77, df = 4, p = 0.943). Compared to the reference climate category, none of the individual climate types showed significant differences: Csa (β = -0.226, SE = 0.906), Csb (β = 0.685, SE = 1.170), Dfb (β = 0.086, SE = 0.979), or Dfc (β = 0.012, SE = 1.170). The intercept was also not significantly different from zero (β = -0.518, SE = 0.739, χ² = 0.49, p = 0.483).

# The random effect of site accounted for the vast majority of variation in the data, with a standard deviation of 1.28 for site-level intercepts compared to a very small residual standard deviation of 0.13. This indicates that between-site variation dominates over within-site variation, and that site-specific factors are much more important than climate type in determining DOC response to fire.

# These results suggest that climate type, as categorized in this analysis, is not a significant predictor of how fire affects DOC dynamics. Instead, site-specific characteristics appear to be the primary driver of variation in fire effects on dissolved organic carbon.

### CLIMATE DOC non-normalized ####
# data frame: doc_es_climate
# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_climate_mixed_nn <-  lmer(Effect_size_no_watershed ~ Climate + (1 | Site), data = doc_es_climate)


# Check for equal variance
performance::check_heteroscedasticity(doc_climate_mixed_nn) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_climate_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.95387, p-value = 0.1195
# when p > 0.05 can assume normality 

doc_climate_mixed_nn

summary(doc_climate_mixed_nn) 

# P-value 
Anova(doc_climate_mixed_nn, type = "III")

# We analyzed the relationship between dissolved organic carbon (DOC) effect size and climate type using a linear mixed-effects model with site as a random effect, excluding watershed-scale studies. The model included 38 observations across 17 sites and five climate categories.

# Climate type had a significant effect on DOC effect size (χ² = 11.08, df = 4, p = 0.026). Compared to the reference climate category, the Csb climate showed the strongest positive effect (β = 0.638, SE = 0.218), followed by Csa (β = 0.278, SE = 0.166), Dfc (β = 0.196, SE = 0.214), and Dfb (β = 0.033, SE = 0.179). The intercept was not significantly different from zero (β = 0.117, SE = 0.130, χ² = 0.80, p = 0.371).

# The random effect structure differed markedly from the full dataset model. Site-level variation was substantially reduced (SD = 0.22) while residual variation remained similar (SD = 0.14), resulting in improved model fit (REML criterion: 30.97 to -8.49). This indicates that excluding watershed studies reduced between-site heterogeneity and allowed climate effects to emerge more clearly.

# These results suggest that climate type is a significant predictor of how fire affects DOC dynamics when watershed-scale studies are excluded. Mediterranean climates (Csa, Csb) appear to show particularly strong positive responses, indicating that climate-driven processes may be more detectable at smaller spatial scales where watershed-level heterogeneity is minimized.

# post hoc test 
a <- emmeans::lsmeans(doc_climate_mixed_nn, pairwise ~ Climate, adjust = "tukey") # This is assessing the pair wise comparisons for the rest of the climate classifications. 

# Or if you want the full contrast results as a data frame
contrast_results <- as.data.frame(a$contrasts)
# no climate types significantly differ from each other:

# We analyzed the relationship between dissolved organic carbon (DOC) effect size and climate type using a linear mixed-effects model with site as a random effect, excluding watershed-scale studies. The model included 38 observations across 17 sites and five climate categories.

# Climate type had a significant overall effect on DOC effect size (χ² = 11.08, df = 4, p = 0.026). However, post-hoc pairwise comparisons using Tukey's adjustment revealed no significant differences between any specific climate pairs (p > 0.05 for all comparisons). The intercept was not significantly different from zero (β = 0.117, SE = 0.130, χ² = 0.80, p = 0.371).

# The random effect structure showed that site-level variation was relatively low (SD = 0.22) compared to residual variation (SD = 0.14), with substantially improved model fit compared to the full dataset (REML criterion: -8.49).

# Interpretation
# The significant overall climate effect with non-significant pairwise comparisons suggests that while climate categories collectively explain significant variation in DOC response to fire, the differences between individual climate types are not sufficiently large to reach statistical significance after multiple comparison correction. This pattern often occurs when there is meaningful biological variation across groups but limited statistical power due to sample size or high within-group variability.

# This indicates that climate influences DOC fire responses, but the specific climate categories used may not capture the most relevant climatic distinctions for DOC dynamics.


### CLIMATE NITRATE normalized ####
# Original workflow 
# data frame: no3_es_climate
no3_es_climate <- no3_es_climate %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
no3_climate_mixed_normal <-  lmer(Effect_size ~ Climate + (1 | Site), data = no3_es_climate)

# Check for equal variance
performance::check_heteroscedasticity(no3_climate_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.963).

# Check for normality
residuals <- resid(no3_climate_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.95187, p-value = 0.004746
# when p > 0.05 can assume normality 

no3_climate_mixed_normal

summary(no3_climate_mixed_normal) 

# P-value 
Anova(no3_climate_mixed_normal, type = "III")

# We analyzed the relationship between nitrate effect size and climate type using a linear mixed-effects model with site as a random effect. The model included 79 observations across 30 sites and eight climate categories.

# Climate type had a highly significant effect on nitrate effect size (χ² = 30.09, df = 7, p < 0.001). Compared to the reference climate category, the climate effects varied substantially: Dsb showed the largest positive effect (β = 2.202, SE = 1.348), followed by Dfc (β = 1.787, SE = 0.789), Cfa (β = 1.320, SE = 1.379), and Cfb (β = 0.775, SE = 0.948). In contrast, Csa showed a negative effect (β = -1.219, SE = 0.861), Csb showed a moderate negative effect (β = -0.604, SE = 0.878), and Dfb showed minimal effect (β = 0.114, SE = 1.090). The intercept was not significantly different from zero (β = -0.080, SE = 0.690, χ² = 0.013, p = 0.908).

# The random effect of site showed moderate variation, with a standard deviation of 1.08 for site-level intercepts compared to a residual standard deviation of 0.72. This indicates that between-site heterogeneity is somewhat larger than within-site variation, suggesting both climate and site-specific factors contribute to variation in nitrate responses.

# These results demonstrate that climate is a strong predictor of how fire affects nitrate dynamics, with continental climates (Dfc, Dsb) showing the most positive responses and Mediterranean climates (Csa, Csb) showing negative or reduced responses to fire.
# post hoc test 

a <- emmeans::lsmeans(no3_climate_mixed_normal, pairwise ~ Climate, adjust = "tukey") # This is assessing the pair wise comparisons for the rest of the climate classifications. 

# Or if you want the full contrast results as a data frame
contrast_results <- as.data.frame(a$contrasts)
# Csa - Dfc
# Csb - Dfc


### CLIMATE NITRATE non-normalized ####
# data frame: no3_es_climate 

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
no3_climate_mixed_nn <-  lmer(Effect_size_no_watershed ~ Climate + (1 | Site), data = no3_es_climate)

# Check for equal variance
performance::check_heteroscedasticity(no3_climate_mixed_nn) # OK: Error variance appears to be homoscedastic (p = 0.433).

# Check for normality
residuals <- resid(no3_climate_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.93762, p-value = 0.000784
# when p > 0.05 can assume normality 
# NON-NORMAL?

no3_climate_mixed_nn

summary(no3_climate_mixed_nn) 

# P-value 
Anova(no3_climate_mixed_nn, type = "III")

# We analyzed the relationship between nitrate effect size and climate type using a linear mixed-effects model with site as a random effect, excluding watershed-scale studies. The model included 79 observations across 30 sites and eight climate categories.

# Climate type did not have a significant effect on nitrate effect size (χ² = 11.36, df = 7, p = 0.124). Compared to the reference climate category, the climate effects were relatively modest and varied in direction: Cfb showed the largest positive effect (β = 0.491, SE = 0.656), followed by Dfc (β = 0.525, SE = 0.558), while Csa showed the largest negative effect (β = -0.788, SE = 0.626), followed by Csb (β = -0.508, SE = 0.630). Other climate types showed smaller effects: Dfb (β = -0.229, SE = 0.779), Dsb (β = -0.132, SE = 0.937), and Cfa (β = -0.117, SE = 0.985). The intercept showed a marginally non-significant positive effect (β = 0.839, SE = 0.493, χ² = 2.90, p = 0.089).

# The random effect structure changed substantially compared to the full dataset model. Site-level variation decreased markedly (SD = 0.67 vs. 1.08) while residual variation increased slightly (SD = 0.75 vs. 0.72), and model fit improved (REML criterion: 210.9 to 199.3). This indicates that excluding watershed studies reduced between-site heterogeneity but also diminished the climate signal.

# These results suggest that when watershed-scale studies are excluded, climate becomes a less important predictor of nitrate fire responses, contrasting with the highly significant climate effects observed in the full dataset. This indicates that climate effects on nitrate dynamics may be more pronounced or detectable at larger spatial scales.


```

### CLIMATE PLOT NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r CLIMATE plot}
climate_long <- ES_nn_normal %>% 
  pivot_longer(
    cols= c(Effect_size, Effect_size_no_watershed),
    names_to = "normalization",
    values_to = "Effect_size")
  
vn = expression(paste(""*N*O[3]^"-"))
dose.labs <- as_labeller(c("DOC" = "DOC", "NO3" = "Nitrate"))

climate_fig_data <- climate_long %>%
  mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
   group_by(Climate, response_var, normalization) %>%
  dplyr::summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>%
  mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect)



# Data frame for the red X marks (only for the DOC facet)
x_marks <- data.frame(
  response_var = "DOC", # must match the facet variable exactly
  Climate = c("Dsb", "Cfa", "BSk"),
  Effect_size = 0 # position the X at x = 0 (you can change this)
)


climate_plot <- climate_fig_data %>% 
  ggplot(aes(Effect_size, Climate, color = normalization))+
  geom_pointrange(aes(xmin = mean_effect - se_effect, xmax = mean_effect + se_effect,
                      color = normalization),
                  position = position_jitter(width = 0, height = 0.2)) +
  facet_grid(~response_var, labeller = dose.labs) +
     scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Normalization"),
                     labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
  scale_y_discrete(labels=c("Dwc" = "Monsoon-influenced subarctic",
                            "Dsb" = "Mediterranean", 
                            "Dfc" = "Subarctic",
                            "Dfb" = "Warm-humid",
                            "Csb" = "Warm-Mediterranean",
                            "Csa" = "Hot-Mediterranean",
                            "Cfb" = "Subtropical highland",
                            "Cfa" = "Humid subtropical",
                            "BSk" = "Cold semi-arid")) +
  ylab("Koppen (Climate Classification)") +
  xlab("Effect Size") +
  # xlim(c(-100,100)) +
    geom_vline(xintercept = 0, linewidth = 0.5, color = "red") +
      theme_bw() +
  theme(legend.position = "top") +
  geom_text(
    data = x_marks,
    aes(x = Effect_size, y = Climate),
    label = "X",
    color = "red",
    size = 6,
    fontface = "bold",
    inherit.aes = FALSE,
    show.legend = FALSE)




climate_plot

ggsave("ES_CLIMATE_nonnormal_vs_normal.pdf",
       path = here("initial_plots", "Temporal_normalization_vs_nonnormal"),
       width = 7, height = 7,  units = "in")

```


# BP MODEL 
```{r BP - nn vs. normal anova}
### DOC ####
# Pivot into long format
long_doc_nn_normal <- doc_es_bp %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_doc_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_doc_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We examined whether the inclusion of watershed-level covariates significantly altered dissolved organic carbon (DOC_Interp) effect size estimates using a one-way ANOVA comparing effect sizes calculated with and without watershed variables. The analysis included 100 observations (df = 98 for residuals, plus 1 df for the effect type factor).

# The ANOVA revealed a highly significant difference between effect sizes calculated with watershed covariates versus those calculated without watershed covariates (F₁,₉₈ = 41.71, p < 0.001). This indicates that the inclusion of watershed-level variables substantially alters the magnitude of DOC effect size estimates in our models.

# The residual mean square error was 0.415, indicating relatively low within-group variability compared to the between-group differences in effect sizes.

### Nitrate ####
# Pivot into long format
long_no3_nn_normal <- no3_es_bp %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_no3_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_no3_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We examined whether the inclusion of watershed-level covariates significantly altered nitrate (NO3_Interp) effect size estimates using a one-way ANOVA comparing effect sizes calculated with and without watershed variables. The analysis included 170 observations (df = 168 for residuals, plus 1 df for the effect type factor).

# The ANOVA revealed a statistically significant difference between effect sizes calculated with watershed covariates versus those calculated without watershed covariates (F₁,₁₆₈ = 5.508, p = 0.020). This indicates that the inclusion of watershed-level variables significantly alters the magnitude of nitrate effect size estimates in our models.

# The residual mean square error was 1.442, indicating the typical within-group variability in effect sizes after accounting for the effect type classification.


```

```{r BP - Normalized & non-normal models}
### BP DOC normalized  ####
# Original workflow 
# data frame: doc_es_bp
doc_es_bp <- doc_es_bp %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_bp_mixed_normal <-  lmer(Effect_size ~ burn_percentage + (1 | Site), data = doc_es_bp)

# Check for equal variance
performance::check_heteroscedasticity(doc_bp_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.968)

# Check for normality
residuals <- resid(doc_bp_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.93252, p-value = 0.006892
# when p > 0.05 can assume normality 

doc_bp_mixed_normal

summary(doc_bp_mixed_normal) 

# P-value 
Anova(doc_bp_mixed_normal, type = "III")

# We analyzed the relationship between dissolved organic carbon (DOC) effect size and burn percentage using a linear mixed-effects model with site as a random effect. The model included 50 observations across 15 sites.

# Burn percentage did not have a significant effect on DOC effect size (β = -0.001, SE = 0.011, χ² = 0.008, p = 0.928). The negligible slope indicates that burn percentage has virtually no relationship with DOC response to fire. The intercept was also not significantly different from zero (β = -0.549, SE = 0.508, χ² = 1.17, p = 0.280).

# The random effect of site dominated the variance structure, with a standard deviation of 1.23 for site-level intercepts compared to a very small residual standard deviation of 0.15. This indicates that between-site variation is approximately eight times larger than within-site variation, suggesting that site-specific characteristics are far more important than burn severity in determining DOC response to fire.

# These results suggest that burn percentage is not a meaningful predictor of how fire affects DOC dynamics. Instead, site-specific factors such as soil type, vegetation, topography, or other local characteristics appear to be the primary drivers of variation in DOC fire responses, regardless of burn percentage


### BP DOC non-normalized ####
# data frame: doc_es_bp
# perform linear mixed effect model for non-normalized effect size and site as a random effect
doc_bp_mixed_nn <-  lmer(Effect_size_no_watershed ~ burn_percentage + (1 | Site), data = doc_es_bp)


# Check for equal variance
performance::check_heteroscedasticity(doc_bp_mixed_nn) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_bp_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.89323, p-value = 0.0002892
# when p > 0.05 can assume normality 

doc_bp_mixed_nn

summary(doc_bp_mixed_nn) 

# P-value 
Anova(doc_bp_mixed_nn, type = "III")

# We analyzed the relationship between dissolved organic carbon (DOC) effect size and burn percentage using a linear mixed-effects model with site as a random effect, excluding watershed-scale studies. The model included 50 observations across 15 sites.

# Burn percentage did not have a significant effect on DOC effect size (β = 0.001, SE = 0.002, χ² = 0.26, p = 0.609). The minimal slope indicates that fire severity, as measured by burn percentage, has essentially no relationship with DOC response to fire. However, the intercept was significantly greater than zero (β = 0.214, SE = 0.105, χ² = 4.13, p = 0.042), indicating a consistent positive baseline effect of fire on DOC dynamics.

# The random effect structure changed substantially compared to the full dataset model. Site-level variation decreased dramatically (SD = 0.23 vs. 1.23) while residual variation remained similar (SD = 0.16 vs. 0.15), resulting in much improved model fit (REML criterion: 33.8 to -6.3). The variance components became more balanced, with site-level variation only slightly larger than residual variation.

# These results suggest that while excluding watershed studies reveals a significant positive baseline effect of fire on DOC, burn percentage remains an unimportant predictor of DOC fire responses regardless of spatial scale. The dramatic reduction in site-level variance indicates that watershed-scale studies contributed substantial heterogeneity that masked the consistent positive fire effect observed at smaller scales.


### BP NITRATE normalized ####
# Original workflow 
# data frame: no3_es_bp
no3_es_bp <- no3_es_bp %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size and site as a random effect 
no3_bp_mixed_normal <-  lmer(Effect_size ~ burn_percentage + (1 | Site), data = no3_es_bp)

# Check for equal variance
performance::check_heteroscedasticity(no3_bp_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.967).

# Check for normality
residuals <- resid(no3_bp_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.93151, p-value = 0.0002218
# when p > 0.05 can assume normality 

no3_bp_mixed_normal

summary(no3_bp_mixed_normal) 

# P-value 
Anova(no3_bp_mixed_normal, type = "III")

# We analyzed the relationship between nitrate effect size and burn percentage using a linear mixed-effects model with site as a random effect. The model included 85 observations across 27 sites.

# Burn percentage did not have a significant effect on nitrate effect size (β = -0.009, SE = 0.011, χ² = 0.73, p = 0.394). The weak negative slope suggests that fire severity, as measured by burn percentage, has minimal relationship with nitrate response to fire. The intercept was also not significantly different from zero (β = 0.723, SE = 0.541, χ² = 1.78, p = 0.182).

# The random effect of site accounted for substantial variation in the data, with a standard deviation of 1.60 for site-level intercepts compared to a residual standard deviation of 0.65. This indicates that between-site variation is approximately 2.5 times larger than within-site variation, suggesting that site-specific characteristics are considerably more important than burn severity in determining nitrate response to fire.

# These results suggest that burn percentage is not a significant predictor of how fire affects nitrate dynamics. Instead, site-specific factors appear to be the primary drivers of variation in nitrate fire responses, with fire severity playing a minimal role in determining the magnitude of nitrogen cycling changes following fire.


### BP NITRATE non-normalized ####
# data frame: no3_es_bp

# perform linear mixed effect model for non-normalized effect size and site as a random effect 
no3_bp_mixed_nn <-  lmer(Effect_size_no_watershed ~ burn_percentage + (1 | Site), data = no3_es_bp)

# Check for equal variance
performance::check_heteroscedasticity(no3_bp_mixed_nn) # OOK: Error variance appears to be homoscedastic (p = 0.414).

# Check for normality
residuals <- resid(no3_bp_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.94004, p-value = 0.0006407
# when p > 0.05 can assume normality 
# NON-NORMAL?

no3_bp_mixed_nn

summary(no3_bp_mixed_nn) 

# P-value 
Anova(no3_bp_mixed_nn, type = "III")

# We analyzed the relationship between nitrate effect size and burn percentage using a linear mixed-effects model with site as a random effect, excluding watershed-scale studies. The model included 85 observations across 27 sites.

# Burn percentage did not have a significant effect on nitrate effect size (β = -0.008, SE = 0.006, χ² = 1.55, p = 0.213). The weak negative slope suggests that fire severity, as measured by burn percentage, has minimal relationship with nitrate response to fire. However, the intercept was highly significant and positive (β = 1.182, SE = 0.309, χ² = 14.63, p < 0.001), indicating a strong baseline positive effect of fire on nitrate dynamics.

# The random effect structure changed substantially compared to the full dataset model. Site-level variation decreased notably (SD = 0.84 vs. 1.60) while residual variation remained similar (SD = 0.67 vs. 0.65), resulting in improved model fit (REML criterion: 248.8 to 224.6). Between-site variation remained larger than within-site variation but the difference was reduced.

# These results demonstrate that while excluding watershed studies reveals a highly significant positive baseline effect of fire on nitrate dynamics, burn percentage remains an unimportant predictor of nitrate fire responses regardless of spatial scale. This suggests that fire consistently enhances nitrate availability, but the magnitude of this effect is not dependent on fire severity. Site-specific factors continue to be more important than burn intensity in determining the variability of nitrate responses to fire.

```

### BP PLOT NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r BP plot}
# Continuous 
bp_long <- BP_model %>% 
  pivot_longer(
    cols= c(Effect_size, Effect_size_no_watershed),
    names_to = "normalization",
    values_to = "Effect_size")
  
vn = expression(paste(""*N*O[3]^"-"))
dose.labs <- as_labeller(c("DOC" = "DOC", "NO3" = "Nitrate"))

bp_fig_data <- bp_long %>%
  mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
   group_by(burn_percentage, response_var, normalization) %>%
  summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>%
  mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect) %>% 
  na.omit()

bp_plot <- bp_fig_data %>%
  ggplot(aes(x = burn_percentage, y = Effect_size, color = normalization)) +
  geom_pointrange(aes(ymin = mean_effect - se_effect, ymax = mean_effect + se_effect),
                  position = position_jitter(width = 0.2, height = 0)) + # Adds jitter
  facet_grid(~response_var, labeller = dose.labs) +
  scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Normalization"),
                     labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
  labs(
    x = "Burn Percentage",
    y = "Effect Size"
  ) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  theme_bw() +
  theme(legend.position = "top")

bp_plot

ggsave("ES_BP_nonnormal_vs_normal.pdf",
       path = here("initial_plots", "Temporal_normalization_vs_nonnormal"),
       width = 7, height = 7,  units = "in")



# # #### 10 % grouping ####
# bp_fig_data_group <- bp_long %>%
#   mutate(grouped_percentage = cut(burn_percentage, breaks = seq(0, 100, by = 10), include.lowest = TRUE, labels = FALSE))
# 
# # make a grouped_percentage
# bp_fig_data_group <- bp_fig_data_group %>% 
#   mutate(grouped_percentage = as.character(grouped_percentage))
# 
# # Make a grouped percentage
# bp_fig_data_group <- bp_fig_data_group %>% 
#   mutate(burn_percentage_group = case_when(grouped_percentage == "1" ~ "0-10",
#                                            grouped_percentage == "2" ~ "11-20",
#                                            grouped_percentage == "3" ~ "21-30",
#                                            grouped_percentage == "4" ~ "31-40",
#                                            grouped_percentage == "5" ~ "41-50",
#                                            grouped_percentage == "6" ~ "51-60",
#                                            grouped_percentage == "7" ~ "61-70",
#                                            grouped_percentage == "8" ~ "71-80",
#                                            grouped_percentage == "9" ~ "81-90",
#                                            grouped_percentage == "10" ~ "91-100"))
#                                            
# # level it so it goes in order 
# bp_fig_data_group$burn_percentage_group <- factor(bp_fig_data_group$burn_percentage_group, levels = c("0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90", "91-100"))
# 
# bp_fig_data_group <- bp_fig_data_group %>% 
#   na.omit(response_var)
# 
# # calculating the mean percent_difference by grouped burn percentage
# mean_group_bp <- bp_fig_data_group %>% 
#   mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
#    group_by(burn_percentage_group, response_var, normalization) %>% 
#    summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
#             sd_effect = sd(Effect_size, na.rm = TRUE),
#             n_effect = n(),
#             Effect_size = mean(Effect_size, na.rm = TRUE)) 
# 
# 
# # %>%
# #   mutate(se_effect = sd_effect / sqrt(n_effect),
# #          lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
# #          upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect)
# 
# burn_percentage_effect_out <- mean_group_bp %>%
#   mutate(across(where(is.character), ~ifelse(is.na(.), "N/A", .))) %>%
#   mutate(across(where(is.numeric), ~ifelse(is.na(.), -9999, .))) %>% 
#   mutate(across(where(is.factor), ~ifelse(is.na(.), -9999, .)))
# 
# # Plot 
# bp_plot_grouped <- burn_percentage_effect_out %>%
#   ggplot(aes(x = burn_percentage_group, y = Effect_size, color = normalization)) +
#   geom_pointrange(aes(ymin = mean_effect - se_effect, ymax = mean_effect + se_effect),
#                   position = position_jitter(width = 0.2, height = 0)) + # Adds jitter
#   facet_grid(~response_var, labeller = dose.labs) +
#   scale_color_manual(values = c("#00AFBB", "#E7B800"),
#                      guide = guide_legend(title = "Normalization"),
#                      labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
#   labs(
#     x = "Burn Percentage",
#     y = "Effect Size"
#   ) +
#   geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
#   theme_bw() +
#   theme(legend.position = "top")
# 
# bp_plot_grouped

```


# BS MODEL 
```{r BP - nn vs. normal anova}
### DOC ####
# Pivot into long format
long_doc_nn_normal <- doc_es_dnbr %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_doc_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_doc_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We examined whether the inclusion of watershed-level covariates significantly altered dissolved organic carbon (DOC_Interp) effect size estimates using a one-way ANOVA comparing effect sizes calculated with and without watershed variables. The analysis included 54 observations (df = 52 for residuals, plus 1 df for the effect type factor).

# The ANOVA revealed a highly significant difference between effect sizes calculated with watershed covariates versus those calculated without watershed covariates (F₁,₅₂ = 19.73, p < 0.001). This indicates that the inclusion of watershed-level variables substantially alters the magnitude of DOC effect size estimates in our models.

# The residual mean square error was 0.345, indicating relatively low within-group variability compared to the between-group differences in effect sizes.

### Nitrate ####
# Pivot into long format
long_no3_nn_normal <- no3_es_dnbr %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_no3_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_no3_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We examined whether the inclusion of watershed-level covariates significantly altered nitrate (NO3_Interp) effect size estimates using a one-way ANOVA comparing effect sizes calculated with and without watershed variables. The analysis included 110 observations (df = 108 for residuals, plus 1 df for the effect type factor).

# The ANOVA revealed no statistically significant difference between effect sizes calculated with watershed covariates versus those calculated without watershed covariates (F₁,₁₀₈ = 1.799, p = 0.183). This suggests that the inclusion of watershed-level variables does not substantially alter the magnitude of nitrate effect size estimates in our models.

# The residual mean square error was 1.709, indicating the typical within-group variability in effect sizes after accounting for the effect type classification.

```

```{r BS - Normalized & non-normal models}
### BS DOC normalized  ####
# Original workflow 
# data frame: doc_es_dnbr
doc_es_dnbr <- doc_es_dnbr %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_bs_mixed_normal <-  lmer(Effect_size ~ mean_dnbr + (1 | Site), data = doc_es_dnbr)

# Check for equal variance
performance::check_heteroscedasticity(doc_bs_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.999).

# Check for normality
residuals <- resid(doc_bs_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.95668, p-value = 0.3097
# when p > 0.05 can assume normality 

doc_bs_mixed_normal

summary(doc_bs_mixed_normal) 

# P-value 
Anova(doc_bs_mixed_normal, type = "III")

# We used a linear mixed-effects model to examine the relationship between burn severity (mean dNBR) and effect size, with site as a random effect to account for spatial clustering of observations. The model included 27 observations across 12 sites.

# Burn severity had a significant negative effect on effect size (β = -0.005, SE = 0.002, t = -2.538, χ² = 6.44, p = 0.011). For each unit increase in mean dNBR (burn severity), the effect size decreased by 0.005 units. The model intercept was also significant (β = 1.606, SE = 0.805, t = 1.995, χ² = 3.98, p = 0.046).

# Random effects showed substantial variation among sites (σ² = 0.622, SD = 0.789), which was considerably larger than the residual variance (σ² = 0.015, SD = 0.122), indicating meaningful site-level differences in effect size that were appropriately accounted for in the mixed-effects framework.


### BS DOC non-normalized ####
# data frame: doc_es_dnbr
# perform linear mixed effect model for non-normalized effect size and site as a random effect
doc_bs_mixed_nn <-  lmer(Effect_size_no_watershed ~ mean_dnbr + (1 | Site), data = doc_es_dnbr)


# Check for equal variance
performance::check_heteroscedasticity(doc_bs_mixed_nn) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_bs_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = W = 0.92746, p-value = 0.05988
# when p > 0.05 can assume normality 

doc_bs_mixed_nn

summary(doc_bs_mixed_nn) 

# P-value 
Anova(doc_bs_mixed_nn, type = "III")

# We used a linear mixed-effects model to examine the relationship between burn severity (mean dNBR) and effect size (excluding watershed effects), with site as a random effect to account for spatial clustering of observations. The model included 27 observations across 12 sites.

# Burn severity showed no significant effect on effect size (β = -9.33 × 10⁻⁵, SE = 7.57 × 10⁻⁴, t = -0.123, χ² = 0.015, p = 0.902). The model intercept was also non-significant (β = 0.298, SE = 0.286, t = 1.041, χ² = 1.083, p = 0.298), indicating no detectable relationship between burn severity and effect size when watershed-level effects were excluded from the response variable.

# Random effects showed moderate variation among sites (σ² = 0.071, SD = 0.267), which was larger than the residual variance (σ² = 0.015, SD = 0.124). This suggests that while there were site-level differences in effect size, the overall effect sizes were smaller and less variable compared to the previous model that included watershed effects.

# Key differences from the previous model:
  #No significant relationships detected
  #Much smaller effect sizes overall (note the scientific notation)
  #Lower between-site variation compared to the watershed-inclusive model
  #Suggests that watershed-level processes may be important drivers of the patterns observed in the first model


### BS NITRATE normalized ####
# Original workflow 
# data frame: no3_es_dnbr
no3_es_dnbr <- no3_es_dnbr %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size and site as a random effect 
no3_bs_mixed_normal <-  lmer(Effect_size ~ mean_dnbr + (1 | Site), data = no3_es_dnbr)

# Check for equal variance
performance::check_heteroscedasticity(no3_bs_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.898).

# Check for normality
residuals <- resid(no3_bs_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96035, p-value = 0.0671
# when p > 0.05 can assume normality 

no3_bs_mixed_normal

summary(no3_bs_mixed_normal) 

# P-value 
Anova(no3_bs_mixed_normal, type = "III")

# We used a linear mixed-effects model to examine the relationship between burn severity (mean dNBR) and nitrate effect size, with site as a random effect to account for spatial clustering of observations. The model included 55 observations across 21 sites.

# Burn severity showed no significant effect on nitrate effect size (β = -1.77 × 10⁻⁵, SE = 2.56 × 10⁻³, t = -0.007, χ² < 0.001, p = 0.995). The model intercept was also non-significant (β = 0.394, SE = 0.990, t = 0.398, χ² = 0.159, p = 0.691), indicating no detectable relationship between burn severity and nitrate effect size.

# Random effects showed substantial variation among sites (σ² = 2.364, SD = 1.538), which was considerably larger than the residual variance (σ² = 0.663, SD = 0.814). This indicates meaningful site-level differences in nitrate effect size that were appropriately accounted for in the mixed-effects framework, despite the lack of a burn severity effect.


### BS NITRATE non-normalized ####
# data frame: no3_es_dnbr
# perform linear mixed effect model for non-normalized effect size and site as a random effect 
no3_bs_mixed_nn <-  lmer(Effect_size_no_watershed ~ mean_dnbr + (1 | Site), data = no3_es_dnbr)

# Check for equal variance
performance::check_heteroscedasticity(no3_bs_mixed_nn) # OK: Error variance appears to be homoscedastic (p = 0.238).

# Check for normality
residuals <- resid(no3_bs_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.93143, p-value = 0.003761
# when p > 0.05 can assume normality 
# NON-NORMAL?

no3_bs_mixed_nn

summary(no3_bs_mixed_nn) 

# P-value 
Anova(no3_bs_mixed_nn, type = "III")

# We used a linear mixed-effects model to examine the relationship between burn severity (mean dNBR) and nitrate effect size (excluding watershed effects), with site as a random effect to account for spatial clustering of observations. The model included 55 observations across 21 sites.

# Burn severity showed no significant effect on nitrate effect size (β = 7.96 × 10⁻⁵, SE = 1.56 × 10⁻³, t = 0.051, χ² = 0.003, p = 0.959). The model intercept was also non-significant (β = 0.809, SE = 0.606, t = 1.337, χ² = 1.787, p = 0.181), indicating no detectable relationship between burn severity and nitrate effect size when watershed-level effects were excluded from the response variable.

# Random effects showed moderate variation among sites (σ² = 0.650, SD = 0.806), which was smaller than the residual variance (σ² = 0.749, SD = 0.865). This pattern contrasts with the watershed-inclusive model, suggesting that watershed-level processes may contribute to site-level variation in nitrate responses.

#Comparison between models:
  # Excluding watershed effects from the nitrate response variable resulted in reduced between-site variation (SD decreased from 1.538 to 0.806) while residual variation increased slightly (SD from 0.814 to 0.865). Unlike the DOC models where watershed inclusion revealed a significant burn severity effect, both nitrate models showed no relationship with burn severity regardless of whether watershed effects were included. This suggests that nitrate responses to fire may be driven by different mechanisms than DOC responses, with burn severity playing a less important role in nitrate dynamics.

```


### BS PLOT NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r BP plot}
# Continuous 
bs_long <- DNBR_model %>% 
  pivot_longer(
    cols= c(Effect_size, Effect_size_no_watershed),
    names_to = "normalization",
    values_to = "Effect_size")
  
vn = expression(paste(""*N*O[3]^"-"))
dose.labs <- as_labeller(c("DOC" = "DOC", "NO3" = "Nitrate"))

bs_fig_data <- bs_long %>%
  mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
   group_by(mean_dnbr, response_var, normalization) %>%
  summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>%
  mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect) %>% 
  na.omit()

bs_plot <- bs_fig_data %>%
  ggplot(aes(x = mean_dnbr, y = Effect_size, color = normalization)) +
  geom_pointrange(aes(ymin = mean_effect - se_effect, ymax = mean_effect + se_effect),
                  position = position_jitter(width = 0.2, height = 0)) + # Adds jitter
  facet_grid(~response_var, labeller = dose.labs) +
  scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Normalization"),
                     labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
  labs(
    x = "Burn Severity",
    y = "Effect Size"
  ) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  theme_bw() +
  theme(legend.position = "top")

bs_plot

ggsave("ES_BS_nonnormal_vs_normal.pdf",
       path = here("initial_plots", "Temporal_normalization_vs_nonnormal"),
       width = 7, height = 7,  units = "in")
```


# Manuscript Summary Table
```{r watershed ratios for each variable of interest}
studies_file_list <- list.files(path = "inputs/Studies/meta_final/", 
                                  recursive=F, 
                                  pattern=".csv", 
                                  full.names=TRUE)

meta_df <-do.call("rbind", lapply(studies_file_list, 
                                     read.csv,
                                     na = c('-9999', 'N/A'),
                                     check.names = FALSE,
                                     stringsAsFactors=FALSE, 
                                     header=T, blank.lines.skip = TRUE, fill=TRUE)) %>% 
  select(Study_ID, Pair, Site, Area_watershed,Area_unit, Burn_Unburn) %>% 
  group_by(Study_ID, Pair) %>% 
  distinct(Pair, .keep_all = TRUE)


# crandall_metadata <- meta_df %>% 
#   filter(Study_ID != "Crandall et al. 2021")
# 
# control_impact <- meta_df %>% 
#   filter(Design_Control_Burn_Pre_Post == "Control_Reference_vs_Impact",
#          Study_ID != "Crandall et al. 2021",
#          Mean_Median_or_IndividualSample == "Individual")

# Changing the units to make sure everything is consistent in mg_N_L or mg_C_L
control_impact_units <- meta_df %>% 
  mutate(Area_watershed_km = case_when(Area_unit == 'ha' ~ Area_watershed * .01,
                                       Area_unit == 'km' ~ print(Area_watershed)))

# Creating a site characteristics data frame that I will use to merge later to add in the proper data 
site_data <- control_impact_units %>% 
  dplyr::select(Study_ID, Pair, Site, Area_watershed_km, Burn_Unburn)

site_data_unique <- site_data %>% 
  group_by(Study_ID, Pair, Area_watershed_km) %>% 
  distinct(Area_watershed_km, .keep_all = TRUE) %>% 
  select(Study_ID, Pair, Site, Area_watershed_km, Burn_Unburn)

# Generic calculation:
# Pivot longer to handle multiple Site columns and calculate ratios individually
# watershed_ratio <- watershed_ratio_wide %>%
#   pivot_longer(
#     cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
#     names_to = "Site",          # Creates a 'Site' column to track sites
#     values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
#   ) %>%
#   mutate(
#     watershed_ratio = case_when(
#       # Configurations based on specific Site and Control pairings
#       !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
#       !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
#       !is.na(Control_3_4_5) & Site %in% c("Site_3", "Site_4", "Site_5") ~ Site_value / Control_3_4_5, # Site_3, Site_4, Site_5 divided by Control_3_4_5
#       !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
#       !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
#       !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
#       TRUE ~ NA_real_ # Default case for rows without valid pairing
#     )
#   ) %>% 
#     filter(!is.na(watershed_ratio)) 

#### BP studies ####
# Number of watershed pairs
# data frames: doc_es_bp; no3_es_bp
# DOC
print(max(doc_es_bp$Site)) # 15 burned watershed pairs 
# NO3
print(max(no3_es_bp$Site)) # 27 burned watershed pairs 

# BP DOC watershed ratio ####
bp_doc <- sort(unique(doc_es_bp$Study_ID))
# Filter out by DOC BP studies
bp_doc_watershed_ratio <- site_data_unique %>%
  filter(Study_ID %in% bp_doc) %>% 
  pivot_wider(
    id_cols = Study_ID,
    names_from = Pair,
    values_from = Area_watershed_km
  )

# Pivot longer to handle multiple Site columns and calculate ratios individually
bp_doc_wr <- bp_doc_watershed_ratio %>%
  pivot_longer(
    cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
    names_to = "Site",          # Creates a 'Site' column to track sites
    values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
  ) %>%
  mutate(
    watershed_ratio = case_when(
      # Configurations based on specific Site and Control pairings
      !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
      !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
      !is.na(Control_3_4) & Site %in% c("Site_3", "Site_4") ~ Site_value / Control_3_4, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
      !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
      !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
      TRUE ~ NA_real_ # Default case for rows without valid pairing
    )
  ) %>% 
    filter(!is.na(watershed_ratio))

sort(unique(doc_es_bp$Study_ID))
sort(unique(bp_doc_wr$Study_ID))

median(bp_doc_wr$watershed_ratio) # 1.930502
mean(bp_doc_wr$watershed_ratio) # 5.325945
sd(bp_doc_wr$watershed_ratio) # 10.27117
range(bp_doc_wr$watershed_ratio) # [0.9327304, 38.6794872]

# BP NO3 watershed ratio ####
bp_no3 <- sort(unique(no3_es_bp$Study_ID))
# Filter out by DOC BP studies
bp_no3_watershed_ratio <- site_data_unique %>%
  filter(Study_ID %in% bp_no3) %>% 
  pivot_wider(
    id_cols = Study_ID,
    names_from = Pair,
    values_from = Area_watershed_km
  )

# Pivot longer to handle multiple Site columns and calculate ratios individually
bp_no3_wr <- bp_no3_watershed_ratio %>%
  pivot_longer(
    cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
    names_to = "Site",          # Creates a 'Site' column to track sites
    values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
  ) %>%
  mutate(
    watershed_ratio = case_when(
      # Configurations based on specific Site and Control pairings
      !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
      !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
      !is.na(Control_3_4) & Site %in% c("Site_3", "Site_4") ~ Site_value / Control_3_4, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control_3_4_5) & Site %in% c("Site_3", "Site_4", "Site_5") ~ Site_value / Control_3_4_5, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
      !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
      !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
      !is.na(Control) & Site == "Site_4" ~ Site_value / Control, # Generic Site_4/Control
      !is.na(Control) & Site == "Site_5" ~ Site_value / Control, # Generic Site_5/Control
      TRUE ~ NA_real_ # Default case for rows without valid pairing
    )
  ) %>% 
    filter(!is.na(watershed_ratio))

# Make sure they match 
sort(unique(no3_es_bp$Study_ID)) # 11
sort(unique(bp_no3_wr$Study_ID)) # 11 - good

median(bp_no3_wr$watershed_ratio) # 1.428571
mean(bp_no3_wr$watershed_ratio) # 3.810139
sd(bp_no3_wr$watershed_ratio) # 7.932648
range(bp_no3_wr$watershed_ratio) # [0.3024718 38.6794872]

#### BS studies ####
# Number of watershed pairs
# data frames: doc_es_dnbr, no3_es_dnbr
# DOC
print(max(doc_es_dnbr$Site)) # 12 burned watershed pairs 
# NO3
print(max(no3_es_dnbr$Site)) # 21 burned watershed pairs 

sort(unique(DNBR_model$Study_ID))

# BS DOC watershed ratio ####
bs_doc <- sort(unique(doc_es_dnbr$Study_ID))
# Filter out by DOC BP studies
bs_doc_watershed_ratio <- site_data_unique %>%
  filter(Study_ID %in% bs_doc) %>% 
  pivot_wider(
    id_cols = Study_ID,
    names_from = Pair,
    values_from = Area_watershed_km
  )

# Pivot longer to handle multiple Site columns and calculate ratios individually
bs_doc_wr <- bs_doc_watershed_ratio %>%
  pivot_longer(
    cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
    names_to = "Site",          # Creates a 'Site' column to track sites
    values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
  ) %>%
  mutate(
    watershed_ratio = case_when(
      # Configurations based on specific Site and Control pairings
      !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
      !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
      !is.na(Control_3_4) & Site %in% c("Site_3", "Site_4") ~ Site_value / Control_3_4, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
      !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
      !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
      !is.na(Control) & Site == "Site_4" ~ Site_value / Control, # Generic Site_4/Control
      TRUE ~ NA_real_ # Default case for rows without valid pairing
    )
  ) %>% 
    filter(!is.na(watershed_ratio))

# Compare - should be the same 
sort(unique(doc_es_dnbr$Study_ID)) # 5
sort(unique(bs_doc_wr$Study_ID)) # 5 - good

median(bs_doc_wr$watershed_ratio) # 1.8
mean(bs_doc_wr$watershed_ratio) #  5.804277
sd(bs_doc_wr$watershed_ratio) # 11.00979
range(bs_doc_wr$watershed_ratio) # [0.9327304 38.6794872]

# BS NO3 watershed ratio ####
bs_no3 <- sort(unique(no3_es_dnbr$Study_ID))
# Filter out by DOC BP studies
bs_no3_watershed_ratio <- site_data_unique %>%
  filter(Study_ID %in% bs_no3) %>% 
  pivot_wider(
    id_cols = Study_ID,
    names_from = Pair,
    values_from = Area_watershed_km
  )

# Pivot longer to handle multiple Site columns and calculate ratios individually
bs_no3_wr <- bs_no3_watershed_ratio %>%
  pivot_longer(
    cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
    names_to = "Site",          # Creates a 'Site' column to track sites
    values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
  ) %>%
  mutate(
    watershed_ratio = case_when(
      # Configurations based on specific Site and Control pairings
      !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
      !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
      !is.na(Control_3_4) & Site %in% c("Site_3", "Site_4") ~ Site_value / Control_3_4, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control_3_4_5) & Site %in% c("Site_3", "Site_4", "Site_5") ~ Site_value / Control_3_4_5, # Site_3, Site_4, Site 5 divided by Control_3_4_5
      !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
      !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
      !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
      !is.na(Control) & Site == "Site_4" ~ Site_value / Control, # Generic Site_4/Control
      !is.na(Control) & Site == "Site_5" ~ Site_value / Control, # Generic Site_5/Control
      TRUE ~ NA_real_ # Default case for rows without valid pairing
    )
  ) %>% 
    filter(!is.na(watershed_ratio))

# Compare - should be the same 
sort(unique(no3_es_dnbr$Study_ID)) # 8
sort(unique(bs_no3_wr$Study_ID)) # 8 - good

median(bs_no3_wr$watershed_ratio) # 1.56
mean(bs_no3_wr$watershed_ratio) # 4.294542
sd(bs_no3_wr$watershed_ratio) # 8.523296
range(bs_no3_wr$watershed_ratio) # [0.3024718, 38.6794872]


#### CLIMATE studies ####
# Number of watershed pairs
# data frames: doc_es_climate, no3_es_climate
# DOC
print(max(doc_es_climate$Site)) # 17 burned watershed pairs 
# NO3
print(max(no3_es_climate$Site)) # 30 burned watershed pairs 

sort(unique(ES_nn_normal$Study_ID))

# CLIMATE DOC watershed ratio ####
climate_doc <- sort(unique(doc_es_climate$Study_ID))
# Filter out by DOC BP studies
climate_doc_watershed_ratio <- site_data_unique %>%
  filter(Study_ID %in% climate_doc) %>% 
  pivot_wider(
    id_cols = Study_ID,
    names_from = Pair,
    values_from = Area_watershed_km
  )

# Pivot longer to handle multiple Site columns and calculate ratios individually
climate_doc_wr <- climate_doc_watershed_ratio %>%
  pivot_longer(
    cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
    names_to = "Site",          # Creates a 'Site' column to track sites
    values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
  ) %>%
  mutate(
    watershed_ratio = case_when(
      # Configurations based on specific Site and Control pairings
      !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
      !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
      !is.na(Control_3_4) & Site %in% c("Site_3", "Site_4") ~ Site_value / Control_3_4, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
      !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
      !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
      !is.na(Control) & Site == "Site_4" ~ Site_value / Control, # Generic Site_4/Control
      TRUE ~ NA_real_ # Default case for rows without valid pairing
    )
  ) %>% 
    filter(!is.na(watershed_ratio))

# Compare - should be the same 
sort(unique(doc_es_climate$Study_ID)) # 8
sort(unique(climate_doc_wr$Study_ID)) # 8 - good

median(climate_doc_wr$watershed_ratio) # 1.85
mean(climate_doc_wr$watershed_ratio) # 4.912453
sd(climate_doc_wr$watershed_ratio) # 9.678452
range(climate_doc_wr$watershed_ratio) # [0.9327304 38.6794872]

# CLIMATE NO3 watershed ratio ####
climate_no3 <- sort(unique(no3_es_climate$Study_ID))
# Filter out by DOC BP studies
climate_no3_watershed_ratio <- site_data_unique %>%
  filter(Study_ID %in% climate_no3) %>% 
  pivot_wider(
    id_cols = Study_ID,
    names_from = Pair,
    values_from = Area_watershed_km
  )

# Pivot longer to handle multiple Site columns and calculate ratios individually
climate_no3_wr <- climate_no3_watershed_ratio %>%
  pivot_longer(
    cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
    names_to = "Site",          # Creates a 'Site' column to track sites
    values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
  ) %>%
  mutate(
    watershed_ratio = case_when(
      # Configurations based on specific Site and Control pairings
      !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
      !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
      !is.na(Control_3_4) & Site %in% c("Site_3", "Site_4") ~ Site_value / Control_3_4, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control_3_4_5) & Site %in% c("Site_3", "Site_4", "Site_5") ~ Site_value / Control_3_4_5, # Site_3, Site_4, Site_5 divided by Control_3_4_5
      !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
      !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
      !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
      !is.na(Control) & Site == "Site_4" ~ Site_value / Control, # Generic Site_4/Control
      !is.na(Control) & Site == "Site_5" ~ Site_value / Control, # Generic Site_5/Control
      TRUE ~ NA_real_ # Default case for rows without valid pairing
    )
  ) %>% 
    filter(!is.na(watershed_ratio))

# Compare - should be the same 
sort(unique(no3_es_climate$Study_ID)) # 13
sort(unique(climate_no3_wr$Study_ID)) # 13 - good

median(climate_no3_wr$watershed_ratio) # 1.493043
mean(climate_no3_wr$watershed_ratio) # 3.557971
sd(climate_no3_wr$watershed_ratio) # 7.554198
range(climate_no3_wr$watershed_ratio) # [0.2428571 38.6794872]


#### TSF studies ####
# Number of watershed pairs
# data frames: doc_es_tsf, no3_es_tsf
# DOC
print(max(doc_es_tsf$Site)) # 21 burned watershed pairs 
# NO3
print(max(no3_es_tsf$Site)) # 34 burned watershed pairs 

sort(unique(ES_nn_normal$Study_ID))


# TSF DOC watershed ratio ####
tsf_doc <- sort(unique(doc_es_tsf$Study_ID))
# Filter out by DOC TSF studies
tsf_doc_watershed_ratio <- site_data_unique %>%
  filter(Study_ID %in% tsf_doc) %>% 
  pivot_wider(
    id_cols = Study_ID,
    names_from = Pair,
    values_from = Area_watershed_km
  ) %>% 
  filter(Study_ID != "Abbott et al. 2021")

# Pivot longer to handle multiple Site columns and calculate ratios individually
tsf_doc_wr <- tsf_doc_watershed_ratio %>%
  pivot_longer(
    cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
    names_to = "Site",          # Creates a 'Site' column to track sites
    values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
  ) %>%
  mutate(
    watershed_ratio = case_when(
      # Configurations based on specific Site and Control pairings
      !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
      !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
      !is.na(Control_3_4) & Site %in% c("Site_3", "Site_4") ~ Site_value / Control_3_4, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
      !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
      !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
      !is.na(Control) & Site == "Site_4" ~ Site_value / Control, # Generic Site_4/Control
      TRUE ~ NA_real_ # Default case for rows without valid pairing
    )
  ) %>% 
    filter(!is.na(watershed_ratio))

# Compare - should be the same 
sort(unique(doc_es_tsf$Study_ID)) # 10
sort(unique(tsf_doc_wr$Study_ID)) # 9 - no Abbott

median(tsf_doc_wr$watershed_ratio) # 1.787272
mean(tsf_doc_wr$watershed_ratio) # 4.300585
sd(tsf_doc_wr$watershed_ratio) #  9.007161
range(tsf_doc_wr$watershed_ratio) # [0.4616477 38.6794872]

# TSF NO3 watershed ratio ####
tsf_no3 <- sort(unique(no3_es_tsf$Study_ID))
# Filter out by DOC TSF studies
tsf_no3_watershed_ratio <- site_data_unique %>%
  filter(Study_ID %in% tsf_no3) %>% 
  pivot_wider(
    id_cols = Study_ID,
    names_from = Pair,
    values_from = Area_watershed_km
  ) %>% 
  filter(Study_ID != "Abbott et al. 2021")

# Pivot longer to handle multiple Site columns and calculate ratios individually
tsf_no3_wr <- tsf_no3_watershed_ratio %>%
  pivot_longer(
    cols = starts_with("Site"), # Pivoting all Site columns (Site_1, Site_2, Site_3, etc.)
    names_to = "Site",          # Creates a 'Site' column to track sites
    values_to = "Site_value"    # Store numeric values of Site in the 'Site_value' column
  ) %>%
  mutate(
    watershed_ratio = case_when(
      # Configurations based on specific Site and Control pairings
      !is.na(Control_1) & Site == "Site_1" ~ Site_value / Control_1, # Site_1/Control_1
      !is.na(Control_2) & Site == "Site_2" ~ Site_value / Control_2, # Site_2/Control_2
      !is.na(Control_3_4) & Site %in% c("Site_3", "Site_4") ~ Site_value / Control_3_4, # Site_3, Site_4, divided by Control_3_4
      !is.na(Control_3_4_5) & Site %in% c("Site_3", "Site_4", "Site_5") ~ Site_value / Control_3_4_5, # Site_3, Site_4, Site_5 divided by Control_3_4_5
      !is.na(Control) & Site == "Site_1" ~ Site_value / Control, # Generic Site_1/Control
      !is.na(Control) & Site == "Site_2" ~ Site_value / Control, # Generic Site_2/Control
      !is.na(Control) & Site == "Site_3" ~ Site_value / Control, # Generic Site_3/Control
      !is.na(Control) & Site == "Site_4" ~ Site_value / Control, # Generic Site_4/Control
      !is.na(Control) & Site == "Site_5" ~ Site_value / Control, # Generic Site_5/Control
      TRUE ~ NA_real_ # Default case for rows without valid pairing
    )
  ) %>% 
    filter(!is.na(watershed_ratio))

# Compare - should be the same 
sort(unique(no3_es_tsf$Study_ID)) # 15
sort(unique(tsf_no3_wr$Study_ID)) # 14 - no Abbott

median(tsf_no3_wr$watershed_ratio) # 1.298625
mean(tsf_no3_wr$watershed_ratio) #  3.310277
sd(tsf_no3_wr$watershed_ratio) #  7.235802
range(tsf_no3_wr$watershed_ratio) # [0.3024718 38.6794872]

```

# Manuscript density figure
```{r density plot}
# Run all chunks before this chunk to get the following data frames:
# bp_doc_wr
# bp_no3_wr

# bs_doc_wr
# bs_no3_wr

# climate_doc_wr
# climate_no3_wr

# tsf_doc_wr
# tsf_no3_wr

# Merge for DOC ####
merged_doc <- bind_rows(
  bp_doc_wr %>% mutate(response_var = "DOC", source = "Burn percent"),
  bs_doc_wr %>% mutate(response_var = "DOC", source = "Burn severity"),
  climate_doc_wr %>% mutate(response_var = "DOC", source = "Climate"),
  tsf_doc_wr %>% mutate(response_var = "DOC", source = "Time-since-fire")
)

# Create the density plot
ggplot(merged_doc, aes(x = watershed_ratio, fill = source)) +
  geom_density(alpha = 0.7) +
  labs(
    x = "Watershed Ratio",
    y = "Density",
    title = "Distribution of Watershed Ratio by Source",
    fill = "Source"
  ) +
  theme_minimal()


# Merge for nitrate ####
merged_no3 <- bind_rows(
  bp_no3_wr %>% mutate(response_var = "Nitrate", source = "Burn percent"),
  bs_no3_wr %>% mutate(response_var = "Nitrate", source = "Burn severity"),
  climate_no3_wr %>% mutate(response_var = "Nitrate", source = "Climate"),
  tsf_no3_wr %>% mutate(response_var = "Nitrate", source = "Time-since-fire")
)

# Combine both datasets
combined_data <- bind_rows(merged_doc, merged_no3)


# 2 panel plot ####
ggplot(combined_data, aes(x = watershed_ratio, color = source)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~ response_var, ncol = 2) +
  labs(
    x = "Watershed Ratio",
    y = "Density",
    fill = "Source"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    panel.spacing = unit(1, "lines"),
    legend.position = c(0.825, 0.85),  
    legend.text = element_text(size = 18),
    legend.title = element_blank(),
    legend.margin = margin(5, 5, 5, 5)
  )

ggsave("watershed_ratio_DOC_nitrate.png",
       path = here("initial_plots", "Temporal_normalization_vs_nonnormal"),
       width = 7, height = 7,  units = "in")

# Filter out big ones
small_data <- combined_data %>% 
  filter(watershed_ratio < 10)


# 2 panel plot ####
ggplot(small_data, aes(x = watershed_ratio, color = source)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~ response_var, ncol = 2) +
  labs(
    x = "Watershed Ratio",
    y = "Density",
    fill = "Source"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    panel.spacing = unit(1, "lines"),
    legend.position = c(0.825, 0.85),  
    legend.text = element_text(size = 18),
    legend.title = element_blank(),
    legend.margin = margin(5, 5, 5, 5)
  )

ggsave("watershed_ratio_DOC_nitrate.png",
       path = here("initial_plots", "Temporal_normalization_vs_nonnormal"),
       width = 7, height = 7,  units = "in")




```





















































# ================================ GRAVEYARD ==================================================#
### OLD DATA and COMPARE 
```{r}
### ORIGINAL DATA MODEL ####
TSF_orig <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "TSF_ES_Fig_data.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, year, Time_Since_Fire, Effect_size_pair, Effect_size)

# Want to make the TSF continuous so I am gonna put actual years to them 
TSF_orig <- TSF_orig %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      Study_ID == "Abbott et al. 2021" & occurrence == 3 ~ "10", # Third value
      Study_ID == "Abbott et al. 2021" & occurrence == 4 ~ "11", # Fourth value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire))

study_site_data <- read_csv(here("inputs", "catchment_characteristics", "Fire_name_Lat_Long.csv"), na = c('-9999', 'N/A')) %>% 
  dplyr::select(Study_ID, Site, Effect_size_pair) %>% 
  na.omit()

unique(study_site_data$Study_ID) # should be all 18 studies. 

TSF_model <- full_join(TSF_orig, study_site_data) %>% 
  na.omit()# joining the TSF dataframe used to make the figure and study site data to match sites because we will be including site as a random effect

doc_es_tsf_OLD <- TSF_model %>% 
  filter(response_var == "DOC_Interp")

no3_es_tsf <- TSF_model %>% 
  filter(response_var == "NO3_Interp")

# Continuous 
doc_tsf_mixed <-  lmer(Effect_size ~ Time_Since_Fire + (1 | Site), data = doc_es_tsf_OLD)

# Check for equal variance
performance::check_heteroscedasticity(doc_tsf_mixed) # OK: Error variance appears to be homoscedastic (p = 0.976).

# Check for normality
residuals <- resid(doc_tsf_mixed)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.97186, p-value = 0.2639
# when p > 0.05 can assume normality 
 
summary(doc_tsf_mixed) # df: 40.9851, t-value: -4.792, p-value: 2.47e-05 # TSF is significant for DOC 





### COMPARE OLD VS. NEW ####
# Find common Study_IDs to see paired comparisons
common_studies <- intersect(doc_es_tsf$Study_ID, doc_es_tsf_OLD$Study_ID)
print(paste("Common studies:", length(common_studies)))

# Check which studies are unique to each dataset
unique_to_new <- setdiff(doc_es_tsf$Study_ID, doc_es_tsf_OLD$Study_ID)
unique_to_old <- setdiff(doc_es_tsf_OLD$Study_ID, doc_es_tsf$Study_ID)

print("Studies only in current dataset:")
print(unique_to_new)
print("Studies only in OLD dataset:")
print(unique_to_old)

library(ggplot2)

# Create combined dataset for plotting
comparison_data <- data.frame(
  Effect_size = c(doc_es_tsf$Effect_size, doc_es_tsf_OLD$Effect_size),
  Dataset = rep(c("Current", "OLD"), c(nrow(doc_es_tsf), nrow(doc_es_tsf_OLD)))
)

# Boxplot comparison
ggplot(comparison_data, aes(x = Dataset, y = Effect_size)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.6) +
  theme_minimal() +
  labs(title = "Effect Size Comparison Between Datasets")

# Density plot
ggplot(comparison_data, aes(x = Effect_size, fill = Dataset)) +
  geom_density(alpha = 0.6) +
  theme_minimal() +
  labs(title = "Effect Size Distribution Comparison")

# Two-sample t-test
t.test(doc_es_tsf$Effect_size, doc_es_tsf_OLD$Effect_size)

# Welch Two Sample t-test (unequal variances)
t.test(doc_es_tsf$Effect_size, doc_es_tsf_OLD$Effect_size, var.equal = FALSE)

# Non-parametric alternative (Mann-Whitney U test)
wilcox.test(doc_es_tsf$Effect_size, doc_es_tsf_OLD$Effect_size)

OLD_summary <- doc_es_tsf_OLD %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(OLD = n())

NEW_summary <- doc_es_tsf %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(NEW = n())

comp_summary <- left_join(OLD_summary, NEW_summary, by = c("Study_ID", "response_var"))
# Old has 5 more Mast & Clow values. Why? 
# Old has 2 more Murphy et al. 2018 values. Why?

# Create comparison dataframe
effect_size_comparison <- doc_es_tsf %>%
  dplyr::select(Study_ID, Effect_size, Time_Since_Fire) %>%
  rename(Current_Effect_size = Effect_size) %>%
  full_join(
    doc_es_tsf_OLD %>%
      dplyr::select(Study_ID, Effect_size, Time_Since_Fire) %>%
      rename(OLD_Effect_size = Effect_size),
    by = "Study_ID"
  ) %>%
  mutate(
    Match = case_when(
      is.na(Current_Effect_size) | is.na(OLD_Effect_size) ~ FALSE,
      Current_Effect_size == OLD_Effect_size ~ TRUE,
      TRUE ~ FALSE
    )
  )


ES_OLD <- doc_es_tsf_OLD %>% 
  filter(Study_ID == "Mast & Clow; 2008") %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Time_Since_Fire)
# DOUBLE EFFECT SIZE VALUES

ES_NEW <- doc_es_tsf %>% 
  filter(Study_ID == "Mast & Clow; 2008") %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Time_Since_Fire)

```

# Normalized Effect size - orig
```{r Normalized Effect size - orig}
ES_orig <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_Size_UPDATED.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, year, Time_Since_Fire, Effect_size_pair, Effect_size)

```

# TSF original data frame
```{r TSF original data frame}
TSF_orig <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "TSF_ES_Fig_data.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, year, Time_Since_Fire, Effect_size_pair, Effect_size)

```

# Compare TSF orig and ES orig
```{r Compare TSF orig and ES orig}
ES_summary <- ES_orig %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(ES_orig = n())

TSF_summary <- TSF_orig %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(TSF_count = n())

comp_summary_2 <- left_join(ES_summary, TSF_summary, by = c("Study_ID", "response_var"))

# TSF_summary has 10 more observations compared to ES_orig. There is more data in the TSF_summary specifically for Hauer & Spencer and Murphy et al. 2018.

```

# Non-normalized Effect size - new
```{r Non-normalized Effect size - new}
ES_nn <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_size_orig_nonnormal.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, Time_Since_Fire, Effect_size_pair, Effect_size, Effect_size_no_watershed)

nn_summary <- ES_nn %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(ES_nn = n())

comp_summary_3 <- left_join(comp_summary_2, nn_summary, by = c("Study_ID", "response_var"))

# Non-normalized effect size matches the ES_Orig here so perhaps we should go with the ES_Orig data frame as our finalized data frame. 

```

# TSF LMER - Normalized (original)
```{r DOC}
### DOC ####
# Read in study site data to get site
study_site_data <- read_csv(here("inputs", "catchment_characteristics", "Fire_name_Lat_Long.csv"), na = c('-9999', 'N/A'))  %>% 
  dplyr::select(Study_ID, Site, Effect_size_pair) %>% 
  na.omit()

# Check to see if all the studies are there.
unique(study_site_data$Study_ID) # should be all 17 studies. 
# good 

# Joining to add the site data because we will be including site as a random effect
ES_orig_model <- full_join(ES_orig, study_site_data) %>% 
  na.omit() 

# Filter data for DOC and nitrate separately
DOC_data <- ES_orig_model %>% 
  filter(response_var == "DOC_Interp")

# Want to make the TSF continuous so I am gonna put actual years to them 
DOC_model <- DOC_data %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire))

unique(DOC_model$Study_ID)

```

# TSF vs climate site number 
```{r}
### WHY DOES CLIMATE N03 HAVE MORE WATERSHEDS THAN TSF NO3 ###
# print(max(no3_es_climate$Site)) # 30 burned watershed pairs 
# print(max(no3_es_tsf$Site)) # 34 burned watershed pairs 
# 
# sort(unique(no3_es_climate$Study_ID)) # 13 studies
# # Coombs, Crandall, Gerla, Gluns, Hauer, Hickenbottom, Mast, Murphy, Neary, Oliver, Tiedemann, Uzun, Writer
# 
# sort(unique(no3_es_tsf$Study_ID)) # 8 studies
# # Coombs, Crandall, Gerla, Hauer, Murphy, Oliver, Uzun, Writer
# # Why no Gluns, Hickenbottom, Mast, Neary, Tiedemann?
# 
# # This is how the data frames are being filtered out. 
# # They are both from the 
# no3_es_tsf <- DNBR_model %>% 
#   filter(response_var == "NO3_Interp") %>% 
#   dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Effect_size_no_watershed, 
#                 Time_Since_Fire)
# 
# #### CLIMATE MODEL DATAFRAME ###
# # Filter out studies that are greater than 10 years 
# no3_es_climate <- ES_nn_normal %>% 
#   filter(response_var == "NO3_Interp") %>% 
#   dplyr::select(Study_ID, response_var, Climate, Biome, Effect_size_pair, Effect_size, Effect_size_no_watershed) %>% 
#   filter(Study_ID != "Abbott et al. 2021") %>% 
#   filter(Study_ID != "Rhea et al. 2021")


```

# AMP START HERE FOR BOOTSTRAPPING VARIANCE ANALYSIS 
Compare effect size variances across all watersheds between normalized/non-normalized data

Chunk 2180 - 2439 is an attempt to do it myself. It looks like # HYPOTHESIS i SUPPORTED (overall): Area normalization significantly reduces effect-size variance. Watershed area explains a meaningful share of cross-watershed variability. Area normalization explains 62% [49%, 72%] of the raw effect-size variance. Anything over 50% tells us that area is the dominant source of cross-watershed variability. When I break it up by nitrate and DOC for looking at variance difference by response variable: both Variables are positive = area normalization captures watershed variance. When looking at variance RATIO by response variable (log scale: both variables >1 = normalization shrinks variance.

```{r Read in final data frame with non-normal and normalized}
rm(list=ls(all=T)) #this clears your Environment

library(pacman)
p_load(tidyverse,
       here,
       forecastML, 
       zoo,
       # hrbrthemes,
       viridis, 
       ggpmisc,
       RobustLinearReg,
       lme4,
       car)

# Read in effect size data frame for normalized and non-normalized data
ES_nn_normal <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_size_orig_nonnormal.csv"))

ES_nn_normal <- ES_nn_normal %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      Study_ID == "Abbott et al. 2021" & occurrence == 3 ~ "10", # Third value
      Study_ID == "Abbott et al. 2021" & occurrence == 4 ~ "11", # Fourth value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire)) %>% 
  dplyr::select(Study_ID:Effect_size_no_watershed)

# Make dataframe long format
ES_long <- ES_nn_normal %>%
  mutate(row_id = row_number()) %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed),
    names_to  = "type",
    values_to = "value"
  ) %>%
  mutate(type = factor(type))


# Levene's Test of all data ####
cat("=== Standard Levene's Test ===\n")
levene_obs <- leveneTest(value ~ type, data = ES_long, center = median)
print(levene_obs)
obs_F <- levene_obs$`F value`[1]

# Quick look at the observed variances
cat("\nVar(Effect_size):", var(ES_nn_normal$Effect_size), "\n")
cat("Var(Effect_size_no_watershed):", var(ES_nn_normal$Effect_size_no_watershed), "\n")
cat("Variance ratio:", 
    var(ES_nn_normal$Effect_size) / var(ES_nn_normal$Effect_size_no_watershed), "\n\n")


# BOOTSTRAPPING ####
set.seed(2024)
n_boot <- 10000

# Pivot to wide so each row_id is one observation with both the raw and area-normalized effect size
ES_wide <- ES_long %>%
  pivot_wider(
    id_cols  = c(Study_ID, response_var, Climate, Biome,
                 Effect_size_pair, Time_Since_Fire, row_id),
    names_from  = type,
    values_from = value
  ) %>%
  drop_na(Effect_size, Effect_size_no_watershed)

# paired bootstrap of variance difference & ratio Resamples row_ids (whole pairs) to keep the pairing intact
boot_var_compare <- function(df, n_boot = 10000) {
  n <- nrow(df)
  replicate(n_boot, {
    bs  <- slice_sample(df, n = n, replace = TRUE)
    v_r <- var(bs$Effect_size,              na.rm = TRUE)
    v_n <- var(bs$Effect_size_no_watershed, na.rm = TRUE)
    c(var_raw = v_r, var_norm = v_n,
      diff = v_r - v_n,              # + → normalization shrinks variance
      ratio = v_r / v_n)            # >1 → normalization shrinks variance
  }) %>% t() %>% as_tibble()
}

# OVERALL comparison (all response variables pooled)
boot_overall <- boot_var_compare(ES_wide, n_boot)

overall_summary <- boot_overall %>%
  pivot_longer(everything(), names_to = "metric") %>%
  group_by(metric) %>%
  summarise(
    mean    = mean(value),
    median  = median(value),
    ci_lo   = quantile(value, 0.025),
    ci_hi   = quantile(value, 0.975),
    .groups = "drop"
  )

overall_summary %>% filter(metric %in% c("diff", "ratio"))

# Quick programmatic check
if (overall_summary %>%
      filter(metric == "diff") %>%
      pull(ci_lo) > 0) {
  cat("✔ HYPOTHESIS i SUPPORTED (overall):
      Area normalization significantly reduces effect-size variance.
      Watershed area explains a meaningful share of cross-watershed variability.\n")
} else if (overall_summary %>%
             filter(metric == "diff") %>%
             pull(ci_hi) < 0) {
  cat("✔ HYPOTHESIS ii SUPPORTED (overall):
      Area normalization increases effect-size variance.
      Additional watershed characteristics likely modulate fire response.\n")
} else {
  cat("⚠ INCONCLUSIVE (overall):
      95% CI of the variance difference spans zero.\n")
}

# HYPOTHESIS i SUPPORTED (overall): Area normalization significantly reduces effect-size variance. Watershed area explains a meaningful share of cross-watershed variability.

# Proportion of raw variance "removed" by normalization
boot_overall <- boot_overall %>%
  mutate(prop_explained = (var_raw - var_norm) / var_raw)

prop_summary <- boot_overall %>%
  summarise(
    mean_prop   = mean(prop_explained),
    ci_lo       = quantile(prop_explained, 0.025),
    ci_hi       = quantile(prop_explained, 0.975)
  )

cat(sprintf(
  "\nArea normalization explains %.0f%% [%.0f%%, %.0f%%] of the raw effect-size variance.\n",
  prop_summary$mean_prop * 100,
  prop_summary$ci_lo     * 100,
  prop_summary$ci_hi     * 100
))

# Area normalization explains 62% [49%, 72%] of the raw effect-size variance.

# 0.50 -> Area is a dominant source of cross-watershed variability 
# 0.20 –> 0.50 -> Area is meaningful but other factors matter too
# 0.20 -> Area alone is a weak first-order correction 
# 0 -> Normalization actually inflates variance → hypothesis ii 

boot_by_response <- boot_by_response %>%
  mutate(
    direction = case_when(
      boot_diff_ci_lo > 0  ~ "i  – norm reduces variance",
      boot_diff_ci_hi < 0  ~ "ii – norm increases variance",
      TRUE                 ~ "inconclusive"
    ),
    prop_explained = obs_diff / obs_var_raw
  )

boot_by_response %>%
  select(response_var, n, obs_var_raw, obs_var_norm,
         obs_diff, prop_explained,
         boot_diff_ci_lo, boot_diff_ci_hi, direction) %>%
  arrange(desc(prop_explained)) %>%
  print(width = Inf)

# One-sided "p-value": proportion of resamples where raw > normalised
p_overall <- mean(boot_overall$diff > 0)

cat("\n===== OVERALL VARIANCE COMPARISON =====\n")
print(overall_summary)
cat(sprintf("\nP(var_raw > var_norm) = %.3f\n", p_overall))

# BY RESPONSE VARIABLE ####
boot_by_response <- ES_wide %>%
  group_by(response_var) %>%
  group_split() %>%
  map_dfr(function(df) {
    rv    <- unique(df$response_var)
    n_obs <- nrow(df)
    b     <- boot_var_compare(df, n_boot)

    tibble(
      response_var       = rv,
      n                  = n_obs,
      obs_var_raw        = var(df$Effect_size,              na.rm = TRUE),
      obs_var_norm       = var(df$Effect_size_no_watershed, na.rm = TRUE),
      obs_diff           = obs_var_raw - obs_var_norm,
      obs_ratio          = obs_var_raw / obs_var_norm,
      boot_diff_mean     = mean(b$diff),
      boot_diff_ci_lo    = quantile(b$diff, 0.025),
      boot_diff_ci_hi    = quantile(b$diff, 0.975),
      boot_ratio_mean    = mean(b$ratio),
      boot_ratio_ci_lo   = quantile(b$ratio, 0.025),
      boot_ratio_ci_hi   = quantile(b$ratio, 0.975),
      p_raw_gt_norm      = mean(b$diff > 0)
    )
  })

cat("\n===== BY RESPONSE VARIABLE =====\n")
print(boot_by_response, width = Inf)

# PLOTS
# 5a – Overlapping densities of bootstrapped variances (overall)
boot_overall %>%
  select(var_raw, var_norm) %>%
  pivot_longer(everything(),
               names_to  = "type",
               values_to = "variance") %>%
  mutate(type = recode(type,
    var_raw  = "Raw effect size",
    var_norm = "Area-normalised effect size")) %>%
  ggplot(aes(variance, fill = type)) +
  geom_density(alpha = 0.45, colour = "grey30") +
  labs(title = "Bootstrapped variance distributions (all responses pooled)",
       x = "Variance of effect sizes", y = "Density", fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")

# 5b – Density of variance DIFFERENCE with zero reference line
boot_overall %>%
  ggplot(aes(diff)) +
  geom_density(fill = "steelblue", alpha = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red", linewidth = 0.8) +
  annotate("text", x = 0, y = Inf, vjust = 2, hjust = -0.1,
           label = paste0("P(diff > 0) = ", round(p_overall, 3)),
           colour = "red", size = 4.5) +
  labs(title = "Bootstrapped variance difference (Raw − Normalised)",
       subtitle = "Positive -> normalization reduces variability  |  Negative -> normalization increases variability",
       x = "Var(raw) − Var(normalized)", y = "Density") +
  theme_minimal(base_size = 13)

# 5c – Forest-style plot: variance difference by response variable
boot_by_response %>%
  ggplot(aes(x = reorder(response_var, obs_diff), y = obs_diff)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "red") +
  geom_pointrange(aes(ymin = boot_diff_ci_lo, ymax = boot_diff_ci_hi),
                  size = 0.6) +
  coord_flip() +
  labs(title = "Variance difference by response variable",
       subtitle = "Positive = area normalization captures watershed variance\nNegative = additional modulators likely",
       x = NULL,
       y = "Var(raw) − Var(normalized)  [95 % bootstrap CI]") +
  theme_minimal(base_size = 13)

# Both Variables are Positive = area normalization captures watershed variance

# 5d – Variance RATIO by response variable (log scale)
boot_by_response %>%
  ggplot(aes(x = reorder(response_var, obs_ratio), y = obs_ratio)) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "red") +
  geom_pointrange(aes(ymin = boot_ratio_ci_lo, ymax = boot_ratio_ci_hi),
                  size = 0.6) +
  scale_y_log10() +
  coord_flip() +
  labs(title = "Variance ratio by response variable",
       subtitle = ">1 = normalisation shrinks variance  |  <1 = normalisation inflates variance",
       x = NULL,
       y = "Var(raw) / Var(normalised)  [95 % CI, log scale]") +
  theme_minimal(base_size = 13)

# Both variables >1 = normalization shrinks variance
```

# BY ANALYTE
```{r}
# library(tidyverse)
# 
# set.seed(2024)
# n_boot <- 10000
# 
# # 1.  Separate DOC and NO3, pivot to wide (raw + normalised)
# make_wide <- function(df, pattern) {
#   df %>%
#     filter(str_detect(response_var, pattern)) %>%
#     pivot_wider(
#       id_cols  = c(Study_ID, response_var, Climate, Biome,
#                    Effect_size_pair, Time_Since_Fire, row_id),
#       names_from  = type,
#       values_from = value
#     ) %>%
#     drop_na(Effect_size, Effect_size_no_watershed)
# }
# 
# doc_data <- make_wide(ES_long, "DOC")
# no3_data <- make_wide(ES_long, "NO3")
# 
# cat(sprintf("DOC observations: %d\nNO3 observations: %d\n",
#             nrow(doc_data), nrow(no3_data)))
# 
# # 2.  Bootstrap Var(DOC)/Var(NO3) for raw AND normalised
# #     DOC and NO3 resampled independently (different studies)
# boot_results <- replicate(n_boot, {
# 
#   doc_bs <- slice_sample(doc_data, n = nrow(doc_data), replace = TRUE)
#   no3_bs <- slice_sample(no3_data, n = nrow(no3_data), replace = TRUE)
# 
#   # --- raw effect sizes ---
#   v_doc_raw <- var(doc_bs$Effect_size, na.rm = TRUE)
#   v_no3_raw <- var(no3_bs$Effect_size, na.rm = TRUE)
# 
# 
#   # --- area-normalised effect sizes ---
#   v_doc_norm <- var(doc_bs$Effect_size_no_watershed, na.rm = TRUE)
#   v_no3_norm <- var(no3_bs$Effect_size_no_watershed, na.rm = TRUE)
# 
#   c(var_doc_raw  = v_doc_raw,
#     var_no3_raw  = v_no3_raw,
#     ratio_raw    = v_doc_raw / v_no3_raw,
#     var_doc_norm = v_doc_norm,
#     var_no3_norm = v_no3_norm,
#     ratio_norm   = v_doc_norm / v_no3_norm)
# 
# }) %>% t() %>% as_tibble()
# 
# # 3.  Summarise the two variance ratios
# summarise_ratio <- function(x, label) {
#   tibble(
#     data_type  = label,
#     mean_ratio = mean(x),
#     median_ratio = median(x),
#     ci_lo      = quantile(x, 0.025),
#     ci_hi      = quantile(x, 0.975),
#     p_DOC_gt   = mean(x > 1)
#   )
# }
# 
# ratio_summary <- bind_rows(
#   summarise_ratio(boot_results$ratio_raw,  "Raw effect sizes"),
#   summarise_ratio(boot_results$ratio_norm, "Area-normalised effect sizes")
# )
# 
# cat("\n===== Var(DOC) / Var(NO3) — Bootstrapped =====\n")
# print(ratio_summary, width = Inf)
# 
# # 4.  Is the DOC-vs-NO3 variance gap ITSELF different between raw and normalized data?
# boot_results <- boot_results %>%
#   mutate(ratio_diff = ratio_raw - ratio_norm)
# 
# ratio_diff_summary <- boot_results %>%
#   summarise(
#     mean_diff = mean(ratio_diff),
#     ci_lo     = quantile(ratio_diff, 0.025),
#     ci_hi     = quantile(ratio_diff, 0.975),
#     p_raw_gap_larger = mean(ratio_diff > 0)
#   )
# 
# cat("\n===== Ratio_raw − Ratio_norm =====\n")
# print(ratio_diff_summary, width = Inf)
# 
# # 5.  Plots
# 
# # 5a — Overlapping densities of the two variance ratios
# boot_results %>%
#   select(ratio_raw, ratio_norm) %>%
#   pivot_longer(everything(),
#                names_to = "type", values_to = "ratio") %>%
#   mutate(type = recode(type,
#     ratio_raw  = "Raw effect sizes",
#     ratio_norm = "Area-normalised")) %>%
#   ggplot(aes(ratio, fill = type)) +
#   geom_density(alpha = 0.45, colour = "grey30") +
#   geom_vline(xintercept = 1, linetype = "dashed", colour = "red",
#              linewidth = 0.8) +
#   labs(title    = "Bootstrapped Var(DOC) / Var(NO₃)",
#        subtitle = "> 1 → DOC more variable than NO₃",
#        x = "Variance ratio", y = "Density", fill = NULL) +
#   theme_minimal(base_size = 13) +
#   theme(legend.position = "top")
# 
# # 5b — Density of the DIFFERENCE between the two ratios
# boot_results %>%
#   ggplot(aes(ratio_diff)) +
#   geom_density(fill = "darkorange", alpha = 0.55) +
#   geom_vline(xintercept = 0, linetype = "dashed", colour = "red",
#              linewidth = 0.8) +
#   annotate("text", x = 0, y = Inf, vjust = 2, hjust = -0.1,
#            label = sprintf("P(gap larger in raw) = %.3f",
#                            ratio_diff_summary$p_raw_gap_larger),
#            colour = "red", size = 4.5) +
#   labs(title = "Does normalisation change the DOC-vs-NO₃ variance gap?",
#        subtitle = "Positive → raw data exaggerates DOC's extra variability",
#        x = "Ratio(raw) − Ratio(normalised)", y = "Density") +
#   theme_minimal(base_size = 13)
# 
# # 5c — Paired dot plot of observed variances
# tibble(
#   solute = rep(c("DOC", "NO₃"), 2),
#   type   = rep(c("Raw", "Area-normalised"), each = 2),
#   variance = c(var(doc_data$Effect_size),
#                var(no3_data$Effect_size),
#                var(doc_data$Effect_size_no_watershed),
#                var(no3_data$Effect_size_no_watershed))
# ) %>%
#   ggplot(aes(type, variance, colour = solute, group = solute)) +
#   geom_point(size = 4) +
#   geom_line(linewidth = 1) +
#   labs(title = "Observed effect-size variance",
#        x = NULL, y = "Variance", colour = NULL) +
#   theme_minimal(base_size = 13) +
#   theme(legend.position = "top")
```

### ATTEMPT 1
```{r}
# 
# # 1. Reshape data to long format for Levene's test
# df_long <- ES_nn_normal %>%
#   mutate(row_id = row_number()) %>%
#   pivot_longer(
#     cols = c(Effect_size, Effect_size_no_watershed),
#     names_to  = "type",
#     values_to = "value"
#   ) %>%
#   mutate(type = factor(type))
# 
# # 2. Standard Levene's Test (for reference)
# cat("=== Standard Levene's Test ===\n")
# levene_obs <- leveneTest(value ~ type, data = df_long, center = median)
# print(levene_obs)
# obs_F <- levene_obs$`F value`[1]
# 
# # Quick look at the observed variances
# cat("\nVar(Effect_size):             ", var(ES_nn_normal$Effect_size), "\n")
# cat("Var(Effect_size_no_watershed):", var(ES_nn_normal$Effect_size_no_watershed), "\n")
# cat("Variance ratio:               ", 
#     var(ES_nn_normal$Effect_size) / var(ES_nn_normal$Effect_size_no_watershed), "\n\n")
# 
# # 3. Bootstrapped Variance Ratio & Difference (paired resampling)
# #    - Resamples ROWS to preserve pairing
# #    - Gives you a CI for how the variances compare
# set.seed(42)
# n_boot <- 10000
# n      <- nrow(ES_nn_normal)
# 
# boot_var_norm    <- numeric(n_boot)
# boot_var_nonorm  <- numeric(n_boot)
# boot_ratio       <- numeric(n_boot)
# boot_diff        <- numeric(n_boot)
# 
# for (i in seq_len(n_boot)) {
#   idx <- sample(1:n, n, replace = TRUE)
#   boot_var_norm[i]   <- var(ES_nn_normal$Effect_size[idx])
#   boot_var_nonorm[i] <- var(ES_nn_normal$Effect_size_no_watershed[idx])
#   boot_ratio[i]      <- boot_var_norm[i] / boot_var_nonorm[i]
#   boot_diff[i]       <- boot_var_norm[i] - boot_var_nonorm[i]
# }
# 
# cat("=== Bootstrapped Variance Ratio (Normalized / Non-normalized) ===\n")
# cat("Mean ratio :", mean(boot_ratio), "\n")
# cat("95% CI     :", quantile(boot_ratio, c(0.025, 0.975)), "\n\n")
# 
# cat("=== Bootstrapped Variance Difference (Normalized - Non-normalized) ===\n")
# cat("Mean diff  :", mean(boot_diff), "\n")
# cat("95% CI     :", quantile(boot_diff, c(0.025, 0.975)), "\n")
# # If the CI for the ratio excludes 1 (or the diff CI excludes 0), variances differ
# cat("CI excludes 1 (ratio)?:", 
#     !(quantile(boot_ratio, 0.025) <= 1 & quantile(boot_ratio, 0.975) >= 1), "\n")
# cat("CI excludes 0 (diff)? :", 
#     !(quantile(boot_diff, 0.025) <= 0 & quantile(boot_diff, 0.975) >= 0), "\n\n")
# 
# # 4. Permutation-based Levene's Test (better p-value for small n)
# #    - Under H0 the two columns are exchangeable within each row
# #    - Randomly swap normalized <-> non-normalized within rows
# set.seed(42)
# n_perm  <- 10000
# perm_F  <- numeric(n_perm)
# 
# for (i in seq_len(n_perm)) {
#   # For each row, randomly decide whether to swap the two values
#   swap <- rbinom(n, 1, 0.5) == 1
#   
#   perm_norm   <- ifelse(swap,
#                         ES_nn_normal$Effect_size_no_watershed,
#                         ES_nn_normal$Effect_size)
#   perm_nonorm <- ifelse(swap,
#                         ES_nn_normal$Effect_size,
#                         ES_nn_normal$Effect_size_no_watershed)
#   
#   perm_long <- tibble(
#     value = c(perm_norm, perm_nonorm),
#     type  = factor(rep(c("norm", "nonorm"), each = n))
#   )
#   
#   perm_F[i] <- leveneTest(value ~ type, data = perm_long, center = median)$`F value`[1]
# }
# 
# perm_p <- mean(perm_F >= obs_F)
# cat("=== Permutation Levene's Test ===\n")
# cat("Observed F :", obs_F, "\n")
# cat("Permutation p-value:", perm_p, "\n\n")
# 
# # 5. Plots
# par(mfrow = c(1, 3))
# 
# # A) Bootstrap distribution of variance ratio
# hist(boot_ratio, breaks = 50, col = "steelblue", border = "white",
#      main = "Bootstrap: Variance Ratio",
#      xlab = "Var(normalized) / Var(non-normalized)")
# abline(v = 1, col = "red", lwd = 2, lty = 2)
# abline(v = quantile(boot_ratio, c(0.025, 0.975)), col = "orange", lwd = 2, lty = 3)
# 
# # B) Bootstrap distribution of variance difference
# hist(boot_diff, breaks = 50, col = "steelblue", border = "white",
#      main = "Bootstrap: Variance Difference",
#      xlab = "Var(normalized) - Var(non-normalized)")
# abline(v = 0, col = "red", lwd = 2, lty = 2)
# abline(v = quantile(boot_diff, c(0.025, 0.975)), col = "orange", lwd = 2, lty = 3)
# 
# # C) Permutation null distribution of Levene's F
# hist(perm_F, breaks = 50, col = "grey70", border = "white",
#      main = "Permutation Null: Levene's F",
#      xlab = "F statistic")
# abline(v = obs_F, col = "red", lwd = 2)
# legend("topright", legend = paste("p =", round(perm_p, 4)), bty = "n", cex = 1.2)
# 
# par(mfrow = c(1, 1))

```


