---
title: "Temporal_normalization_vs_nonnormal"
output: html_document
date: "2026-02-12"
editor_options: 
  chunk_output_type: console
---

## Read in libraries 
```{r Jake/Mac Load Packages and}
#for Jake/mac

rm(list=ls(all=T)) #this clears your Environment

library(pacman)
p_load(tidyverse,
       here,
       forecastML, 
       zoo,
       # hrbrthemes,
       viridis, 
       ggpmisc,
       RobustLinearReg,
       lme4,
       car)

```

### Build Model data frame for all variables of interest - BP, BS, TSF, CLIMATE 
```{r Build model data frames}
# Read in effect size data frame for normalized and non-normalized data
ES_nn_normal <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_size_orig_nonnormal.csv"))

ES_nn_normal <- ES_nn_normal %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      Study_ID == "Abbott et al. 2021" & occurrence == 3 ~ "10", # Third value
      Study_ID == "Abbott et al. 2021" & occurrence == 4 ~ "11", # Fourth value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire)) %>% 
  dplyr::select(Study_ID:Effect_size_no_watershed)

# Read in the study and site csv 
study_site_data <- read_csv(here("inputs", "catchment_characteristics", "Fire_name_Lat_Long.csv"), na = c('-9999', 'N/A')) %>% 
  dplyr::select(Study_ID, Site, Fire_year, Effect_size_pair, burn_percentage) %>% 
  rename(year = Fire_year) %>% 
  filter(!Study_ID == "Murphy et al. 2015") %>% 
  na.omit()

unique(study_site_data$Study_ID) # Should be 13 studies because we are missing Abbott.
# Abbott because of merging burned and unburned sites. 
# Murphy et al. 2015 was already accounted for in the Murphy et al. 2018 


### BURN PERCENTAGE MODEL DATAFRAME ####
bp_clip <- study_site_data %>% 
  dplyr::select(Study_ID, Site, year, Effect_size_pair, burn_percentage)

BP_ES <- full_join(bp_clip, ES_nn_normal) %>% 
  na.omit() %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Effect_size_no_watershed, burn_percentage) # joining the BP dataframe used to make the figure and study site data to match sites because we will be including site as a random effect

unique(BP_ES$Study_ID) # 13 studies

# Merge to include site
BP_model <- full_join(BP_ES, study_site_data) %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Site, Effect_size, Effect_size_no_watershed, burn_percentage)

# filter DOC and no3 out 
doc_es_bp <- BP_model %>% 
  filter(response_var == "DOC_Interp")

no3_es_bp <- BP_model %>% 
  filter(response_var == "NO3_Interp")

#### DNBR BURN SEVERITY MODEL DATAFRAME ##### 
DNBR_df <- read_csv(here("Output_for_analysis", "14_Meta_calculate_burn_severity", "Burn_severity_ES_fig_data.csv")) %>% 
  dplyr::select(Study_ID, response_var, fire_name, Effect_size_pair, mean_dnbr, burn_severity_category, burn_percentage, Time_Since_Fire)

DNBR_model <- left_join(DNBR_df, ES_nn_normal, by = c("Study_ID", "Effect_size_pair", "response_var", "Time_Since_Fire")) %>% 
  na.omit()

# Filter by response 
doc_es_dnbr <- DNBR_model %>% 
  filter(response_var == "DOC_Interp")

no3_es_dnbr <- DNBR_model %>% 
  filter(response_var == "NO3_Interp")

#### TSF MODEL DATAFRAME ##### 
doc_es_tsf <- ES_nn_normal %>% 
  filter(response_var == "DOC_Interp") %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Effect_size_no_watershed, 
                Time_Since_Fire)

no3_es_tsf <- DNBR_model %>% 
  filter(response_var == "NO3_Interp") %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Effect_size_no_watershed, 
                Time_Since_Fire)

#### CLIMATE MODEL DATAFRAME ##### 
doc_es_climate <- ES_nn_normal %>% 
  filter(response_var == "DOC_Interp") %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, Effect_size_pair, Effect_size, Effect_size_no_watershed)

no3_es_climate <- ES_nn_normal %>% 
  filter(response_var == "NO3_Interp") %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, Effect_size_pair, Effect_size, Effect_size_no_watershed)

```

# TSF NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r}
# Pivot into long format
long_doc_nn_normal <- doc_es_tsf %>%
  pivot_longer(
    cols = c(Effect_size, Effect_size_no_watershed), # Columns to pivot
    names_to = "Effect_type",                      # New column to identify which column the value comes from
    values_to = "Effect_value"                     # New column to store the values
  )

# Create a list to store ANOVA results for each response_var group
anova_results <- list()

# Perform ANOVA for each unique response_var
unique_response_vars <- unique(long_doc_nn_normal$response_var)

for (response in unique_response_vars) {
  # Filter data for the current response_var group
  group_data <- long_doc_nn_normal %>% filter(response_var == response)
  
  # Perform ANOVA
  anova_result <- aov(Effect_value ~ Effect_type, data = group_data)
  
  # Store the result in the list
  anova_results[[response]] <- summary(anova_result)
}

# Print each ANOVA result by response_var group
for (response in names(anova_results)) {
  cat("\nANOVA Results for Response Variable:", response, "\n")
  print(anova_results[[response]])
}

# We conducted a one-way ANOVA to compare dissolved organic carbon (DOC) effect sizes calculated with and without watershed-level controls. The analysis included 88 observations and revealed a statistically significant difference between the two effect size calculations (F₁,₈₆ = 21.23, p < 0.001).

# The effect type (with vs. without watershed controls) explained a substantial portion of the variance in effect sizes, with an effect size sum of squares of 11.15 compared to residual sum of squares of 45.17. This significant difference indicates that the inclusion or exclusion of watershed-level controls meaningfully impacts the calculated effect sizes for fire impacts on DOC concentrations.

# The significant difference between effect size calculations suggests that watershed-level factors play an important role in determining the magnitude of fire effects on dissolved organic carbon. This highlights the importance of controlling for watershed characteristics when assessing fire impacts, as failure to account for these factors may lead to biased effect size estimates.

```

# TSF MODEL 
```{r TSF - Normalized & non-normal models}
### TSF DOC normalized  ####
# Original workflow 
# data frame: doc_es_tsf 
doc_es_tsf <- doc_es_tsf %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_tsf_mixed_normal <-  lmer(Effect_size ~ Time_Since_Fire + (1 | Site), data = doc_es_tsf)


# Check for equal variance
performance::check_heteroscedasticity(doc_tsf_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.993).

# Check for normality
residuals <- resid(doc_tsf_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96893, p-value = 0.2772
# when p > 0.05 can assume normality 

doc_tsf_mixed_normal

summary(doc_tsf_mixed_normal) 

# P-value 
Anova(doc_tsf_mixed_normal, type = "III")

# We analyzed the relationship between time since fire and dissolved organic carbon (DOC) effect sizes using a linear mixed-effects model with site as a random effect to account for multiple observations from the same study locations. The model included 44 observations from 21 sites.

# The mixed-effects model revealed a marginally significant negative relationship between time since fire and DOC effect size (β = -0.034, SE = 0.018, t = -1.876, p = 0.071). This suggests that the magnitude of fire effects on DOC may decrease over time, with effect sizes becoming less negative (closer to zero) as time progresses following fire disturbance.

# The model intercept was -0.321 (SE = 0.246, t = -1.305, p = 0.205), indicating that at time zero (immediately post-fire), the predicted effect size was negative but not significantly different from zero. The random effect of site accounted for substantial variation in the data (variance = 1.165, SD = 1.079), while residual variance was relatively low (variance = 0.016, SD = 0.125), suggesting that site-level differences were an important source of variation in fire effects on DOC.

# A Type III Wald chi-square test confirmed the marginal significance of the time since fire effect (χ² = 3.519, df = 1, p = 0.061), indicating a trend toward recovery of DOC levels over time following fire disturbance.


### TSF DOC non-normalized ####
# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_tsf_mixed_nn <-  lmer(Effect_size_no_watershed ~ Time_Since_Fire + (1 | Site), data = doc_es_tsf)


# Check for equal variance
performance::check_heteroscedasticity(doc_tsf_mixed_nn) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_tsf_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.97221, p-value = 0.3613
# when p > 0.05 can assume normality 

doc_tsf_mixed_nn

summary(doc_tsf_mixed_nn) 

# P-value 
Anova(doc_tsf_mixed_nn, type = "III")

# We analyzed the relationship between time since fire and dissolved organic carbon (DOC) effect sizes calculated without watershed controls using a linear mixed-effects model with site as a random effect. The model included 44 observations from 21 sites.

#The mixed-effects model revealed a highly significant negative relationship between time since fire and DOC effect size without watershed controls (β = -0.057, SE = 0.010, t = -5.454, p < 0.001). This indicates that fire effects on DOC decrease substantially over time, with effect sizes becoming smaller as time progresses following fire disturbance.

#The model intercept was significantly positive (β = 0.391, SE = 0.072, t = 5.402, p < 0.001), indicating that immediately post-fire (time zero), there was a strong positive effect of fire on DOC concentrations. The random effect of site showed moderate variation (variance = 0.069, SD = 0.263), while residual variance remained low (variance = 0.016, SD = 0.127), suggesting that while site-level differences contributed to variation, the temporal pattern was consistent across sites.

# Type III Wald chi-square tests confirmed the highly significant effects of both the intercept (χ² = 29.18, df = 1, p < 0.001) and time since fire (χ² = 29.74, df = 1, p < 0.001). These results demonstrate a clear pattern of DOC recovery over time, with initially elevated post-fire concentrations declining significantly as ecosystems recover from fire disturbance.

# Note: The strong significance in this model contrasts with the marginal effects found when watershed controls were included, suggesting that watershed-level factors may mask or confound the temporal recovery patterns in DOC concentrations following fire.

### TSF NITRATE normalized ####
# Original workflow 
# data frame: doc_es_tsf 
no3_es_tsf <- no3_es_tsf %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
no3_tsf_mixed_normal <-  lmer(Effect_size ~ Time_Since_Fire + (1 | Site), data = no3_es_tsf)

# Check for equal variance
performance::check_heteroscedasticity(no3_tsf_mixed_normal) # OOK: Error variance appears to be homoscedastic (p = 0.876).

# Check for normality
residuals <- resid(no3_tsf_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96477, p-value = 0.1067
# when p > 0.05 can assume normality 

no3_tsf_mixed_normal

summary(no3_tsf_mixed_normal) 

# P-value 
Anova(no3_tsf_mixed_normal, type = "III")

# We examined the relationship between time since fire and nitrate (NO₃) effect sizes using a linear mixed effects model with site as a random effect. The model included 55 observations across 21 sites.

# The analysis revealed no significant relationship between time since fire and NO₃ effect size (β = 0.029, SE = 0.092, t = 0.312, p = 0.757). The intercept was also not significantly different from zero (β = 0.346, SE = 0.377, t = 0.917, p = 0.368).

# Random effects showed substantial variation among sites (SD = 1.47), indicating considerable site-specific differences in baseline conditions, while residual variation was moderate (SD = 0.83). The Type III Wald test confirmed the non-significant effect of time since fire (χ² = 0.097, df = 1, p = 0.755).

# These results suggest that nitrate levels in the studied systems do not show a clear temporal pattern of recovery following fire disturbance. The large among-site variation indicates that site-specific factors may be more important than time since fire in determining nitrate response patterns post-fire.

### TSF NITRATE non-normalized ####
# data frame: no3_es_tsf 

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
no3_tsf_mixed_non_normal <-  lmer(Effect_size_no_watershed ~ Time_Since_Fire + (1 | Site), data = no3_es_tsf)

# Check for equal variance
performance::check_heteroscedasticity(no3_tsf_mixed_non_normal) # OK: Error variance appears to be homoscedastic (p = 0.067).

# Check for normality
residuals <- resid(no3_tsf_mixed_non_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.89441, p-value = 0.00016
# when p > 0.05 can assume normality 
# NON-NORMAL?

no3_tsf_mixed_non_normal

summary(no3_tsf_mixed_non_normal) 

# P-value 
Anova(no3_tsf_mixed_non_normal, type = "III")

# We examined the relationship between time since fire and nitrate (NO₃) effect sizes (excluding watershed-scale studies) using a linear mixed effects model with site as a random effect. The model included 55 observations across 21 sites.

# The analysis revealed no significant relationship between time since fire and NO₃ effect size (β = 0.125, SE = 0.088, t = 1.43, p = 0.160). However, the intercept was significantly greater than zero (β = 0.692, SE = 0.239, t = 2.90, p = 0.008), indicating that nitrate effect sizes were significantly positive immediately following fire disturbance.

# Random effects showed relatively low variation among sites (SD = 0.39), while residual variation was higher (SD = 0.99), suggesting that within-site variability was greater than among-site differences. The Type III Wald test confirmed the non-significant effect of time since fire (χ² = 2.04, df = 1, p = 0.154) and the significant intercept effect (χ² = 8.42, df = 1, p = 0.004).

# These results suggest that fire initially produces a significant positive effect on nitrate levels, but this effect does not change systematically over time since fire. The significant baseline effect indicates that fire disturbance consistently elevates nitrate concentrations in the immediate post-fire period across the studied systems.

# COMPARE
# Effect_size model (normalized): No significant baseline effect (intercept p = 0.368) and no relationship with time since fire (p = 0.757). This model showed higher among-site variation (SD = 1.47) relative to residual variation (SD = 0.83).

# Effect_size_no_watershed model (non-normal): Significant positive baseline effect immediately post-fire (intercept = 0.692, p = 0.008) but still no relationship with time since fire (p = 0.160). This model showed lower among-site variation (SD = 0.39) but higher residual variation (SD = 0.99).

# The key difference is that excluding watershed-scale studies reveals a consistent, significant positive effect of fire on nitrate concentrations in the immediate post-fire period. This suggests that watershed-scale studies may dilute or mask the direct effects of fire on nitrate, possibly due to differences in spatial scale, methodology, or the influence of broader landscape processes. The reduced among-site variation in the no-watershed model indicates that watershed studies contributed substantially to site-level heterogeneity in the full dataset.

```

### TSF PLOT NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r TSF plot}
tsf_nn_normal_long <-  doc_es_tsf %>%
  pivot_longer(
    cols= c(Effect_size, Effect_size_no_watershed),
    names_to = "normalization",
    values_to = "Effect_size"
  )

vn = expression(paste(""*N*O[3]^"-"))
dose.labs <- as_labeller(c("DOC" = "DOC", "NO3" = "Nitrate"))

# Means and SD
tsf_nn_normal_plot = tsf_nn_normal_long %>%
  mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
   group_by(Time_Since_Fire, response_var, normalization) %>%
  summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>%
  mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect)
View(tsf_nn_normal_plot)

tsf_plot = tsf_nn_normal_plot %>% 
  ggplot(aes(Time_Since_Fire, Effect_size, color = normalization))+
  geom_pointrange(aes(ymin = mean_effect - se_effect, ymax = mean_effect + se_effect,
                      color = normalization)) +
  facet_grid(~response_var, labeller = dose.labs) +
     scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Normalization"),
                     labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
  xlab("Time Since Fire") +
  ylab("Effect Size") +
    geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
      theme_bw()

tsf_plot

```


# CLIMATE MODEL 
```{r CLIMATE - Normalized & non-normal models}
### CLIMATE DOC normalized  ####
# Original workflow 
# data frame: doc_es_climate
doc_es_climate <- doc_es_climate %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_climate_mixed_normal <-  lmer(Effect_size ~ Climate + (1 | Site), data = doc_es_climate)

# Check for equal variance
performance::check_heteroscedasticity(doc_climate_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.994)

# Check for normality
residuals <- resid(doc_climate_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96704, p-value = 0.237
# when p > 0.05 can assume normality 

doc_climate_mixed_normal

summary(doc_climate_mixed_normal) 

# P-value 
Anova(doc_climate_mixed_normal, type = "III")

# We examined the relationship between climate type and dissolved organic carbon (DOC) effect sizes using a linear mixed effects model with site as a random effect. The model included 44 observations across 21 sites, comparing six climate types (reference climate plus Csa, Csb, Dfb, Dfc, and Dwc).

# The analysis revealed no significant differences in DOC effect sizes among climate types (overall climate effect: χ² = 1.12, df = 5, p = 0.952). None of the individual climate contrasts were significant: Csa (β = -0.227, SE = 0.835, t = -0.272, p = 0.789), Csb (β = 0.685, SE = 1.078, t = 0.636, p = 0.535), Dfb (β = 0.086, SE = 0.902, t = 0.096, p = 0.925), Dfc (β = 0.114, SE = 0.862, t = 0.132, p = 0.897), and Dwc (β = 0.578, SE = 1.363, t = 0.424, p = 0.678). The reference climate intercept was also not significantly different from zero (β = -0.518, SE = 0.680, t = -0.762, p = 0.458).

# Random effects showed substantial variation among sites (SD = 1.18) compared to very low residual variation (SD = 0.13), indicating that site-specific factors account for most of the variability in DOC response to fire, while within-site measurements were highly consistent.

# These results suggest that climate type does not significantly influence how DOC concentrations respond to fire disturbance, with site-specific characteristics being the primary driver of variability in post-fire DOC dynamics.

### CLIMATE DOC non-normalized ####
# data frame: doc_es_climate
# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_climate_mixed_nn <-  lmer(Effect_size_no_watershed ~ Climate + (1 | Site), data = doc_es_climate)


# Check for equal variance
performance::check_heteroscedasticity(doc_climate_mixed_nn) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_climate_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.9674, p-value = 0.2441
# when p > 0.05 can assume normality 

doc_climate_mixed_nn

summary(doc_climate_mixed_nn) 

# P-value 
Anova(doc_climate_mixed_nn, type = "III")

# We examined the relationship between climate type and dissolved organic carbon (DOC) effect sizes on non-normalized data using a linear mixed effects model with site as a random effect. The model included 44 observations across 21 sites, comparing six climate types (reference climate plus Csa, Csb, Dfb, Dfc, and Dwc).
# Comparing Cfb to the rest of them?


# The analysis revealed a significant overall effect of climate type on DOC effect sizes (χ² = 13.87, df = 5, p = 0.016). Among individual climate contrasts, the Csb climate showed a marginally significant positive effect (β = 0.641, SE = 0.324, t = 1.98, p = 0.067), while other climate types showed non-significant effects: Csa (β = 0.286, SE = 0.249, t = 1.15, p = 0.270), Dfb (β = 0.030, SE = 0.269, t = 0.11, p = 0.914), Dfc (β = -0.341, SE = 0.257, t = -1.33, p = 0.206), and Dwc (β = 0.013, SE = 0.407, t = 0.032, p = 0.975). The reference climate intercept was not significantly different from zero (β = 0.117, SE = 0.200, t = 0.582, p = 0.570).

# Random effects showed moderate variation among sites (SD = 0.34) with very low residual variation (SD = 0.13), indicating that site-specific factors still account for most variability, but this variation was substantially reduced compared to the full dataset.

# These results suggest that when watershed-scale studies are excluded, climate type does significantly influence DOC response to fire, with Mediterranean climates (particularly Csb) showing the strongest positive responses to fire disturbance.

# post hoc test 
a <- emmeans::lsmeans(doc_climate_mixed_nn, pairwise ~ Climate, adjust = "tukey") # This is assessing the pair wise comparisons for the rest of the climate classifications. 

# Or if you want the full contrast results as a data frame
contrast_results <- as.data.frame(a$contrasts)
# Csb to Dfc is significantly different 


### CLIMATE NITRATE normalized ####
# Original workflow 
# data frame: no3_es_climate
no3_es_climate <- no3_es_climate %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
no3_climate_mixed_normal <-  lmer(Effect_size ~ Climate + (1 | Site), data = no3_es_climate)

# Check for equal variance
performance::check_heteroscedasticity(no3_climate_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.956).

# Check for normality
residuals <- resid(no3_climate_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.95116, p-value = 0.00281
# when p > 0.05 can assume normality 

no3_climate_mixed_normal

summary(no3_climate_mixed_normal) 

# P-value 
Anova(no3_climate_mixed_normal, type = "III")

# We examined the relationship between climate type and nitrate (NO₃) effect sizes using a linear mixed effects model with site as a random effect. The model included 85 observations across 34 sites, comparing nine climate types (reference climate plus Cfa, Cfb, Csa, Csb, Dfb, Dfc, Dsb, and Dwc).

# The analysis revealed a highly significant overall effect of climate type on NO₃ effect sizes (χ² = 40.02, df = 8, p < 0.001). Individual climate contrasts showed varying effects relative to the reference climate, though none reached statistical significance in the individual t-tests: Cfa (β = 1.320, SE = 1.357, t = 0.973, p = 0.341), Cfb (β = 0.775, SE = 0.933, t = 0.831, p = 0.416), Csa (β = -1.220, SE = 0.847, t = -1.439, p = 0.163), Csb (β = -0.604, SE = 0.864, t = -0.699, p = 0.492), and Dfb (β = 0.114, SE = 1.073, t = 0.106, p = 0.916). The reference climate intercept was not significantly different from zero (β = -0.080, SE = 0.679, t = -0.117, p = 0.908).

# Random effects showed substantial variation among sites (SD = 1.06) and moderate residual variation (SD = 0.72), indicating that both site-specific factors and within-site variability contribute meaningfully to overall variance in nitrate responses.

# These results demonstrate that climate type is a significant predictor of how nitrate concentrations respond to fire disturbance, with the overall pattern suggesting systematic differences among climate zones despite individual contrasts not reaching significance, likely due to multiple comparison effects.

# post hoc test 
a <- emmeans::lsmeans(no3_climate_mixed_normal, pairwise ~ Climate, adjust = "tukey") # This is assessing the pair wise comparisons for the rest of the climate classifications. 

# Or if you want the full contrast results as a data frame
contrast_results <- as.data.frame(a$contrasts)
# Csa - Dfc
# Csb - Dfc


### CLIMATE NITRATE non-normalized ####
# data frame: no3_es_climate 

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
no3_climate_mixed_nn <-  lmer(Effect_size_no_watershed ~ Climate + (1 | Site), data = no3_es_climate)

# Check for equal variance
performance::check_heteroscedasticity(no3_climate_mixed_nn) # OK: Error variance appears to be homoscedastic (p = 0.555).

# Check for normality
residuals <- resid(no3_climate_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.94189, p-value = 0.0008131
# when p > 0.05 can assume normality 
# NON-NORMAL?

no3_climate_mixed_nn

summary(no3_climate_mixed_nn) 

# P-value 
Anova(no3_climate_mixed_nn, type = "III")

# A linear mixed-effects model was fitted using REML to assess the effect of climate classification on Effect_size_no_watershed, with Site included as a random intercept. The model included 85 observations across 34 sites (REML criterion = 214.6).
# Random Effects:
  # Site variance = 0.514 (SD = 0.717)
  # Residual variance = 0.547 (SD = 0.740)

# Fixed Effects:
# The intercept was estimated at 0.839 (SE = 0.512, p = 0.120). None of the individual climate categories showed statistically significant effects (all p > 0.05), though ClimateDfc had the largest positive estimate (0.754, SE = 0.568, p = 0.203), and ClimateCsa had the largest negative estimate (-0.801, SE = 0.649, p = 0.232).


```

### CLIMATE PLOT NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r CLIMATE plot}
climate_long <- ES_nn_normal %>% 
  pivot_longer(
    cols= c(Effect_size, Effect_size_no_watershed),
    names_to = "normalization",
    values_to = "Effect_size")
  
vn = expression(paste(""*N*O[3]^"-"))
dose.labs <- as_labeller(c("DOC" = "DOC", "NO3" = "Nitrate"))

climate_fig_data <- climate_long %>%
  mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
   group_by(Climate, response_var, normalization) %>%
  dplyr::summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>%
  mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect)



# Data frame for the red X marks (only for the DOC facet)
x_marks <- data.frame(
  response_var = "DOC", # must match the facet variable exactly
  Climate = c("Dsb", "Cfa", "BSk"),
  Effect_size = 0 # position the X at x = 0 (you can change this)
)


climate_plot <- climate_fig_data %>% 
  ggplot(aes(Effect_size, Climate, color = normalization))+
  geom_pointrange(aes(xmin = mean_effect - se_effect, xmax = mean_effect + se_effect,
                      color = normalization),
                  position = position_jitter(width = 0, height = 0.2)) +
  facet_grid(~response_var, labeller = dose.labs) +
     scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Normalization"),
                     labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
  scale_y_discrete(labels=c("Dwc" = "Monsoon-influenced subarctic",
                            "Dsb" = "Mediterranean", 
                            "Dfc" = "Subarctic",
                            "Dfb" = "Warm-humid",
                            "Csb" = "Warm-Mediterranean",
                            "Csa" = "Hot-Mediterranean",
                            "Cfb" = "Subtropical highland",
                            "Cfa" = "Humid subtropical",
                            "BSk" = "Cold semi-arid")) +
  ylab("Koppen (Climate Classification)") +
  xlab("Effect Size") +
  # xlim(c(-100,100)) +
    geom_vline(xintercept = 0, linewidth = 0.5, color = "red") +
      theme_bw() +
  theme(legend.position = "top") +
  geom_text(
    data = x_marks,
    aes(x = Effect_size, y = Climate),
    label = "X",
    color = "red",
    size = 6,
    fontface = "bold",
    inherit.aes = FALSE,
    show.legend = FALSE)




climate_plot

ggsave("ES_CLIMATE_nonnormal_vs_normal.pdf",
       path = here("initial_plots", "Temporal_normalization_vs_nonnormal"),
       width = 7, height = 7,  units = "in")

```


# BP MODEL 
```{r BP - Normalized & non-normal models}
### BP DOC normalized  ####
# Original workflow 
# data frame: doc_es_bp
doc_es_bp <- doc_es_bp %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_bp_mixed_normal <-  lmer(Effect_size ~ burn_percentage + (1 | Site), data = doc_es_bp)

# Check for equal variance
performance::check_heteroscedasticity(doc_bp_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.994)

# Check for normality
residuals <- resid(doc_bp_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96704, p-value = 0.237
# when p > 0.05 can assume normality 

doc_bp_mixed_normal

summary(doc_bp_mixed_normal) 

# P-value 
Anova(doc_bp_mixed_normal, type = "III")


### BP DOC non-normalized ####
# data frame: doc_es_bp
# perform linear mixed effect model for non-normalized effect size and site as a random effect
doc_bp_mixed_nn <-  lmer(Effect_size_no_watershed ~ burn_percentage + (1 | Site), data = doc_es_bp)


# Check for equal variance
performance::check_heteroscedasticity(doc_bp_mixed_nn) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_bp_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.9674, p-value = 0.2441
# when p > 0.05 can assume normality 

doc_bp_mixed_nn

summary(doc_bp_mixed_nn) 

# P-value 
Anova(doc_bp_mixed_nn, type = "III")


### BP NITRATE normalized ####
# Original workflow 
# data frame: no3_es_bp
no3_es_bp <- no3_es_bp %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size and site as a random effect 
no3_bp_mixed_normal <-  lmer(Effect_size ~ burn_percentage + (1 | Site), data = no3_es_bp)

# Check for equal variance
performance::check_heteroscedasticity(no3_bp_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.956).

# Check for normality
residuals <- resid(no3_bp_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.95116, p-value = 0.00281
# when p > 0.05 can assume normality 

no3_bp_mixed_normal

summary(no3_bp_mixed_normal) 

# P-value 
Anova(no3_bp_mixed_normal, type = "III")


### BP NITRATE non-normalized ####
# data frame: no3_es_bp

# perform linear mixed effect model for non-normalized effect size and site as a random effect 
no3_bp_mixed_nn <-  lmer(Effect_size_no_watershed ~ Climate + (1 | Site), data = no3_es_bp)

# Check for equal variance
performance::check_heteroscedasticity(no3_bp_mixed_nn) # OK: Error variance appears to be homoscedastic (p = 0.555).

# Check for normality
residuals <- resid(no3_bp_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.94189, p-value = 0.0008131
# when p > 0.05 can assume normality 
# NON-NORMAL?

no3_bp_mixed_nn

summary(no3_bp_mixed_nn) 

# P-value 
Anova(no3_bp_mixed_nn, type = "III")


```

### BP PLOT NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r BP plot}
# Continuous 
bp_long <- BP_model %>% 
  pivot_longer(
    cols= c(Effect_size, Effect_size_no_watershed),
    names_to = "normalization",
    values_to = "Effect_size")
  
vn = expression(paste(""*N*O[3]^"-"))
dose.labs <- as_labeller(c("DOC" = "DOC", "NO3" = "Nitrate"))

bp_fig_data <- bp_long %>%
  mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
   group_by(burn_percentage, response_var, normalization) %>%
  summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>%
  mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect) %>% 
  na.omit()

bp_plot <- bp_fig_data %>%
  ggplot(aes(x = burn_percentage, y = Effect_size, color = normalization)) +
  geom_pointrange(aes(ymin = mean_effect - se_effect, ymax = mean_effect + se_effect),
                  position = position_jitter(width = 0.2, height = 0)) + # Adds jitter
  facet_grid(~response_var, labeller = dose.labs) +
  scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Normalization"),
                     labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
  labs(
    x = "Burn Percentage",
    y = "Effect Size"
  ) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  theme_bw() +
  theme(legend.position = "top")

bp_plot

ggsave("ES_BP_nonnormal_vs_normal.pdf",
       path = here("initial_plots", "Temporal_normalization_vs_nonnormal"),
       width = 7, height = 7,  units = "in")



# # #### 10 % grouping ####
# bp_fig_data_group <- bp_long %>%
#   mutate(grouped_percentage = cut(burn_percentage, breaks = seq(0, 100, by = 10), include.lowest = TRUE, labels = FALSE))
# 
# # make a grouped_percentage
# bp_fig_data_group <- bp_fig_data_group %>% 
#   mutate(grouped_percentage = as.character(grouped_percentage))
# 
# # Make a grouped percentage
# bp_fig_data_group <- bp_fig_data_group %>% 
#   mutate(burn_percentage_group = case_when(grouped_percentage == "1" ~ "0-10",
#                                            grouped_percentage == "2" ~ "11-20",
#                                            grouped_percentage == "3" ~ "21-30",
#                                            grouped_percentage == "4" ~ "31-40",
#                                            grouped_percentage == "5" ~ "41-50",
#                                            grouped_percentage == "6" ~ "51-60",
#                                            grouped_percentage == "7" ~ "61-70",
#                                            grouped_percentage == "8" ~ "71-80",
#                                            grouped_percentage == "9" ~ "81-90",
#                                            grouped_percentage == "10" ~ "91-100"))
#                                            
# # level it so it goes in order 
# bp_fig_data_group$burn_percentage_group <- factor(bp_fig_data_group$burn_percentage_group, levels = c("0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90", "91-100"))
# 
# bp_fig_data_group <- bp_fig_data_group %>% 
#   na.omit(response_var)
# 
# # calculating the mean percent_difference by grouped burn percentage
# mean_group_bp <- bp_fig_data_group %>% 
#   mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
#    group_by(burn_percentage_group, response_var, normalization) %>% 
#    summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
#             sd_effect = sd(Effect_size, na.rm = TRUE),
#             n_effect = n(),
#             Effect_size = mean(Effect_size, na.rm = TRUE)) 
# 
# 
# # %>%
# #   mutate(se_effect = sd_effect / sqrt(n_effect),
# #          lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
# #          upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect)
# 
# burn_percentage_effect_out <- mean_group_bp %>%
#   mutate(across(where(is.character), ~ifelse(is.na(.), "N/A", .))) %>%
#   mutate(across(where(is.numeric), ~ifelse(is.na(.), -9999, .))) %>% 
#   mutate(across(where(is.factor), ~ifelse(is.na(.), -9999, .)))
# 
# # Plot 
# bp_plot_grouped <- burn_percentage_effect_out %>%
#   ggplot(aes(x = burn_percentage_group, y = Effect_size, color = normalization)) +
#   geom_pointrange(aes(ymin = mean_effect - se_effect, ymax = mean_effect + se_effect),
#                   position = position_jitter(width = 0.2, height = 0)) + # Adds jitter
#   facet_grid(~response_var, labeller = dose.labs) +
#   scale_color_manual(values = c("#00AFBB", "#E7B800"),
#                      guide = guide_legend(title = "Normalization"),
#                      labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
#   labs(
#     x = "Burn Percentage",
#     y = "Effect Size"
#   ) +
#   geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
#   theme_bw() +
#   theme(legend.position = "top")
# 
# bp_plot_grouped

```


# BS MODEL 
```{r BS - Normalized & non-normal models}
### BS DOC normalized  ####
# Original workflow 
# data frame: doc_es_dnbr
doc_es_dnbr <- doc_es_dnbr %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size # Effect_size_pair (Site) as a random effect 
doc_bs_mixed_normal <-  lmer(Effect_size ~ mean_dnbr + (1 | Site), data = doc_es_dnbr)

# Check for equal variance
performance::check_heteroscedasticity(doc_bs_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.999).

# Check for normality
residuals <- resid(doc_bs_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.95668, p-value = 0.3097
# when p > 0.05 can assume normality 

doc_bs_mixed_normal

summary(doc_bs_mixed_normal) 

# P-value 
Anova(doc_bs_mixed_normal, type = "III")

# We used a linear mixed-effects model to examine the relationship between burn severity (mean dNBR) and effect size, with site as a random effect to account for spatial clustering of observations. The model included 27 observations across 12 sites.

# Burn severity had a significant negative effect on effect size (β = -0.005, SE = 0.002, t = -2.538, χ² = 6.44, p = 0.011). For each unit increase in mean dNBR (burn severity), the effect size decreased by 0.005 units. The model intercept was also significant (β = 1.606, SE = 0.805, t = 1.995, χ² = 3.98, p = 0.046).

# Random effects showed substantial variation among sites (σ² = 0.622, SD = 0.789), which was considerably larger than the residual variance (σ² = 0.015, SD = 0.122), indicating meaningful site-level differences in effect size that were appropriately accounted for in the mixed-effects framework.


### BS DOC non-normalized ####
# data frame: doc_es_dnbr
# perform linear mixed effect model for non-normalized effect size and site as a random effect
doc_bs_mixed_nn <-  lmer(Effect_size_no_watershed ~ mean_dnbr + (1 | Site), data = doc_es_dnbr)


# Check for equal variance
performance::check_heteroscedasticity(doc_bs_mixed_nn) # Warning: Heteroscedasticity (non-constant error variance) detected (p < .001).

# Check for normality
residuals <- resid(doc_bs_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = W = 0.92746, p-value = 0.05988
# when p > 0.05 can assume normality 

doc_bs_mixed_nn

summary(doc_bs_mixed_nn) 

# P-value 
Anova(doc_bs_mixed_nn, type = "III")

# We used a linear mixed-effects model to examine the relationship between burn severity (mean dNBR) and effect size (excluding watershed effects), with site as a random effect to account for spatial clustering of observations. The model included 27 observations across 12 sites.

# Burn severity showed no significant effect on effect size (β = -9.33 × 10⁻⁵, SE = 7.57 × 10⁻⁴, t = -0.123, χ² = 0.015, p = 0.902). The model intercept was also non-significant (β = 0.298, SE = 0.286, t = 1.041, χ² = 1.083, p = 0.298), indicating no detectable relationship between burn severity and effect size when watershed-level effects were excluded from the response variable.

# Random effects showed moderate variation among sites (σ² = 0.071, SD = 0.267), which was larger than the residual variance (σ² = 0.015, SD = 0.124). This suggests that while there were site-level differences in effect size, the overall effect sizes were smaller and less variable compared to the previous model that included watershed effects.

# Key differences from the previous model:
  #No significant relationships detected
  #Much smaller effect sizes overall (note the scientific notation)
  #Lower between-site variation compared to the watershed-inclusive model
  #Suggests that watershed-level processes may be important drivers of the patterns observed in the first model


### BS NITRATE normalized ####
# Original workflow 
# data frame: no3_es_dnbr
no3_es_dnbr <- no3_es_dnbr %>%
  group_by(Study_ID, Effect_size_pair) %>%
  mutate(Site = cur_group_id()) %>%
  ungroup()

# perform linear mixed effect model for normalized effect size and site as a random effect 
no3_bs_mixed_normal <-  lmer(Effect_size ~ mean_dnbr + (1 | Site), data = no3_es_dnbr)

# Check for equal variance
performance::check_heteroscedasticity(no3_bs_mixed_normal) # OK: Error variance appears to be homoscedastic (p = 0.898).

# Check for normality
residuals <- resid(no3_bs_mixed_normal)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.96035, p-value = 0.0671
# when p > 0.05 can assume normality 

no3_bs_mixed_normal

summary(no3_bs_mixed_normal) 

# P-value 
Anova(no3_bs_mixed_normal, type = "III")

# We used a linear mixed-effects model to examine the relationship between burn severity (mean dNBR) and nitrate effect size, with site as a random effect to account for spatial clustering of observations. The model included 55 observations across 21 sites.

# Burn severity showed no significant effect on nitrate effect size (β = -1.77 × 10⁻⁵, SE = 2.56 × 10⁻³, t = -0.007, χ² < 0.001, p = 0.995). The model intercept was also non-significant (β = 0.394, SE = 0.990, t = 0.398, χ² = 0.159, p = 0.691), indicating no detectable relationship between burn severity and nitrate effect size.

# Random effects showed substantial variation among sites (σ² = 2.364, SD = 1.538), which was considerably larger than the residual variance (σ² = 0.663, SD = 0.814). This indicates meaningful site-level differences in nitrate effect size that were appropriately accounted for in the mixed-effects framework, despite the lack of a burn severity effect.


### BS NITRATE non-normalized ####
# data frame: no3_es_dnbr
# perform linear mixed effect model for non-normalized effect size and site as a random effect 
no3_bs_mixed_nn <-  lmer(Effect_size_no_watershed ~ mean_dnbr + (1 | Site), data = no3_es_dnbr)

# Check for equal variance
performance::check_heteroscedasticity(no3_bs_mixed_nn) # OK: Error variance appears to be homoscedastic (p = 0.238).

# Check for normality
residuals <- resid(no3_bs_mixed_nn)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.93143, p-value = 0.003761
# when p > 0.05 can assume normality 
# NON-NORMAL?

no3_bs_mixed_nn

summary(no3_bs_mixed_nn) 

# P-value 
Anova(no3_bs_mixed_nn, type = "III")

# We used a linear mixed-effects model to examine the relationship between burn severity (mean dNBR) and nitrate effect size (excluding watershed effects), with site as a random effect to account for spatial clustering of observations. The model included 55 observations across 21 sites.

# Burn severity showed no significant effect on nitrate effect size (β = 7.96 × 10⁻⁵, SE = 1.56 × 10⁻³, t = 0.051, χ² = 0.003, p = 0.959). The model intercept was also non-significant (β = 0.809, SE = 0.606, t = 1.337, χ² = 1.787, p = 0.181), indicating no detectable relationship between burn severity and nitrate effect size when watershed-level effects were excluded from the response variable.

# Random effects showed moderate variation among sites (σ² = 0.650, SD = 0.806), which was smaller than the residual variance (σ² = 0.749, SD = 0.865). This pattern contrasts with the watershed-inclusive model, suggesting that watershed-level processes may contribute to site-level variation in nitrate responses.

#Comparison between models:
  # Excluding watershed effects from the nitrate response variable resulted in reduced between-site variation (SD decreased from 1.538 to 0.806) while residual variation increased slightly (SD from 0.814 to 0.865). Unlike the DOC models where watershed inclusion revealed a significant burn severity effect, both nitrate models showed no relationship with burn severity regardless of whether watershed effects were included. This suggests that nitrate responses to fire may be driven by different mechanisms than DOC responses, with burn severity playing a less important role in nitrate dynamics.

```


### BS PLOT NON-NORMAL vs. NORMAL EFFECT SIZE 
```{r BP plot}
# Continuous 
bs_long <- DNBR_model %>% 
  pivot_longer(
    cols= c(Effect_size, Effect_size_no_watershed),
    names_to = "normalization",
    values_to = "Effect_size")
  
vn = expression(paste(""*N*O[3]^"-"))
dose.labs <- as_labeller(c("DOC" = "DOC", "NO3" = "Nitrate"))

bs_fig_data <- bs_long %>%
  mutate(response_var = stringr::str_remove(response_var, "_Interp")) %>%
   group_by(mean_dnbr, response_var, normalization) %>%
  summarise(mean_effect = mean(Effect_size, na.rm = TRUE),
            sd_effect = sd(Effect_size, na.rm = TRUE),
            n_effect = n(),
            Effect_size = mean(Effect_size, na.rm = TRUE)) %>%
  mutate(se_effect = sd_effect / sqrt(n_effect),
         lower_ci_effect = mean_effect - qt(1 - (0.05 / 2), n_effect - 1) * se_effect,
         upper_ci_effect = mean_effect + qt(1 - (0.05 / 2), n_effect - 1) * se_effect) %>% 
  na.omit()

bs_plot <- bs_fig_data %>%
  ggplot(aes(x = mean_dnbr, y = Effect_size, color = normalization)) +
  geom_pointrange(aes(ymin = mean_effect - se_effect, ymax = mean_effect + se_effect),
                  position = position_jitter(width = 0.2, height = 0)) + # Adds jitter
  facet_grid(~response_var, labeller = dose.labs) +
  scale_color_manual(values = c("#00AFBB", "#E7B800"),
                     guide = guide_legend(title = "Normalization"),
                     labels = c('Normalized Effect Size', 'Non-Normalized Effect Size')) +
  labs(
    x = "Burn Severity",
    y = "Effect Size"
  ) +
  geom_hline(yintercept = 0, linewidth = 0.5, color = "red") +
  theme_bw() +
  theme(legend.position = "top")

bs_plot

ggsave("ES_BS_nonnormal_vs_normal.pdf",
       path = here("initial_plots", "Temporal_normalization_vs_nonnormal"),
       width = 7, height = 7,  units = "in")
```



























### OLD DATA and COMPARE 
```{r}
### ORIGINAL DATA MODEL ####
TSF_orig <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "TSF_ES_Fig_data.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, year, Time_Since_Fire, Effect_size_pair, Effect_size)

# Want to make the TSF continuous so I am gonna put actual years to them 
TSF_orig <- TSF_orig %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(
    occurrence = row_number(), # Create an occurrence tracker within Study_ID
    Time_Since_Fire = case_when(
      Study_ID == "Abbott et al. 2021" & occurrence == 1 ~ "10", # First value
      Study_ID == "Abbott et al. 2021" & occurrence == 2 ~ "11", # Second value
      Study_ID == "Abbott et al. 2021" & occurrence == 3 ~ "10", # Third value
      Study_ID == "Abbott et al. 2021" & occurrence == 4 ~ "11", # Fourth value
      TRUE ~ Time_Since_Fire),
    Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>%
  ungroup() %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire))

study_site_data <- read_csv(here("inputs", "catchment_characteristics", "Fire_name_Lat_Long.csv"), na = c('-9999', 'N/A')) %>% 
  dplyr::select(Study_ID, Site, Effect_size_pair) %>% 
  na.omit()

unique(study_site_data$Study_ID) # should be all 18 studies. 

TSF_model <- full_join(TSF_orig, study_site_data) %>% 
  na.omit()# joining the TSF dataframe used to make the figure and study site data to match sites because we will be including site as a random effect

doc_es_tsf_OLD <- TSF_model %>% 
  filter(response_var == "DOC_Interp")

no3_es_tsf <- TSF_model %>% 
  filter(response_var == "NO3_Interp")

# Continuous 
doc_tsf_mixed <-  lmer(Effect_size ~ Time_Since_Fire + (1 | Site), data = doc_es_tsf_OLD)

# Check for equal variance
performance::check_heteroscedasticity(doc_tsf_mixed) # OK: Error variance appears to be homoscedastic (p = 0.976).

# Check for normality
residuals <- resid(doc_tsf_mixed)
qqnorm(residuals)
qqline(residuals)
shapiro.test(residuals) # W = 0.97186, p-value = 0.2639
# when p > 0.05 can assume normality 
 
summary(doc_tsf_mixed) # df: 40.9851, t-value: -4.792, p-value: 2.47e-05 # TSF is significant for DOC 





### COMPARE OLD VS. NEW ####
# Find common Study_IDs to see paired comparisons
common_studies <- intersect(doc_es_tsf$Study_ID, doc_es_tsf_OLD$Study_ID)
print(paste("Common studies:", length(common_studies)))

# Check which studies are unique to each dataset
unique_to_new <- setdiff(doc_es_tsf$Study_ID, doc_es_tsf_OLD$Study_ID)
unique_to_old <- setdiff(doc_es_tsf_OLD$Study_ID, doc_es_tsf$Study_ID)

print("Studies only in current dataset:")
print(unique_to_new)
print("Studies only in OLD dataset:")
print(unique_to_old)

library(ggplot2)

# Create combined dataset for plotting
comparison_data <- data.frame(
  Effect_size = c(doc_es_tsf$Effect_size, doc_es_tsf_OLD$Effect_size),
  Dataset = rep(c("Current", "OLD"), c(nrow(doc_es_tsf), nrow(doc_es_tsf_OLD)))
)

# Boxplot comparison
ggplot(comparison_data, aes(x = Dataset, y = Effect_size)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.6) +
  theme_minimal() +
  labs(title = "Effect Size Comparison Between Datasets")

# Density plot
ggplot(comparison_data, aes(x = Effect_size, fill = Dataset)) +
  geom_density(alpha = 0.6) +
  theme_minimal() +
  labs(title = "Effect Size Distribution Comparison")

# Two-sample t-test
t.test(doc_es_tsf$Effect_size, doc_es_tsf_OLD$Effect_size)

# Welch Two Sample t-test (unequal variances)
t.test(doc_es_tsf$Effect_size, doc_es_tsf_OLD$Effect_size, var.equal = FALSE)

# Non-parametric alternative (Mann-Whitney U test)
wilcox.test(doc_es_tsf$Effect_size, doc_es_tsf_OLD$Effect_size)

OLD_summary <- doc_es_tsf_OLD %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(OLD = n())

NEW_summary <- doc_es_tsf %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(NEW = n())

comp_summary <- left_join(OLD_summary, NEW_summary, by = c("Study_ID", "response_var"))
# Old has 5 more Mast & Clow values. Why? 
# Old has 2 more Murphy et al. 2018 values. Why?

# Create comparison dataframe
effect_size_comparison <- doc_es_tsf %>%
  dplyr::select(Study_ID, Effect_size, Time_Since_Fire) %>%
  rename(Current_Effect_size = Effect_size) %>%
  full_join(
    doc_es_tsf_OLD %>%
      dplyr::select(Study_ID, Effect_size, Time_Since_Fire) %>%
      rename(OLD_Effect_size = Effect_size),
    by = "Study_ID"
  ) %>%
  mutate(
    Match = case_when(
      is.na(Current_Effect_size) | is.na(OLD_Effect_size) ~ FALSE,
      Current_Effect_size == OLD_Effect_size ~ TRUE,
      TRUE ~ FALSE
    )
  )


ES_OLD <- doc_es_tsf_OLD %>% 
  filter(Study_ID == "Mast & Clow; 2008") %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Time_Since_Fire)
# DOUBLE EFFECT SIZE VALUES

ES_NEW <- doc_es_tsf %>% 
  filter(Study_ID == "Mast & Clow; 2008") %>% 
  dplyr::select(Study_ID, response_var, Effect_size_pair, Effect_size, Time_Since_Fire)

```



























# ================================ GRAVEYARD ==================================================#
# Normalized Effect size - orig
```{r Normalized Effect size - orig}
ES_orig <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_Size_UPDATED.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, year, Time_Since_Fire, Effect_size_pair, Effect_size)

```

# TSF original data frame
```{r TSF original data frame}
TSF_orig <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "TSF_ES_Fig_data.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, year, Time_Since_Fire, Effect_size_pair, Effect_size)

```

# Compare TSF orig and ES orig
```{r Compare TSF orig and ES orig}
ES_summary <- ES_orig %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(ES_orig = n())

TSF_summary <- TSF_orig %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(TSF_count = n())

comp_summary_2 <- left_join(ES_summary, TSF_summary, by = c("Study_ID", "response_var"))

# TSF_summary has 10 more observations compared to ES_orig. There is more data in the TSF_summary specifically for Hauer & Spencer and Murphy et al. 2018.

```

# Non-normalized Effect size - new
```{r Non-normalized Effect size - new}
ES_nn <- read_csv(here("Output_for_analysis", "04_Meta_merge_all_studies_effect_size", "archive", "Effect_size_orig_nonnormal.csv")) %>% 
  dplyr::select(Study_ID, response_var, Climate, Biome, Time_Since_Fire, Effect_size_pair, Effect_size, Effect_size_no_watershed)

nn_summary <- ES_nn %>% 
  group_by(Study_ID, response_var) %>% 
  summarise(ES_nn = n())

comp_summary_3 <- left_join(comp_summary_2, nn_summary, by = c("Study_ID", "response_var"))

# Non-normalized effect size matches the ES_Orig here so perhaps we should go with the ES_Orig data frame as our finalized data frame. 

```









# TSF LMER - Normalized (original)
```{r DOC}
### DOC ####
# Read in study site data to get site
study_site_data <- read_csv(here("inputs", "catchment_characteristics", "Fire_name_Lat_Long.csv"), na = c('-9999', 'N/A'))  %>% 
  dplyr::select(Study_ID, Site, Effect_size_pair) %>% 
  na.omit()

# Check to see if all the studies are there.
unique(study_site_data$Study_ID) # should be all 17 studies. 
# good 

# Joining to add the site data because we will be including site as a random effect
ES_orig_model <- full_join(ES_orig, study_site_data) %>% 
  na.omit() 

# Filter data for DOC and nitrate separately
DOC_data <- ES_orig_model %>% 
  filter(response_var == "DOC_Interp")

# Want to make the TSF continuous so I am gonna put actual years to them 
DOC_model <- DOC_data %>% 
  group_by(Study_ID) %>% # Group by Study_ID to handle occurrences within the group
  mutate(Time_Since_Fire = case_when(Study_ID == 'Rhea et al. 2021' ~ '15',
                                     TRUE ~ Time_Since_Fire)) %>% 
  mutate(Time_Since_Fire = as.numeric(Time_Since_Fire))

unique(DOC_model$Study_ID)

```



